### 2024-03-10 paper/paper.qmd
updated the structure of the paper

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Jeongwoo Kim
  - Jiwon Choi
thanks: "Code and data are available at: https://github.com/Kjeongwoo99/STA302H_Paper3"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Survey Data

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

## Post-stratification Data

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-03-10 paper/paper.qmd
Updated 2.1 Survey Data in Data section and references

@@
 
+We have used two datasets for this study.
+One is the U.S election survey data of Democracy Fund + UCLA Nationscape dataset from the Voter Study Group, conducted on October 3, 2019.
+Second is the census data from IPUMS America Census Service, which is used as the post-stratification data for the survey data to adjust the weight.
+
 ## Survey Data
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from from July 2019 to February 2021 (@citeSurveyData).
+The sample is weighted to represent the U.S.
+adult population (@citeSurveyData).
+Nationscape groups weight on the following important factors: gender, the four major census regions, race, Hispanic ethnicity, household income, education, age, language spoken at home, nativity, 2016 presidential vote, and the urban-rural mix of the respondent's ZIP code (@citeSurveyData).
+According to the data, Male make up 48.3% while female make up 51.3% (@citeSurveyData).
+74.2% of the respondents are White, 6.8% are Asian/Pacific, 12% are Black (@citeSurveyData).
+20.4% are those between 18-29, 33.4% are 30-49, 32.4% are 50-69, 3.3% are 70+ (@citeSurveyData).
+On average, 5.1 percent declined immediately among those who are selected for the survey.
+16.7 percent of the respondents did not complete the survey.
+Another 5.9 percent were categorized as speeding or straight-line which means they completed the survey in less than 6 minutes or selected the same response for every question in the three policy question batteries.
+Leaving these out leave 72.4 percent of the original sample for the analysis.
+
+The Nationscape survey's strength lies in its methodological rigor - the effectiveness in collecting large samples from the U.S.
+citizen and its weighting strategy desinged to mirror the U.S.
+adult population by including weight factors such as age, gender, race and income and more.
+As they filter out inaccurate or missing data, it makes sure that the data collected are accurate and ensures data integrity.
+While other datasets such as the General Social Survey (GSS) and the American National Election Studies (ANES) are available, the Nationscape dataset's frequency (surveys collected every week) give it an advantage in analyzing electoral trends and shifts in real-time.
+Its' extensive sample size also justifies the choice of this dataset.
+
+For our analysis, we decided to focus on five demographics: age, gender, education, race and state.
+Age is important because in general, voters tend to become more conservative as they get older.
+To account for the age difference, we divided the age group into four categories: 18-29, 30-49, 50-69 and 70+.
+
+Gender is also an important category because in general, men tend be more conservative and women tend to be more liberal.
+Recently, gender issues are growing social issues and this may affect the election, hence we wanted to explore how this affects our model.
+
+Education is also an interesting factor.
+In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans {@Harris}.
+However, there has been a switch in this trend as 61 percent of non-college white voters showed their support wheres just 45 percent of college-educated white voters did in the exit polls {@Harris}.
+Only 37 percent of those without a degree cast their votes for Democrats while 53 percent with a degree did so {@Harris}.
+We categorized education into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
+
+Race also needs some attention because normally non-white groups are highly in favour of Democrats regardless of candidates and white swing by depending on candidates.
+According to the statistics collected in 2016, 93% of black, 71% of Latino, 68% of Asian support democrats while only 41% of white support democrats {@Prokop}.
+As white voters make up 74% of the voting population, it is really important for both parties to attain this demographic group.
+
+Lastly, states are very important as some states historically favor conservatives while some states vote for democrats.
+In general, the west and the east coasts are democrat supporters whereas south are conservative supporters.
 
@@
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
 ```
@@
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
 ```
@@
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
 ```
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
 ```
@@
 
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-03-10 scripts/02-data_cleaning-post-strat.R
update on age group

@@ -44,8 +44,8 @@ cleaned_data <- raw_data %>%
          age_group = case_when(
            age >= 18 & age <= 29 ~ "18-29",
            age >= 30 & age <= 49 ~ "30-49",
-           age >= 50 & age <= 64 ~ "50-64",
-           age >= 65 ~ "65+",
+           age >= 50 & age <= 64 ~ "50-69",
+           age >= 70 ~ "70+",
            TRUE ~ "Unknown"
          ),
          race = case_when(
@@ -53,14 +53,14 @@ cleaned_data <- raw_data %>%
            race == 2 ~ "Black",
            race == 3 ~ "American Indian",
            race %in% c(4, 5, 6) ~ "Asian",
-           race %in% c(7, 8, 9) ~ "Others",
+           race %in% c(7, 8, 9) ~ "Other",
            TRUE ~ "Unknown"
          ),
          education = case_when(
-           educd %in% c(0:60, 62, 63) ~ "High School or Less",
-           educd %in% c(64, 65:71, 73:81) ~ "Some College",
-           educd == 72 ~ "College Degree",
-           educd >= 82 ~ "Postgrad",
+           educd %in% c(0:60, 62, 63) ~ "High school or less",
+           educd %in% c(64, 65:71, 73:80) ~ "Some college",
+           educd %in% c(81, 82:113) ~ "College degree",
+           educd >= 114 ~ "Postgrad",
            TRUE ~ "Unknown"
          )) %>%
   select(-age, -educd) %>% # Remove the original 'age' and 'educd' columns to keep only the age group and education group

### 2024-03-10 scripts/02-data_cleaning-survey.R

@@ -43,7 +43,7 @@ cleaned_data_survey <- survey_data %>%
       TRUE ~ "Unknown"
     ),
     education = case_when(
-      education %in% 1:4 ~ "Highschool or less",
+      education %in% 1:4 ~ "High school or less",
       education %in% 5:7 ~ "Some college",
       education == 8 ~ "College degree",
       education %in% 9:11 ~ "Postgrad",
@@ -59,10 +59,10 @@ cleaned_data_survey <- survey_data %>%
       TRUE ~ "Unknown"
     ),
     age_group = case_when(
-      age >= 18 & age <= 29 ~ "18~29",
-      age >= 30 & age <= 49 ~ "30~49",
-      age >= 50 & age <= 64 ~ "50~64",
-      age >= 65 ~ "65+",
+      age >= 18 & age <= 29 ~ "18-29",
+      age >= 30 & age <= 49 ~ "30-49",
+      age >= 50 & age <= 69 ~ "50-69",
+      age >= 70 ~ "70+",
       TRUE ~ "Unknown" # Handles any missing or unexpected age values
     ),
     gender = case_when(

### 2024-03-10 scripts/03-test_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Tests the cleaned data from "data/analysis_data"
+# Purpose: Tests the cleaned data sets from "data/analysis_data"
 # Author: Jeongwoo Kim, Jiwon Choi
 # Date: 9 March 2024
 # Contact: jwon.choi@mail.utoronto.ca

### 2024-03-10 paper/paper.qmd
major update on the code for the data section

--- 2024-03-10 paper/paper.qmd
@@
 library(tidyverse)
+library(readr)
 library(palmerpenguins)
+library(ggrepel)
+library(maps)
+library(here)
+library(ggplot2)
+library(dplyr)
+library(knitr)
+library(kableExtra)
+library(statebins)
+
+#### Read in cleaned data ####
+cleaned_data_post_strat <-read_csv(file = here("data/analysis_data/usa_00002_cleaned.csv"), show_col_types = FALSE)
+cleaned_data_survey <- read_csv(file = here("data/analysis_data/ns20191003_cleaned.csv"), show_col_types = FALSE)
 ```
@@
 # Data {#sec-data}
+```{r fig.width=10, fig.height=5}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-age-group
+#| fig-cap: Distribution of Sample and Population by Age Group
+# Read in the survey data
+survey_data <- cleaned_data_survey
+
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+
+# Count the number of occurrences in each age group and calculate percentages for the survey data
+survey_counts <- survey_data %>%
+  group_by(age_group) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Count the number of occurrences in each age group and calculate percentages for the post-strat data
+post_strat_counts <- post_strat_data %>%
+  group_by(age_group) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Combine the percentages into a single data frame for plotting
+combined_percentages <- rbind(
+  data.frame(dataset = 'Survey', age_group = survey_counts$age_group, percentage = survey_counts$percentage),
+  data.frame(dataset = 'Post-Strat', age_group = post_strat_counts$age_group, percentage = post_strat_counts$percentage)
+)
+
+# Plot the percentages with specified bar width and percentage on y-axis, and label each bar
+ggplot(combined_percentages, aes(x = age_group, y = percentage, fill = dataset)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
+  geom_text(aes(label = sprintf("%.1f%%", percentage)),
+            position = position_dodge(width = 0.8), 
+            vjust = -0.3, # Adjust text position
+            color = "black",
+            size = 3) + 
+  labs(x = "Age Group", y = "Proportion", title = "Distribution of Sample and Population by Age Group") +
+  theme_minimal() +
+  scale_fill_manual(values = c("orange", "green"))
+
+```
+
+```{r fig.width=10, fig.height=5}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-gender
+#| fig-cap: Distribution of Sample and Population by Gender
+# Read in the survey data
+survey_data <- cleaned_data_survey
+
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+
+# Calculate percentages for the survey data by gender
+survey_gender <- survey_data %>%
+  group_by(gender) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Calculate percentages for the post-strat data by gender
+post_strat_gender <- post_strat_data %>%
+  group_by(sex) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Combine the gender percentages into a single data frame for plotting
+combined_gender_percentages <- rbind(
+  data.frame(dataset = 'Survey', gender = survey_gender$gender, percentage = survey_gender$percentage),
+  data.frame(dataset = 'Post-Strat', gender = post_strat_gender$sex, percentage = post_strat_gender$percentage)
+)
+
+# Plot the gender percentages with specified bar width and percentage on y-axis, and label each bar
+ggplot(combined_gender_percentages, aes(x = gender, y = percentage, fill = dataset)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
+  geom_text(aes(label = sprintf("%.1f%%", percentage)),
+            position = position_dodge(width = 0.8), 
+            vjust = -0.3, # Adjust text position
+            color = "black",
+            size = 3) + 
+  labs(x = "Gender", y = "Proportion", title = "Distribution of Sample and Population by Gender") +
+  theme_minimal() +
+  scale_fill_manual(values = c("orange", "green"))
+
+```
+
+```{r fig.width=10, fig.height=5}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-race
+#| fig-cap: Distribution of Sample and Population by Race Ethnicity
+# Read in the survey data
+survey_data <- cleaned_data_survey
+
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+
+# Calculate percentages for the survey data by race
+survey_race <- survey_data %>%
+  group_by(race_ethnicity) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Calculate percentages for the post-strat data by race
+post_strat_race <- post_strat_data %>%
+  group_by(race) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Combine the race percentages into a single data frame for plotting
+combined_race_percentages <- rbind(
+  data.frame(dataset = 'Survey', race = survey_race$race_ethnicity, percentage = survey_race$percentage),
+  data.frame(dataset = 'Post-Strat', race = post_strat_race$race, percentage = post_strat_race$percentage)
+)
+
+# Adjust the factor levels for the 'race' column to specify the order
+combined_race_percentages$race <- factor(combined_race_percentages$race, 
+                                                    levels = c("White", "Black", "Asian", "American Indian", "Other"))
+
+# Plot the race percentages with specified bar width and percentage on y-axis, and label each bar
+ggplot(combined_race_percentages, aes(x = race, y = percentage, fill = dataset)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
+  geom_text(aes(label = sprintf("%.1f%%", percentage)),
+            position = position_dodge(width = 0.8), 
+            vjust = -0.3, # Adjust text position
+            color = "black",
+            size = 3) + 
+  labs(x = "Race", y = "Proportion", title = "Distribution of Sample and Population by Race Ethnicity") +
+  theme_minimal() +
+  scale_fill_manual(values = c("orange", "green"))
+```
+
+```{r fig.width=10, fig.height=5}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-education
+#| fig-cap: Distribution of Sample and Population by Education
+# Read in the survey data
+survey_data <- cleaned_data_survey
+
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+
+# Calculate percentages for the survey data by education
+survey_education <- survey_data %>%
+  group_by(education) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Calculate percentages for the post-strat data by education
+post_strat_education <- post_strat_data %>%
+  group_by(education) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Combine the education percentages into a single data frame for plotting
+combined_education_percentages <- rbind(
+  data.frame(dataset = 'Survey', education = survey_education$education, percentage = survey_education$percentage),
+  data.frame(dataset = 'Post-Strat', education = post_strat_education$education, percentage = post_strat_education$percentage)
+)
+
+# Adjust the factor levels for the 'education' column to specify the order
+combined_education_percentages$education <- factor(combined_education_percentages$education, 
+                                                    levels = c("High school or less", "Some college", "College degree", "Postgrad"))
+
+# Plot the education percentages with the adjusted order
+ggplot(combined_education_percentages, aes(x = education, y = percentage, fill = dataset)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
+  geom_text(aes(label = sprintf("%.1f%%", percentage)),
+            position = position_dodge(width = 0.8), 
+            vjust = -0.3, # Adjust text position
+            color = "black",
+            size = 3) + 
+  labs(x = "Education", y = "Proportion", title = "Distribution of Sample and Population by Education") +
+  theme_minimal() +
+  scale_fill_manual(values = c("orange", "green"))
+
+```
 
@@
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-voters-intention-to-vote-for-trump
+#| tbl-cap: Voters Intention to Support Trump
+data <- cleaned_data_survey
 
+# Prepare the data with specified order
+trump_counts <- data %>%
+  mutate(consider_trump = factor(consider_trump, levels = c("Yes", "No", "Other"))) %>%
+  count(consider_trump, name = "Number of Respondents") %>%
+  mutate(`Proportion (%)` = (`Number of Respondents` / sum(`Number of Respondents`)) * 100) %>%
+  mutate(`Proportion (%)` = round(`Proportion (%)`, 2)) %>%
+  rename(`Response` = consider_trump)
+
+# Use kable to create the table
+kable(trump_counts, format = "latex", booktabs = TRUE, align = "c") %>%
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-voters-intention-of-their-primary-party
+#| tbl-cap: Voters Intention of Their Primary Party
+data <- cleaned_data_survey
+
+# Prepare the data with specified order
+party_preference_counts <- data %>%
+  mutate(primary_party = factor(primary_party, levels = c("Democratic", "Republican", "Other"))) %>%
+  count(primary_party, name = "Number of Respondents") %>%
+  mutate(`Proportion (%)` = (`Number of Respondents` / sum(`Number of Respondents`)) * 100) %>%
+  mutate(`Proportion (%)` = round(`Proportion (%)`, 2)) %>%
+  rename(`Party Preference` = primary_party)
+
+# Use kable to create the table
+kable(party_preference_counts, format = "latex", booktabs = TRUE, align = "c") %>%
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
 ```
@@
 Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-
-```
 
@@
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+# analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+# analysis_data |> 
+#  ggplot(aes(x = width, y = length)) +
+#  geom_point(alpha = 0.8) +
+#  theme_minimal() +
+#  labs(x = "Wing width (mm)",
+#       y = "Wing length (mm)")
 ```
@@
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+# first_model <- readRDS(file = here::here("models/first_model.rds"))
 ```
@@
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+# modelsummary::modelsummary(
+#  list(
+#    "First model" = first_model
+#  ),
+#  statistic = "mad",
+#  fmt = 2
+# )
 ```
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
+#pp_check(first_model) +
+#  theme_classic() +
+#  theme(legend.position = "bottom")
 
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
+#posterior_vs_prior(first_model) +
+#  theme_minimal() +
+#  scale_color_brewer(palette = "Set1") +
+#  theme(legend.position = "bottom") +
+#  coord_flip()
 ```
@@
 
-plot(first_model, "trace")
+#plot(first_model, "trace")
 
-plot(first_model, "rhat")
+#plot(first_model, "rhat")
 ```

### 2024-03-10 paper/paper.qmd
Merged the codes and the writings to the master branch. Major changes are 2.1 Survey Data section, 2.2 Post-stratification - graphs and figures, and the reference page.

--- 2024-03-10 paper/paper.qmd
@@
 Education is also an interesting factor.
-In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans {@Harris}.
-However, there has been a switch in this trend as 61 percent of non-college white voters showed their support wheres just 45 percent of college-educated white voters did in the exit polls {@Harris}.
-Only 37 percent of those without a degree cast their votes for Democrats while 53 percent with a degree did so {@Harris}.
+In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans (@Harris).
+However, there has been a switch in this trend as 61 percent of non-college white voters showed their support wheres just 45 percent of college-educated white voters did in the exit polls (@Harris).
+Only 37 percent of those without a degree cast their votes for Democrats while 53 percent with a degree did so (@Harris).
 We categorized education into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
@@
 Race also needs some attention because normally non-white groups are highly in favour of Democrats regardless of candidates and white swing by depending on candidates.
-According to the statistics collected in 2016, 93% of black, 71% of Latino, 68% of Asian support democrats while only 41% of white support democrats {@Prokop}.
+According to the statistics collected in 2016, 93% of black, 71% of Latino, 68% of Asian support democrats while only 41% of white support democrats (@Prokop).
 As white voters make up 74% of the voting population, it is really important for both parties to attain this demographic group.
@@
 In general, the west and the east coasts are democrat supporters whereas south are conservative supporters.
+
+## Post-stratification Data
 
@@
 
-## Post-stratification Data
+```{r fig.width=10, fig.height=5}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-age-group
+#| fig-cap: Distribution of Sample and Population by Age Group
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+# Read in the survey data
+survey_data <- cleaned_data_survey
+
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+
+# Count the number of occurrences in each age group and calculate percentages for the survey data
+survey_counts <- survey_data %>%
+  group_by(age_group) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Count the number of occurrences in each age group and calculate percentages for the post-strat data
+post_strat_counts <- post_strat_data %>%
+  group_by(age_group) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Combine the percentages into a single data frame for plotting
+combined_percentages <- rbind(
+  data.frame(dataset = 'Survey', age_group = survey_counts$age_group, percentage = survey_counts$percentage),
+  data.frame(dataset = 'Post-Strat', age_group = post_strat_counts$age_group, percentage = post_strat_counts$percentage)
+)
+
+# Plot the percentages with specified bar width and percentage on y-axis, and label each bar
+ggplot(combined_percentages, aes(x = age_group, y = percentage, fill = dataset)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
+  geom_text(aes(label = sprintf("%.1f%%", percentage)),
+            position = position_dodge(width = 0.8), 
+            vjust = -0.3, # Adjust text position
+            color = "black",
+            size = 3) + 
+  labs(x = "Age Group", y = "Proportion", title = "Distribution of Sample and Population by Age Group") +
+  theme_minimal() +
+  scale_fill_manual(values = c("orange", "green"))
+```
+
+```{r fig.width=10, fig.height=5}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-gender
+#| fig-cap: Distribution of Sample and Population by Gender
+
+# Read in the survey data
+survey_data <- cleaned_data_survey
+
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+
+# Calculate percentages for the survey data by gender
+survey_gender <- survey_data %>%
+  group_by(gender) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Calculate percentages for the post-strat data by gender
+post_strat_gender <- post_strat_data %>%
+  group_by(sex) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Combine the gender percentages into a single data frame for plotting
+combined_gender_percentages <- rbind(
+  data.frame(dataset = 'Survey', gender = survey_gender$gender, percentage = survey_gender$percentage),
+  data.frame(dataset = 'Post-Strat', gender = post_strat_gender$sex, percentage = post_strat_gender$percentage)
+)
+
+# Plot the gender percentages with specified bar width and percentage on y-axis, and label each bar
+ggplot(combined_gender_percentages, aes(x = gender, y = percentage, fill = dataset)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
+  geom_text(aes(label = sprintf("%.1f%%", percentage)),
+            position = position_dodge(width = 0.8), 
+            vjust = -0.3, # Adjust text position
+            color = "black",
+            size = 3) + 
+  labs(x = "Gender", y = "Proportion", title = "Distribution of Sample and Population by Gender") +
+  theme_minimal() +
+  scale_fill_manual(values = c("orange", "green"))
+```
 
 ```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-voters-intention-to-vote-for-trump
+#| tbl-cap: Voters Intention to Support Trump
+data <- cleaned_data_survey
 
+# Prepare the data with specified order
+trump_counts <- data %>%
+  mutate(consider_trump = factor(consider_trump, levels = c("Yes", "No", "Other"))) %>%
+  count(consider_trump, name = "Number of Respondents") %>%
+  mutate(`Proportion (%)` = (`Number of Respondents` / sum(`Number of Respondents`)) * 100) %>%
+  mutate(`Proportion (%)` = round(`Proportion (%)`, 2)) %>%
+  rename(`Response` = consider_trump)
+
+# Use kable to create the table
+kable(trump_counts, format = "latex", booktabs = TRUE, align = "c") %>%
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
 ```
 
-Talk more about it.
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-voters-intention-of-their-primary-party
+#| tbl-cap: Voters Intention of Their Primary Party
+data <- cleaned_data_survey
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+# Prepare the data with specified order
+party_preference_counts <- data %>%
+  mutate(primary_party = factor(primary_party, levels = c("Democratic", "Republican", "Other"))) %>%
+  count(primary_party, name = "Number of Respondents") %>%
+  mutate(`Proportion (%)` = (`Number of Respondents` / sum(`Number of Respondents`)) * 100) %>%
+  mutate(`Proportion (%)` = round(`Proportion (%)`, 2)) %>%
+  rename(`Party Preference` = primary_party)
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
+# Use kable to create the table
+kable(party_preference_counts, format = "latex", booktabs = TRUE, align = "c") %>%
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
 ```
-
-Talk way more about it.
 
@@
 
+```{=tex}
 \begin{align} 
@@
 \end{align}
-
+```
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 

### 2024-03-10 scripts/post-strat/02-data_cleaning-post-strat.R
major updates and modification on scripts folder

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
+# Purpose: Cleans the raw post-strat data
 # Author: Jeongwoo Kim, Jiwon Choi
 # Date: 9 March 2024
 # Contact: jwon.choi@mail.utoronto.ca

### 2024-03-10 scripts/post-strat/03-test_data-post-strat.R

@@ -1,0 +1,46 @@
+#### Preamble ####
+# Purpose: Tests the cleaned data sets from "data/analysis_data/usa_00002_cleaned.csv", which is the post-strat data
+# Author: Jeongwoo Kim, Jiwon Choi
+# Date: 9 March 2024
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: scripts/01-download_data.R, scripts/post-strat/02-data_cleaning-post-strat.R
+
+#### Workspace setup ####
+library(tidyverse)
+library(lubridate)
+
+### Test for the cleaned post-strat data ###
+cleaned_data <- read_csv("data/analysis_data/usa_00002_cleaned.csv")
+
+# Check if 'stateicp' contains only valid state abbreviations
+all(cleaned_data$stateicp %in% state_abbreviations)
+
+# Check if 'sex' contains only valid categories
+all(cleaned_data$sex %in% sex_categories)
+
+# Check if 'age_group' is categorized correctly
+valid_age_groups <- c("18-29", "30-49", "50-69", "70+")
+all(cleaned_data$age_group %in% valid_age_groups)
+
+# Check if 'race' is categorized correctly
+valid_races <- c("White", "Black", "American Indian", "Asian", "Other")
+all(cleaned_data$race %in% valid_races)
+
+# Check if 'education' is categorized correctly
+valid_education_levels <- c("High school or less", "Some college", "College degree", "Postgrad")
+all(cleaned_data$education %in% valid_education_levels)
+
+# Check for absence of 'Unknown' in key columns
+!any(cleaned_data$sex == "Unknown" | cleaned_data$age_group == "Unknown" | cleaned_data$race == "Unknown" | cleaned_data$education == "Unknown")
+
+# Ensure that original 'age' and 'educd' columns are removed
+!"age" %in% names(cleaned_data)
+!"educd" %in% names(cleaned_data)
+
+# Applying is.character to each column and checking if all are TRUE
+columns_to_check <- c("sex", "stateicp", "race", "education", "age_group")
+all(sapply(cleaned_data[columns_to_check], is.character))
+
+#### Test result ####
+# Result: All TRUE

### 2024-03-10 scripts/survey/00-simulate_data-survey.R

@@ -1,0 +1,7 @@
+#### Preamble ####
+# Purpose: Simulates the survey data set
+# Author: Jeongwoo Kim, Jiwon Choi
+# Date: 9 March 2024
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None

### 2024-03-10 scripts/survey/02-data_cleaning-survey.R

@@ -1,10 +1,10 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
+# Purpose: Cleans the raw survey data
 # Author: Jeongwoo Kim, Jiwon Choi
 # Date: 9 March 2024
 # Contact: jwon.choi@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 01-download_data.R
+# Pre-requisites: scripts/01-download_data.R
 
 #### Workspace setup ####
 library(haven)

### 2024-03-10 scripts/survey/03-test_data-survey.R

@@ -1,0 +1,55 @@
+#### Preamble ####
+# Purpose: Tests the cleaned data sets from "data/analysis_data/ns20191003_cleaned.csv", which is the survey data
+# Author: Jeongwoo Kim, Jiwon Choi
+# Date: 9 March 2024
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: scripts/01-download_data.R, scripts/survey02-data_cleaning-survey.R
+
+### Test for the cleaned survey data ###
+cleaned_data_survey <- read_csv("data/analysis_data/ns20191003_cleaned.csv")
+
+# Check unique values for 'vote_intention'
+valid_vote_intention <- c("Yes", "No", "Other")
+all(cleaned_data_survey$vote_intention %in% valid_vote_intention)
+
+# Check unique values for 'consider_trump'
+valid_consider_trump <- c("Yes", "No", "Other")
+all(cleaned_data_survey$consider_trump %in% valid_consider_trump)
+
+# Check unique values for 'not_trump'
+valid_not_trump <- c("Yes", "No", "Other")
+all(cleaned_data_survey$not_trump %in% valid_not_trump)
+
+# Check unique values for 'primary_party'
+valid_primary_party <- c("Democratic", "Republican", "Other")
+all(cleaned_data_survey$primary_party %in% valid_primary_party)
+
+# Check unique values for 'education'
+valid_education <- c("High school or less", "Some college", "College degree", "Postgrad", "Other")
+all(cleaned_data_survey$education %in% valid_education)
+
+# Check unique values for 'race_ethnicity'
+valid_race_ethnicity <- c("White", "Black", "American Indian", "Asian", "Other")
+all(cleaned_data_survey$race_ethnicity %in% valid_race_ethnicity)
+
+# Check unique values for 'age_group'
+valid_age_group <- c("18-29", "30-49", "50-69", "70+")
+all(cleaned_data_survey$age_group %in% valid_age_group)
+
+# Check unique values for 'gender'
+valid_gender <- c("Female", "Male", "Unknown") # Including 'Unknown' if it was intentionally retained
+all(cleaned_data_survey$gender %in% valid_gender)
+
+# Ensure no 'Unknown' values in key columns
+!any(cleaned_data_survey$vote_intention == "Unknown" |
+       cleaned_data_survey$consider_trump == "Unknown" |
[…]

### 2024-03-11 scripts/survey/00-simulate_data-survey.R
Update 00-simulate_data-survey.R

@@ -5,3 +5,92 @@
 # Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
+
+# Load necessary libraries
+library(ggplot2)
+library(tidyverse)
+
+# Set the seed for reproducibility
+set.seed(123)
+
+# Define the number of samples you want to simulate for vote intention
+n_vote_intention <- 6566
+
+# Define the proportions from your vote intention data
+proportions_vote_intention <- c(Yes = 34.39, No = 55.15, Other = 10.46)
+
+# Adjust the proportions to be out of 1 for vote intention
+proportions_vote_intention <- proportions_vote_intention / sum(proportions_vote_intention)
+
+# Define the responses based on your vote intention categories
+vote_intentions <- c("Yes", "No", "Other")
+
+# Simulate the vote intentions based on the given proportions
+simulated_vote_intentions <- sample(vote_intentions, size = n_vote_intention, replace = TRUE, prob = proportions_vote_intention)
+
+# Calculate the frequency table of the simulated vote intentions
+frequency_table_vote_intention <- table(simulated_vote_intentions)
+
+# Create a tibble of frequencies and proportions for vote intention
+result_table_vote_intention <- tibble(
+  VoteIntention = names(frequency_table_vote_intention),
+  Frequency = as.numeric(frequency_table_vote_intention),
+  Proportion = as.numeric(frequency_table_vote_intention) / sum(frequency_table_vote_intention)
+)
+
+# Print the result table for vote intention
+print(result_table_vote_intention)
+
+# Bar plot for vote intention
+ggplot(data = result_table_vote_intention, aes(x = VoteIntention, y = Frequency, fill = VoteIntention)) +
+  geom_bar(stat = "identity") +
+  theme_minimal() +
+  labs(title = "Bar Plot of Vote Intention to Trump", x = "Vote Intention", y = "Frequency")
+
+# Pie chart for vote intention
+pie(result_table_vote_intention$Frequency, labels = paste(result_table_vote_intention$VoteIntention, ": ", round(result_table_vote_intention$Proportion*100, 2), "%", sep=""),
+    main = "Pie Chart of Vote Intention to Trump")
+
+
+### Simulation 2) Data for the primary party ###
+# Set the seed for reproducibility
+set.seed(123)
+
+# Define the number of samples you want to simulate
+n_primary_party <- 6068
+
+# Define the proportions from your data
[…]

### 2024-03-11 scripts/post-strat/01-data-cleaning-poststratification.R
Updated the README.md, changed the name of simulation/data-cleaning/test files in scripts, removed unnecessary files in paper folder.

@@ -2,7 +2,7 @@
 # Purpose: Cleans the raw post-strat data
 # Author: Jeongwoo Kim, Jiwon Choi
 # Date: 9 March 2024
-# Contact: jwon.choi@mail.utoronto.ca
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 01-download_data.R
 

### 2024-03-11 scripts/survey/01-data-cleaning-survey.R

@@ -2,7 +2,7 @@
 # Purpose: Cleans the raw survey data
 # Author: Jeongwoo Kim, Jiwon Choi
 # Date: 9 March 2024
-# Contact: jwon.choi@mail.utoronto.ca
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: scripts/01-download_data.R
 

### 2024-03-11 scripts/survey/01-data-cleaning-survey.R
Fix age group category for survey data

@@ -59,10 +59,10 @@ cleaned_data_survey <- survey_data %>%
       TRUE ~ "Unknown"
     ),
     age_group = case_when(
-      age >= 18 & age <= 29 ~ "18~29",
-      age >= 30 & age <= 49 ~ "30~49",
-      age >= 50 & age <= 64 ~ "50~64",
-      age >= 65 ~ "65+",
+      age >= 18 & age <= 29 ~ "18-29",
+      age >= 30 & age <= 49 ~ "30-49",
+      age >= 50 & age <= 64 ~ "50-69",
+      age >= 70 ~ "70+",
       TRUE ~ "Unknown" # Handles any missing or unexpected age values
     ),
     gender = case_when(

### 2024-03-11 paper/paper.qmd
fix some errors in several categories

@@
 
+```{r fig.width=10, fig.height=5}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-race
+#| fig-cap: Distribution of Sample and Population by Race Ethnicity
+# Read in the survey data
+survey_data <- cleaned_data_survey
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+# Calculate percentages for the survey data by race
+survey_race <- survey_data %>%
+  group_by(race_ethnicity) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+# Calculate percentages for the post-strat data by race
+post_strat_race <- post_strat_data %>%
+  group_by(race) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+# Combine the race percentages into a single data frame for plotting
+combined_race_percentages <- rbind(
+  data.frame(dataset = 'Survey', race = survey_race$race_ethnicity, percentage = survey_race$percentage),
+  data.frame(dataset = 'Post-Strat', race = post_strat_race$race, percentage = post_strat_race$percentage)
+)
+# Adjust the factor levels for the 'race' column to specify the order
+combined_race_percentages$race <- factor(combined_race_percentages$race, 
+                                                    levels = c("White", "Black", "Asian", "American Indian", "Other"))
+# Plot the race percentages with specified bar width and percentage on y-axis, and label each bar
+ggplot(combined_race_percentages, aes(x = race, y = percentage, fill = dataset)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
+  geom_text(aes(label = sprintf("%.1f%%", percentage)),
+            position = position_dodge(width = 0.8), 
+            vjust = -0.3, # Adjust text position
+            color = "black",
+            size = 3) + 
+  labs(x = "Race", y = "Proportion", title = "Distribution of Sample and Population by Race Ethnicity") +
+  theme_minimal() +
+  scale_fill_manual(values = c("orange", "green"))
+```
+
+```{r fig.width=10, fig.height=5}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-education
+#| fig-cap: Distribution of Sample and Population by Education
+# Read in the survey data
+survey_data <- cleaned_data_survey
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+# Calculate percentages for the survey data by education
+survey_education <- survey_data %>%
+  group_by(education) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+# Calculate percentages for the post-strat data by education
+post_strat_education <- post_strat_data %>%
+  group_by(education) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+# Combine the education percentages into a single data frame for plotting
+combined_education_percentages <- rbind(
+  data.frame(dataset = 'Survey', education = survey_education$education, percentage = survey_education$percentage),
+  data.frame(dataset = 'Post-Strat', education = post_strat_education$education, percentage = post_strat_education$percentage)
+)
+# Adjust the factor levels for the 'education' column to specify the order
+combined_education_percentages$education <- factor(combined_education_percentages$education, 
+                                                    levels = c("High school or less", "Some college", "College degree", "Postgrad"))
+# Plot the education percentages with the adjusted order
+ggplot(combined_education_percentages, aes(x = education, y = percentage, fill = dataset)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
+  geom_text(aes(label = sprintf("%.1f%%", percentage)),
+            position = position_dodge(width = 0.8), 
+            vjust = -0.3, # Adjust text position
+            color = "black",
+            size = 3) + 
+  labs(x = "Education", y = "Proportion", title = "Distribution of Sample and Population by Education") +
+  theme_minimal() +
+  scale_fill_manual(values = c("orange", "green"))
+```
+
 ```{r}

### 2024-03-11 scripts/post-strat/01-data-cleaning-poststratification.R

@@ -44,8 +44,8 @@ cleaned_data <- raw_data %>%
          age_group = case_when(
            age >= 18 & age <= 29 ~ "18-29",
            age >= 30 & age <= 49 ~ "30-49",
-           age >= 50 & age <= 64 ~ "50-64",
-           age >= 65 ~ "65+",
+           age >= 50 & age <= 64 ~ "50-69",
+           age >= 70 ~ "70+",
            TRUE ~ "Unknown"
          ),
          race = case_when(
@@ -53,14 +53,14 @@ cleaned_data <- raw_data %>%
            race == 2 ~ "Black",
            race == 3 ~ "American Indian",
            race %in% c(4, 5, 6) ~ "Asian",
-           race %in% c(7, 8, 9) ~ "Others",
+           race %in% c(3, 7, 8, 9) ~ "Other",
            TRUE ~ "Unknown"
          ),
          education = case_when(
-           educd %in% c(0:60, 62, 63) ~ "High School or Less",
-           educd %in% c(64, 65:71, 73:81) ~ "Some College",
-           educd == 72 ~ "College Degree",
-           educd >= 82 ~ "Postgrad",
+           educd %in% c(0:60, 62, 63) ~ "High school or less",
+           educd %in% c(64, 65:71, 73:80) ~ "Some college",
+           educd %in% c(81, 82:113) ~ "College degree",
+           educd >= 114 ~ "Postgrad",
            TRUE ~ "Unknown"
          )) %>%
   select(-age, -educd) %>% # Remove the original 'age' and 'educd' columns to keep only the age group and education group

### 2024-03-11 scripts/survey/01-data-cleaning-survey.R

@@ -43,7 +43,7 @@ cleaned_data_survey <- survey_data %>%
       TRUE ~ "Unknown"
     ),
     education = case_when(
-      education %in% 1:4 ~ "Highschool or less",
+      education %in% 1:4 ~ "High school or less",
       education %in% 5:7 ~ "Some college",
       education == 8 ~ "College degree",
       education %in% 9:11 ~ "Postgrad",

### 2024-03-11 paper/paper.qmd
updated 2.2 post-stratification data section

@@
 
+IPUMS ("Integrated Public Use Microdata Series") is a website that offers database of samples of the American population from the American Community Surveys of 2000-present.
+These samples provide rich qualitative information on the long-term changes in the population.
+We selected the '' data as the post-stratification dataset for our research.
+
+The variables we have decided to use in this analysis from the dataset are 'sex', 'race', 'stateicp', 'age_group', and 'education'.
+We have filtered out the respondents who have answered 'other' or no available sex data from the raw dataset for simplicity of the analysis.
+Therefore, 'Sex' is categorized into 'Male' and 'Female'.
+'Race' is divided into 'White', 'Black', 'Asian', 'American Indian' and 'other'.
+As mentioned earlier, the fact that white, black and Asian together make up around 93 percent of the U.S population, this categorization is justified.
+'stateicp' is a record of all the states in the U.S.
+We have used the abbreviations for each state, for example, 'CT' for Connecticut, 'ME' for Maine, 'MA' for Massachusetts.
+We have the data of 55 states in total.
+'age_group' categorized the respondents' age into four groups: '18-29', '30-49', '50-64' and '65+'.
+'education' is divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
+We filtered out those unknowns, and cleaned the dataset by assigning each respondent to the right category accordingly.
+
 ```{r}
@@
 ```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 

### 2024-03-11 paper/paper.qmd
updated 2.2 post-stratification data section

@@
 We selected the '' data as the post-stratification dataset for our research.
+The ACS is an ongoing survey that collects data monthly, which is then combined into 1-year, 3-year, and 5-year aggregates.
+It then uses stratified sampling where the U.S population is broken down into sub-groups and initial weights are assigned to each respondent.
+
+One strength of the IPUMS survey is the fact that it provides a comprehensive and extensive data with detailed demographic of the U.S.
+population with social, economic and housing characteristics, which is very useful in our analysis of the 2020 U.S presidential election forecast.
+The longitudinal data of this survey also allows researchers to analyze trends over time.
+The U.S.
+Census Bureau offers credibility of the data with high quality checks.
+The post-stratification process ensures correcting for sampling biases and non-response.
+On the other hand, since the survey relies on self-report, there lies a risk of response bias inherently.
+While it is an ongoing survey, there is still a time lag between the data collection and data availability.
+However, the large sample size, consistency and reliability of the data collection, the integrated data over time with post-stratification can justify the decision to utilize IPUMS data over other sources.
 

### 2024-03-11 scripts/04-model.R
update 04-model.R

@@ -1,37 +1,43 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Model political vote intention using logistic regression.
+# Author: Jeongwoo Kim, Jiwon Choi
+# Date: 9 March 2024
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: need to have properly cleaned survey data
 
 #### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
+library(tidyverse)   # For data manipulation and visualization
+library(rstanarm)    # For Bayesian models including logistic regression
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+survey_data <- read_csv("data/analysis_data/ns20191003_cleaned.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+set.seed(853)
 
+survey_reduced <- 
+  survey_data |> 
+  slice_sample(n = 1000)
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+# Recode 'consider_trump' into a binary variable: 1 for 'Yes', 0 for 'No' and 'Other'
+survey_data$consider_trump_binary <- ifelse(survey_data$consider_trump == "Yes", 1, 0)
+
+# Fit the logistic regression model using the binary outcome variable
+political_preferences <- stan_glm(
+  consider_trump_binary ~ gender + education + age_group + race_ethnicity + state,
+  data = survey_data,
+  family = binomial(link = "logit"),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
[…]

### 2024-03-12 scripts/post-strat/01-data-cleaning-poststratification.R
update category names to match each other

@@ -63,6 +63,7 @@ cleaned_data <- raw_data %>%
            educd >= 114 ~ "Postgrad",
            TRUE ~ "Unknown"
          )) %>%
+  rename(state = stateicp) %>%
   select(-age, -educd) %>% # Remove the original 'age' and 'educd' columns to keep only the age group and education group
   # Filter step to remove rows with "Unknown" in any of the transformed columns
   filter(!if_any(c(sex, age_group, race, education), ~ .x == "Unknown"))

### 2024-03-12 scripts/post-strat/02-test-data-poststratification.R

@@ -13,8 +13,8 @@ library(lubridate)
 ### Test for the cleaned post-strat data ###
 cleaned_data <- read_csv("data/analysis_data/usa_00002_cleaned.csv")
 
-# Check if 'stateicp' contains only valid state abbreviations
-all(cleaned_data$stateicp %in% state_abbreviations)
+# Check if 'state' contains only valid state abbreviations
+all(cleaned_data$state %in% state_abbreviations)
 
 # Check if 'sex' contains only valid categories
 all(cleaned_data$sex %in% sex_categories)
@@ -39,7 +39,7 @@ all(cleaned_data$education %in% valid_education_levels)
 !"educd" %in% names(cleaned_data)
 
 # Applying is.character to each column and checking if all are TRUE
-columns_to_check <- c("sex", "stateicp", "race", "education", "age_group")
+columns_to_check <- c("sex", "state", "race", "education", "age_group")
 all(sapply(cleaned_data[columns_to_check], is.character))
 
 #### Test result ####

### 2024-03-12 scripts/survey/01-data-cleaning-survey.R

@@ -71,9 +71,10 @@ cleaned_data_survey <- survey_data %>%
       TRUE ~ "Unknown" # Handles any other cases as "Unknown"
     )
   ) %>%
+  rename(sex = gender, race = race_ethnicity) %>%
   select(-age) %>%
   # Filter step to remove rows with "Unknown" in any of the transformed columns
-  filter(!if_any(c(vote_intention, consider_trump, not_trump, primary_party, education, state, race_ethnicity, age_group, gender), ~ .x == "Unknown"))
+  filter(!if_any(c(vote_intention, consider_trump, not_trump, primary_party, education, state, race, age_group, sex), ~ .x == "Unknown"))
 
 # Save the cleaned data
 write_csv(cleaned_data_survey, "data/analysis_data/ns20191003_cleaned.csv")

### 2024-03-12 scripts/survey/02-test-data-survey.R

@@ -31,7 +31,7 @@ all(cleaned_data_survey$education %in% valid_education)
 
 # Check unique values for 'race_ethnicity'
 valid_race_ethnicity <- c("White", "Black", "American Indian", "Asian", "Other")
-all(cleaned_data_survey$race_ethnicity %in% valid_race_ethnicity)
+all(cleaned_data_survey$race %in% valid_race_ethnicity)
 
 # Check unique values for 'age_group'
 valid_age_group <- c("18-29", "30-49", "50-69", "70+")
@@ -39,7 +39,7 @@ all(cleaned_data_survey$age_group %in% valid_age_group)
 
 # Check unique values for 'gender'
 valid_gender <- c("Female", "Male", "Unknown") # Including 'Unknown' if it was intentionally retained
-all(cleaned_data_survey$gender %in% valid_gender)
+all(cleaned_data_survey$sex %in% valid_gender)
 
 # Ensure no 'Unknown' values in key columns
 !any(cleaned_data_survey$vote_intention == "Unknown" |
@@ -47,9 +47,9 @@ all(cleaned_data_survey$gender %in% valid_gender)
        cleaned_data_survey$not_trump == "Unknown" |
        cleaned_data_survey$primary_party == "Unknown" |
        cleaned_data_survey$education == "Unknown" |
-       cleaned_data_survey$race_ethnicity == "Unknown" |
+       cleaned_data_survey$race == "Unknown" |
        cleaned_data_survey$age_group == "Unknown" |
-       cleaned_data_survey$gender == "Unknown")
+       cleaned_data_survey$sex == "Unknown")
 
 #### Test result ####
 # Result: All TRUE

### 2024-03-12 scripts/post-strat/04-model-poststratification.R
update model

@@ -1,0 +1,35 @@
+#### Preamble ####
+# Purpose: Model political vote intention using logistic regression.
+# Author: Jeongwoo Kim, Jiwon Choi
+# Date: 9 March 2024
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: need to have properly cleaned censusa data
+
+# Load required libraries
+library(tidyverse)
+library(rstanarm)
+
+# Load the trained logistic regression model
+political_preferences <- readRDS("models/consider_trump.rds")
+
+# Load the Census data (assuming it's been cleaned and optimized)
+census_data <- read_csv("data/analysis_data/usa_00002_cleaned.csv")
+
+
+census_reduced <- 
+  census_data |> 
+  slice_sample(n = 100000)
+
+# Ensure that 'census_data' includes necessary columns and they are optimized for analysis
+
+# Use predict() to apply the model to the Census data
+# Here, type = "response" gives probabilities, which is what we want
+predictions <- predict(political_preferences, newdata = census_reduced, type = "response")
+
+# Add these predictions to the census_data dataframe
+census_reduced$predicted_consider_trump_probability <- predictions
+
+# Save the updated census_data with predictions to an RDS file
+saveRDS(census_reduced, file = "data/census_data_with_predictions.rds")
+

### 2024-03-12 scripts/survey/04-model-survey.R

@@ -1,0 +1,43 @@
+#### Preamble ####
+# Purpose: Model political vote intention using logistic regression.
+# Author: Jeongwoo Kim, Jiwon Choi
+# Date: 9 March 2024
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: need to have properly cleaned survey data
+
+#### Workspace setup ####
+library(tidyverse)   # For data manipulation and visualization
+library(rstanarm)    # For Bayesian models including logistic regression
+
+#### Read data ####
+survey_data <- read_csv("data/analysis_data/ns20191003_cleaned.csv")
+
+set.seed(853)
+
+survey_reduced <- 
+  survey_data |> 
+  slice_sample(n = 1000)
+
+# Recode 'consider_trump' into a binary variable: 1 for 'Yes', 0 for 'No' and 'Other'
+survey_data$consider_trump_binary <- ifelse(survey_data$consider_trump == "Yes", 1, 0)
+
+# Fit the logistic regression model using the binary outcome variable
+political_preferences <- stan_glm(
+  consider_trump_binary ~ sex + education + age_group + race + state,
+  data = survey_data,
+  family = binomial(link = "logit"),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  seed = 853,
+  # Increased iteration values due to warning message that indicates Bulk Effective Samples Size (ESS) 
+  # and Tail Effective Samples Size (ESS) are too low
+  iter = 4000, # Default is usually 2000
+  warmup = 2000 # Half of iter is a common choice
+)
+
+
+saveRDS(
+  political_preferences,
+  file = "models/consider_trump.rds"
+)

### 2024-03-12 scripts/04-model.R
Delete 04-model.R

@@ -1,43 +1,0 @@
-#### Preamble ####
-# Purpose: Model political vote intention using logistic regression.
-# Author: Jeongwoo Kim, Jiwon Choi
-# Date: 9 March 2024
-# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: need to have properly cleaned survey data
-
-#### Workspace setup ####
-library(tidyverse)   # For data manipulation and visualization
-library(rstanarm)    # For Bayesian models including logistic regression
-
-#### Read data ####
-survey_data <- read_csv("data/analysis_data/ns20191003_cleaned.csv")
-
-set.seed(853)
-
-survey_reduced <- 
-  survey_data |> 
-  slice_sample(n = 1000)
-
-# Recode 'consider_trump' into a binary variable: 1 for 'Yes', 0 for 'No' and 'Other'
-survey_data$consider_trump_binary <- ifelse(survey_data$consider_trump == "Yes", 1, 0)
-
-# Fit the logistic regression model using the binary outcome variable
-political_preferences <- stan_glm(
-  consider_trump_binary ~ gender + education + age_group + race_ethnicity + state,
-  data = survey_data,
-  family = binomial(link = "logit"),
-  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-  seed = 853,
-  # Increased iteration values due to warning message that indicates Bulk Effective Samples Size (ESS) 
-  # and Tail Effective Samples Size (ESS) are too low
-  iter = 4000, # Default is usually 2000
-  warmup = 2000 # Half of iter is a common choice
-)
-
-
-saveRDS(
-  political_preferences,
-  file = "models/consider_trump.rds"
-)

### 2024-03-12 scripts/post-strat/04-model-poststratification.R
Update 04-model-poststratification.R

@@ -13,15 +13,14 @@ library(rstanarm)
 # Load the trained logistic regression model
 political_preferences <- readRDS("models/consider_trump.rds")
 
-# Load the Census data (assuming it's been cleaned and optimized)
+# Load the Census data
 census_data <- read_csv("data/analysis_data/usa_00002_cleaned.csv")
 
-
+# reduce the size of data due to memory limit
 census_reduced <- 
   census_data |> 
   slice_sample(n = 100000)
 
-# Ensure that 'census_data' includes necessary columns and they are optimized for analysis
 
 # Use predict() to apply the model to the Census data
 # Here, type = "response" gives probabilities, which is what we want
@@ -31,5 +30,5 @@ predictions <- predict(political_preferences, newdata = census_reduced, type = "
 census_reduced$predicted_consider_trump_probability <- predictions
 
 # Save the updated census_data with predictions to an RDS file
-saveRDS(census_reduced, file = "data/census_data_with_predictions.rds")
+saveRDS(census_reduced, file = "models/census_data_with_predictions.rds")
 

### 2024-03-12 scripts/post-strat/01-data-cleaning-poststratification.R
changed the variable name 'sex' to 'gender'

@@ -63,10 +63,10 @@ cleaned_data <- raw_data %>%
            educd >= 114 ~ "Postgrad",
            TRUE ~ "Unknown"
          )) %>%
-  rename(state = stateicp) %>%
+  rename(state = stateicp, gender = sex, ) %>%
   select(-age, -educd) %>% # Remove the original 'age' and 'educd' columns to keep only the age group and education group
   # Filter step to remove rows with "Unknown" in any of the transformed columns
-  filter(!if_any(c(sex, age_group, race, education), ~ .x == "Unknown"))
+  filter(!if_any(c(gender, age_group, race, education), ~ .x == "Unknown"))
 
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/usa_00002_cleaned.csv")

### 2024-03-12 scripts/post-strat/02-test-data-poststratification.R

@@ -16,8 +16,8 @@ cleaned_data <- read_csv("data/analysis_data/usa_00002_cleaned.csv")
 # Check if 'state' contains only valid state abbreviations
 all(cleaned_data$state %in% state_abbreviations)
 
-# Check if 'sex' contains only valid categories
-all(cleaned_data$sex %in% sex_categories)
+# Check if 'gender' contains only valid categories
+all(cleaned_data$gender %in% sex_categories)
 
 # Check if 'age_group' is categorized correctly
 valid_age_groups <- c("18-29", "30-49", "50-69", "70+")
@@ -32,14 +32,14 @@ valid_education_levels <- c("High school or less", "Some college", "College degr
 all(cleaned_data$education %in% valid_education_levels)
 
 # Check for absence of 'Unknown' in key columns
-!any(cleaned_data$sex == "Unknown" | cleaned_data$age_group == "Unknown" | cleaned_data$race == "Unknown" | cleaned_data$education == "Unknown")
+!any(cleaned_data$gender == "Unknown" | cleaned_data$age_group == "Unknown" | cleaned_data$race == "Unknown" | cleaned_data$education == "Unknown")
 
 # Ensure that original 'age' and 'educd' columns are removed
 !"age" %in% names(cleaned_data)
 !"educd" %in% names(cleaned_data)
 
 # Applying is.character to each column and checking if all are TRUE
-columns_to_check <- c("sex", "state", "race", "education", "age_group")
+columns_to_check <- c("gender", "state", "race", "education", "age_group")
 all(sapply(cleaned_data[columns_to_check], is.character))
 
 #### Test result ####

### 2024-03-12 scripts/survey/01-data-cleaning-survey.R

@@ -71,10 +71,10 @@ cleaned_data_survey <- survey_data %>%
       TRUE ~ "Unknown" # Handles any other cases as "Unknown"
     )
   ) %>%
-  rename(sex = gender, race = race_ethnicity) %>%
+  rename(race = race_ethnicity) %>%
   select(-age) %>%
   # Filter step to remove rows with "Unknown" in any of the transformed columns
-  filter(!if_any(c(vote_intention, consider_trump, not_trump, primary_party, education, state, race, age_group, sex), ~ .x == "Unknown"))
+  filter(!if_any(c(vote_intention, consider_trump, not_trump, primary_party, education, state, race, age_group, gender), ~ .x == "Unknown"))
 
 # Save the cleaned data
 write_csv(cleaned_data_survey, "data/analysis_data/ns20191003_cleaned.csv")

### 2024-03-12 scripts/survey/02-test-data-survey.R

@@ -39,7 +39,7 @@ all(cleaned_data_survey$age_group %in% valid_age_group)
 
 # Check unique values for 'gender'
 valid_gender <- c("Female", "Male", "Unknown") # Including 'Unknown' if it was intentionally retained
-all(cleaned_data_survey$sex %in% valid_gender)
+all(cleaned_data_survey$gender %in% valid_gender)
 
 # Ensure no 'Unknown' values in key columns
 !any(cleaned_data_survey$vote_intention == "Unknown" |
@@ -49,7 +49,7 @@ all(cleaned_data_survey$sex %in% valid_gender)
        cleaned_data_survey$education == "Unknown" |
        cleaned_data_survey$race == "Unknown" |
        cleaned_data_survey$age_group == "Unknown" |
-       cleaned_data_survey$sex == "Unknown")
+       cleaned_data_survey$gender == "Unknown")
 
 #### Test result ####
 # Result: All TRUE

### 2024-03-12 scripts/survey/04-model-survey.R

@@ -24,7 +24,7 @@ survey_data$consider_trump_binary <- ifelse(survey_data$consider_trump == "Yes",
 
 # Fit the logistic regression model using the binary outcome variable
 political_preferences <- stan_glm(
-  consider_trump_binary ~ sex + education + age_group + race + state,
+  consider_trump_binary ~ gender + education + age_group + race + state,
   data = survey_data,
   family = binomial(link = "logit"),
   prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-03-12 paper/paper.qmd
changed codes with variable sex to gender, race_ethnicity to race

@@
 post_strat_gender <- post_strat_data %>%
-  group_by(sex) %>%
+  group_by(gender) %>%
   summarise(count = n()) %>%
@@
   data.frame(dataset = 'Survey', gender = survey_gender$gender, percentage = survey_gender$percentage),
-  data.frame(dataset = 'Post-Strat', gender = post_strat_gender$sex, percentage = post_strat_gender$percentage)
+  data.frame(dataset = 'Post-Strat', gender = post_strat_gender$gender, percentage = post_strat_gender$percentage)
 )
@@
 survey_race <- survey_data %>%
-  group_by(race_ethnicity) %>%
+  group_by(race) %>%
   summarise(count = n()) %>%
@@
 combined_race_percentages <- rbind(
-  data.frame(dataset = 'Survey', race = survey_race$race_ethnicity, percentage = survey_race$percentage),
+  data.frame(dataset = 'Survey', race = survey_race$race, percentage = survey_race$percentage),
   data.frame(dataset = 'Post-Strat', race = post_strat_race$race, percentage = post_strat_race$percentage)

### 2024-03-12 paper/paper.qmd
Update paper.qmd

@@
 
-#### Read in cleaned data ####
+#### Read in necessary data and model ####
 cleaned_data_post_strat <-read_csv(file = here("data/analysis_data/usa_00002_cleaned.csv"), show_col_types = FALSE)
 cleaned_data_survey <- read_csv(file = here("data/analysis_data/ns20191003_cleaned.csv"), show_col_types = FALSE)
+survey_model <- readRDS(file = here("models/consider_trump.rds"))
+predicton_model <- readRDS(file = here("models/census_data_with_predictions.rds"))
 ```
@@
 
-The variables we have decided to use in this analysis from the dataset are 'sex', 'race', 'stateicp', 'age_group', and 'education'.
+The variables we have decided to use in this analysis from the dataset are 'sex', 'race', 'stateicp', 'age_group', and 'educd'.
 We have filtered out the respondents who have answered 'other' or no available sex data from the raw dataset for simplicity of the analysis.
 Therefore, 'Sex' is categorized into 'Male' and 'Female'.
-'Race' is divided into 'White', 'Black', 'Asian', 'American Indian' and 'other'.
+'Race' is divided into 'White', 'Black', 'Asian', 'American Indian' and 'Other'.
 As mentioned earlier, the fact that white, black and Asian together make up around 93 percent of the U.S population, this categorization is justified.
@@
 We filtered out those unknowns, and cleaned the dataset by assigning each respondent to the right category accordingly.
-
-```{r}
-
-```
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+For our study, we employ a technique called multilevel regression with post-stratification (MRP).
+This approach involves creating a model based on a smaller data set, such as our survey data, and then extending the model's findings to a larger population.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#run the regression model using glm.
+# gender, age_groups, race, stateicp, education_level and hispanic are used to 
+# predit whether a person will vote biden or not
+
+model <- survey_model
+
+# Assuming 'model' is your model object from rstanarm
+library(broom.mixed)
+
+coefficients <- broom.mixed::tidy(model, conf.int = TRUE)
+
+# Assuming you have knitr and kableExtra already installed and loaded
+library(knitr)
+library(kableExtra)
+
+knitr::kable(coefficients,
+             caption = "Coefficients from the Model",
+             digits = 2) %>%
+  kableExtra::kable_styling(latex_options = "scale_down")
+
+```
 

### 2024-03-12 paper/paper.qmd
Update paper.qmd

@@
 #| warning: false
-#run the regression model using glm.
-# gender, age_groups, race, stateicp, education_level and hispanic are used to 
-# predit whether a person will vote biden or not
-
 model <- survey_model

### 2024-03-13 paper/paper.tex
commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={My title},
  pdfauthor={Jeongwoo Kim; Jiwon Choi},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{My title\thanks{Code and data are available at:
https://github.com/Kjeongwoo99/STA302H\_Paper3}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{My subtitle if needed}
\author{Jeongwoo Kim \and Jiwon Choi}
\date{March 13, 2024}

\begin{document}
\maketitle
\begin{abstract}
First sentence. Second sentence. Third sentence. Fourth sentence.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, frame hidden, enhanced, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

You can and should cross-reference sections and sub-sections. We use R
Core Team (2023) and Wickham et al. (2019).

The remainder of this paper is structured as follows.
Section~\ref{sec-data}\ldots.

\hypertarget{sec-data}{%
\section{Data}\label{sec-data}}

We have used two datasets for this study. One is the U.S election survey
data of Democracy Fund + UCLA Nationscape dataset from the Voter Study
Group, conducted on October 3, 2019. Second is the census data from
IPUMS America Census Service, which is used as the post-stratification
data for the survey data to adjust the weight.

\hypertarget{survey-data}{%
\subsection{Survey Data}\label{survey-data}}

This survey data is an 18-month election study conducted by UCLA
researchers with roughly 6250 online interviews each from from July 2019
to February 2021 (Tausanovitch and Vavreck (2020)). The sample is
weighted to represent the U.S. adult population (Tausanovitch and
Vavreck (2020)). Nationscape groups weight on the following important
factors: gender, the four major census regions, race, Hispanic
ethnicity, household income, education, age, language spoken at home,
nativity, 2016 presidential vote, and the urban-rural mix of the
respondent's ZIP code (Tausanovitch and Vavreck (2020)). According to
the data, Male make up 48.3\% while female make up 51.3\% (Tausanovitch
and Vavreck (2020)). 74.2\% of the respondents are White, 6.8\% are
Asian/Pacific, 12\% are Black (Tausanovitch and Vavreck (2020)). 20.4\%
are those between 18-29, 33.4\% are 30-49, 32.4\% are 50-69, 3.3\% are
70+ (Tausanovitch and Vavreck (2020)). On average, 5.1 percent declined
immediately among those who are selected for the survey. 16.7 percent of
the respondents did not complete the survey. Another 5.9 percent were
categorized as speeding or straight-line which means they completed the
survey in less than 6 minutes or selected the same response for every
question in the three policy question batteries. Leaving these out leave
72.4 percent of the original sample for the analysis.

The Nationscape survey's strength lies in its methodological rigor - the
effectiveness in collecting large samples from the U.S. citizen and its
weighting strategy desinged to mirror the U.S. adult population by
including weight factors such as age, gender, race and income and more.
As they filter out inaccurate or missing data, it makes sure that the
data collected are accurate and ensures data integrity. While other
datasets such as the General Social Survey (GSS) and the American
National Election Studies (ANES) are available, the Nationscape
dataset's frequency (surveys collected every week) give it an advantage
in analyzing electoral trends and shifts in real-time. Its' extensive
sample size also justifies the choice of this dataset.

For our analysis, we decided to focus on five demographics: age, gender,
education, race and state. Age is important because in general, voters
tend to become more conservative as they get older. To account for the
age difference, we divided the age group into four categories: 18-29,
30-49, 50-69 and 70+.

Gender is also an important category because in general, men tend be
more conservative and women tend to be more liberal. Recently, gender
issues are growing social issues and this may affect the election, hence
we wanted to explore how this affects our model.

Education is also an interesting factor. In the past, non-college white
voters used to support Democrats while college-educated white voters
supported Republicans (Harris (2018)). However, there has been a switch
in this trend as 61 percent of non-college white voters showed their
support wheres just 45 percent of college-educated white voters did in
the exit polls (Harris (2018)). Only 37 percent of those without a
degree cast their votes for Democrats while 53 percent with a degree did
so (Harris (2018)). We categorized education into four categories: `High
school or less', `Some college', `College degree', `Postgrad'.

Race also needs some attention because normally non-white groups are
highly in favour of Democrats regardless of candidates and white swing
by depending on candidates. According to the statistics collected in
2016, 93\% of black, 71\% of Latino, 68\% of Asian support democrats
while only 41\% of white support democrats (Prokop (2021)). As white
voters make up 74\% of the voting population, it is really important for
both parties to attain this demographic group.

Lastly, states are very important as some states historically favor
conservatives while some states vote for democrats. In general, the west
and the east coasts are democrat supporters whereas south are
conservative supporters.

\hypertarget{post-stratification-data}{%
\subsection{Post-stratification Data}\label{post-stratification-data}}

IPUMS (``Integrated Public Use Microdata Series'') is a website that
offers database of samples of the American population from the American
Community Surveys of 2000-present. These samples provide rich
qualitative information on the long-term changes in the population. We
selected the '\,' data as the post-stratification dataset for our
research. The ACS is an ongoing survey that collects data monthly, which
is then combined into 1-year, 3-year, and 5-year aggregates. It then
uses stratified sampling where the U.S population is broken down into
sub-groups and initial weights are assigned to each respondent.

One strength of the IPUMS survey is the fact that it provides a
comprehensive and extensive data with detailed demographic of the U.S.
population with social, economic and housing characteristics, which is
very useful in our analysis of the 2020 U.S presidential election
forecast. The longitudinal data of this survey also allows researchers
to analyze trends over time. The U.S. Census Bureau offers credibility
of the data with high quality checks. The post-stratification process
ensures correcting for sampling biases and non-response. On the other
hand, since the survey relies on self-report, there lies a risk of
response bias inherently. While it is an ongoing survey, there is still
a time lag between the data collection and data availability. However,
the large sample size, consistency and reliability of the data
collection, the integrated data over time with post-stratification can
justify the decision to utilize IPUMS data over other sources.

The variables we have decided to use in this analysis from the dataset
are `sex', `race', `stateicp', `age\_group', and `educd'. We have
filtered out the respondents who have answered `other' or no available
sex data from the raw dataset for simplicity of the analysis. Therefore,
`Sex' is categorized into `Male' and `Female'. `Race' is divided into
`White', `Black', `Asian', `American Indian' and `Other'. As mentioned
earlier, the fact that white, black and Asian together make up around 93
percent of the U.S population, this categorization is justified.
`stateicp' is a record of all the states in the U.S. We have used the
abbreviations for each state, for example, `CT' for Connecticut, `ME'
for Maine, `MA' for Massachusetts. We have the data of 55 states in
total. `age\_group' categorized the respondents' age into four groups:
`18-29', `30-49', `50-64' and `65+'. `education' is divided into four
categories: `High school or less', `Some college', `College degree',
`Postgrad'. We filtered out those unknowns, and cleaned the dataset by
assigning each respondent to the right category accordingly.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-distribution-by-age-group-1.pdf}

}

\caption{\label{fig-distribution-by-age-group}Distribution of Sample and
Population by Age Group}

\end{figure}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-distribution-by-gender-1.pdf}

}

\caption{\label{fig-distribution-by-gender}Distribution of Sample and
Population by Gender}

\end{figure}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-distribution-by-race-1.pdf}

}

\caption{\label{fig-distribution-by-race}Distribution of Sample and
Population by Race Ethnicity}

\end{figure}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-distribution-by-education-1.pdf}

}

\caption{\label{fig-distribution-by-education}Distribution of Sample and
Population by Education}

\end{figure}

\hypertarget{tbl-voters-intention-to-vote-for-trump}{}
\begin{table}[!h]
\caption{\label{tbl-voters-intention-to-vote-for-trump}Voters Intention to Support Trump }\tabularnewline

\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}{ccc}
\toprule
Response & Number of Respondents & Proportion (\%)\\
\midrule
\cellcolor{gray!10}{Yes} & \cellcolor{gray!10}{1908} & \cellcolor{gray!10}{34.25}\\
No & 3080 & 55.30\\
\cellcolor{gray!10}{Other} & \cellcolor{gray!10}{582} & \cellcolor{gray!10}{10.45}\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{tbl-voters-intention-of-their-primary-party}{}
\begin{table}[!h]
\caption{\label{tbl-voters-intention-of-their-primary-party}Voters Intention of Their Primary Party }\tabularnewline

\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}{ccc}
\toprule
Party Preference & Number of Respondents & Proportion (\%)\\
\midrule
\cellcolor{gray!10}{Democratic} & \cellcolor{gray!10}{2180} & \cellcolor{gray!10}{39.14}\\
Republican & 1533 & 27.52\\
\cellcolor{gray!10}{Other} & \cellcolor{gray!10}{1857} & \cellcolor{gray!10}{33.34}\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{model}{%
\section{Model}\label{model}}

For our study, we employ a technique called multilevel regression with
post-stratification (MRP). This approach involves creating a model based
on a smaller data set, such as our survey data, and then extending the
model's findings to a larger population.

\begin{longtable}[t]{lrrrr}
\caption{Coefficients from the Model}\\
\toprule
term & estimate & std.error & conf.low & conf.high\\
\midrule
(Intercept) & -1.13 & 0.82 & -2.59 & 0.19\\
genderMale & 0.65 & 0.06 & 0.55 & 0.74\\
educationHigh school or less & -0.02 & 0.09 & -0.18 & 0.13\\
educationPostgrad & -0.04 & 0.10 & -0.21 & 0.13\\
educationSome college & 0.13 & 0.08 & 0.00 & 0.26\\
\addlinespace
age\_group30-49 & 0.51 & 0.09 & 0.37 & 0.65\\
age\_group50-69 & 0.68 & 0.09 & 0.51 & 0.83\\
age\_group70+ & 0.90 & 0.13 & 0.69 & 1.11\\
raceAsian & -0.79 & 0.30 & -1.28 & -0.27\\
raceBlack & -1.71 & 0.28 & -2.15 & -1.24\\
\addlinespace
raceOther & -0.97 & 0.29 & -1.44 & -0.47\\
raceWhite & 0.19 & 0.25 & -0.21 & 0.61\\
stateAL & 0.18 & 0.83 & -1.15 & 1.66\\
stateAR & -0.07 & 0.83 & -1.44 & 1.40\\
stateAZ & -0.46 & 0.80 & -1.75 & 0.99\\
\addlinespace
stateCA & -0.58 & 0.79 & -1.85 & 0.84\\
stateCO & -0.41 & 0.82 & -1.75 & 1.05\\
stateCT & -0.90 & 0.87 & -2.27 & 0.61\\
stateDC & 0.13 & 0.94 & -1.46 & 1.74\\
stateDE & -0.03 & 0.92 & -1.55 & 1.57\\
\addlinespace
stateFL & 0.02 & 0.80 & -1.26 & 1.46\\
stateGA & 0.23 & 0.81 & -1.06 & 1.67\\
stateHI & -0.40 & 1.05 & -2.21 & 1.37\\
stateIA & -0.28 & 0.85 & -1.68 & 1.22\\
stateID & -0.57 & 0.91 & -2.12 & 1.02\\
\addlinespace
stateIL & -0.45 & 0.79 & -1.74 & 0.98\\
stateIN & -0.11 & 0.80 & -1.38 & 1.32\\
stateKS & -0.92 & 0.87 & -2.32 & 0.61\\
stateKY & 0.19 & 0.82 & -1.12 & 1.67\\
stateLA & 0.51 & 0.85 & -0.83 & 2.01\\
\addlinespace
stateMA & -1.01 & 0.81 & -2.33 & 0.43\\
stateMD & 0.25 & 0.82 & -1.05 & 1.71\\
stateME & 0.33 & 0.90 & -1.13 & 1.89\\
stateMI & -0.62 & 0.81 & -1.92 & 0.83\\
stateMN & -0.17 & 0.83 & -1.50 & 1.30\\
\addlinespace
stateMO & 0.26 & 0.82 & -1.03 & 1.72\\
stateMS & 1.00 & 0.86 & -0.34 & 2.50\\
stateMT & 0.92 & 1.04 & -0.75 & 2.74\\
stateNC & -0.32 & 0.80 & -1.64 & 1.13\\
stateND & -1.38 & 1.15 & -3.42 & 0.51\\
\addlinespace
stateNE & -0.74 & 0.85 & -2.17 & 0.79\\
stateNH & -1.16 & 1.00 & -2.83 & 0.53\\
stateNJ & -0.28 & 0.82 & -1.57 & 1.19\\
stateNM & 0.13 & 0.88 & -1.29 & 1.68\\
stateNV & -0.49 & 0.83 & -1.83 & 0.99\\
\addlinespace
stateNY & -0.40 & 0.79 & -1.69 & 1.03\\
stateOH & -0.28 & 0.80 & -1.55 & 1.17\\
stateOK & -0.55 & 0.82 & -1.89 & 0.91\\
stateOR & -0.61 & 0.84 & -1.98 & 0.88\\
statePA & -0.36 & 0.80 & -1.65 & 1.09\\
\addlinespace
stateRI & -0.87 & 0.96 & -2.46 & 0.80\\
stateSC & 0.05 & 0.83 & -1.26 & 1.50\\
stateSD & -0.13 & 1.00 & -1.73 & 1.58\\
stateTN & 0.08 & 0.81 & -1.22 & 1.55\\
stateTX & -0.11 & 0.79 & -1.37 & 1.34\\
\addlinespace
stateUT & 0.20 & 0.84 & -1.15 & 1.68\\
stateVA & -0.05 & 0.80 & -1.36 & 1.38\\
stateVT & -0.47 & 1.10 & -2.29 & 1.39\\
stateWA & -0.72 & 0.82 & -2.03 & 0.75\\
stateWI & -0.25 & 0.80 & -1.56 & 1.20\\
\addlinespace
stateWV & -0.16 & 0.88 & -1.56 & 1.36\\
stateWY & 0.82 & 1.03 & -0.85 & 2.57\\
\bottomrule
\end{longtable}

\hypertarget{model-set-up}{%
\subsection{Model set-up}\label{model-set-up}}

Define \(y_i\) as the number of seconds that the plane remained aloft.
Then \(\beta_i\) is the wing width and \(\gamma_i\) is the wing length,
both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R (R Core Team 2023) using the \texttt{rstanarm}
package of Goodrich et al. (2022). We use the default priors from
\texttt{rstanarm}.

\hypertarget{results}{%
\section{Results}\label{results}}

\begin{table}

\end{table}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/stateplot-1.pdf}

}

\caption{Proportion of each State Voting for Trump}

\end{figure}

Figure @ref(fig:stateplot) shows us the estimated of proportion of
support for Trump and Biden by state using MRP with the inclusion of
error terms. Each dot represents the point estimate of the proportion of
support for Biden (blue) or Trump (red) in each state. Horizontal Lines
extending from the dots represent confidence intervals for these
estimates. The length of each line indicates the uncertainty associated
with each estimate. For instance, we can see that this uncertainty lies
between 50 percent to slightly higher than 80 percent for Trump in MT
(Massachusetts). The dashed green line in the middle at the 50 percent
mark represents the threshold for majority support. On the y-axis, each
state is listed with its abbreviations and is ordered based on the
proportion of support for Trump from the highest at the top to the
lowest at the bottom.

From figure @ref(fig:stateplot), it seems that majority of the states
support Biden. Only 7 states out of 51 have its point estimate greater
than 50 percent for Trump. However, the horizontal lines of confidence
intervals of some states overlapping the green mark give some hope for
the Republicans. However, excluding these contesting states, our model
suggests that only 3 states are definitely in favor of Trump whereas 35
states are definitely supporting Biden.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/educationplot-1.pdf}

}

\caption{How different education levels of voters affect voting for
Trump}

\end{figure}

Figure @ref(fig:educationplot) presents the estimated proportion of
voters for Trump by education level, divided into four categories: `High
school or less', `Some college', `College degree', `Postgrad'. Each
black dot represents the point estimate of the proportion of voters
within the corresponding education category who are predicted to vote
for Trump. The horizontal lines extending to the left and right of each
dot represent the confidence intervals around the estimate, which
reflect the uncertainty.

It shows that regardless of education level, the level of support for
Trump lies below 40 percent. The Republican party does not have
majority, including the error bars across all education levels. Voters
with ``High school or less'' education appear to have the lowest
estimated support for Trump, which does not align with various exit
polls and analyses from the 2020 election suggesting that Trump had
substantial support among voters without a college degree. Conversely,
Voters with `Some college' and `Postgrad' education are the two groups
that are more in support of Trump, which is exactly the opposite of what
we have expected.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{sec-first-point}{%
\subsection{First discussion point}\label{sec-first-point}}

\hypertarget{second-discussion-point}{%
\subsection{Second discussion point}\label{second-discussion-point}}

\hypertarget{third-discussion-point}{%
\subsection{Third discussion point}\label{third-discussion-point}}

\hypertarget{weaknesses-and-next-steps}{%
\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}}

Weaknesses and next steps should also be included.

\newpage

\appendix

\hypertarget{appendix}{%
\section*{Appendix}\label{appendix}}
\addcontentsline{toc}{section}{Appendix}

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-rstanarm}{}}%
Goodrich, Ben, Jonah Gabry, Imad Ali, and Sam Brilleman. 2022.
{``Rstanarm: {Bayesian} Applied Regression Modeling via {Stan}.''}
\url{https://mc-stan.org/rstanarm/}.

\leavevmode\vadjust pre{\hypertarget{ref-Harris}{}}%
Harris, Adam. 2018. {``AMERICA IS DIVIDED BY EDUCATION.''} \emph{The
Atlantic}.
\url{https://www.theatlantic.com/education/archive/2018/11/education-gap-explains-american-politics/575113/}.

\leavevmode\vadjust pre{\hypertarget{ref-Prokop}{}}%
Prokop, Andrew. 2021. {``A New Report Complicates Simplistic Narratives
about Race and the 2020 Election.''} \emph{Vox}, May.
\url{https://www.vox.com/2021/5/10/22425178/catalist-report-2020-election-biden-trump-demographics}.

\leavevmode\vadjust pre{\hypertarget{ref-citeR}{}}%
R Core Team. 2023. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-citeSurveyData}{}}%
Tausanovitch, Chris, and Lynn Vavreck. 2020. \emph{Democracy Fund + UCLA
Nationscape}. October 10-17, 2019 (version 20200814).
\url{https://www.voterstudygroup.org/publication/nationscape-data-set}.

\leavevmode\vadjust pre{\hypertarget{ref-rohan}{}}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\end{CSLReferences}

\end{document}

### 2024-03-13 paper/paper.qmd
changes

--- 2024-03-12 paper/paper.qmd
@@
 
-#### Read in cleaned data ####
+#### Read in necessary data and model ####
 cleaned_data_post_strat <-read_csv(file = here("data/analysis_data/usa_00002_cleaned.csv"), show_col_types = FALSE)
 cleaned_data_survey <- read_csv(file = here("data/analysis_data/ns20191003_cleaned.csv"), show_col_types = FALSE)
+
+survey_model <- readRDS(file = here("models/consider_trump.rds"))
+predicton_model <- readRDS(file = here("models/census_data_with_predictions.rds"))
 ```
@@
 
-This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from from July 2019 to February 2021 (@citeSurveyData).
+This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from from July 2019 to February 2021 @citeSurveyData.
 The sample is weighted to represent the U.S.
-adult population (@citeSurveyData).
-Nationscape groups weight on the following important factors: gender, the four major census regions, race, Hispanic ethnicity, household income, education, age, language spoken at home, nativity, 2016 presidential vote, and the urban-rural mix of the respondent's ZIP code (@citeSurveyData).
-According to the data, Male make up 48.3% while female make up 51.3% (@citeSurveyData).
-74.2% of the respondents are White, 6.8% are Asian/Pacific, 12% are Black (@citeSurveyData).
-20.4% are those between 18-29, 33.4% are 30-49, 32.4% are 50-69, 3.3% are 70+ (@citeSurveyData).
+adult population @citeSurveyData.
+Nationscape groups weight on the following important factors: gender, the four major census regions, race, Hispanic ethnicity, household income, education, age, language spoken at home, nativity, 2016 presidential vote, and the urban-rural mix of the respondent's ZIP code @citeSurveyData.
+According to the data, Male make up 48.3% while female make up 51.3% @citeSurveyData.
+74.2% of the respondents are White, 6.8% are Asian/Pacific, 12% are Black @citeSurveyData.
+20.4% are those between 18-29, 33.4% are 30-49, 32.4% are 50-69, 3.3% are 70+ @citeSurveyData.
 On average, 5.1 percent declined immediately among those who are selected for the survey.
@@
 Education is also an interesting factor.
-In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans (@Harris).
-However, there has been a switch in this trend as 61 percent of non-college white voters showed their support wheres just 45 percent of college-educated white voters did in the exit polls (@Harris).
-Only 37 percent of those without a degree cast their votes for Democrats while 53 percent with a degree did so (@Harris).
+In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans @Harris.
+However, there has been a switch in this trend as 61 percent of non-college white voters showed their support wheres just 45 percent of college-educated white voters did in the exit polls @Harris.
+Only 37 percent of those without a degree cast their votes for Democrats while 53 percent with a degree did so @Harris.
 We categorized education into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
@@
 Race also needs some attention because normally non-white groups are highly in favour of Democrats regardless of candidates and white swing by depending on candidates.
-According to the statistics collected in 2016, 93% of black, 71% of Latino, 68% of Asian support democrats while only 41% of white support democrats (@Prokop).
+According to the statistics collected in 2016, 93% of black, 71% of Latino, 68% of Asian support democrats while only 41% of white support democrats @Prokop.
 As white voters make up 74% of the voting population, it is really important for both parties to attain this demographic group.
@@
 
-```{r fig.width=10, fig.height=5}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-distribution-by-gender
-#| fig-cap: Distribution of Sample and Population by Gender
-
-# Read in the survey data
-survey_data <- cleaned_data_survey
-
-# Read in the post-stratification data
-post_strat_data <- cleaned_data_post_strat
-
-# Calculate percentages for the survey data by gender
-survey_gender <- survey_data %>%
-  group_by(gender) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
-
-# Calculate percentages for the post-strat data by gender
-post_strat_gender <- post_strat_data %>%
-  group_by(gender) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
-
-# Combine the gender percentages into a single data frame for plotting
-combined_gender_percentages <- rbind(
-  data.frame(dataset = 'Survey', gender = survey_gender$gender, percentage = survey_gender$percentage),
-  data.frame(dataset = 'Post-Strat', gender = post_strat_gender$gender, percentage = post_strat_gender$percentage)
-)
-
-# Plot the gender percentages with specified bar width and percentage on y-axis, and label each bar
-ggplot(combined_gender_percentages, aes(x = gender, y = percentage, fill = dataset)) +
-  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
-  geom_text(aes(label = sprintf("%.1f%%", percentage)),
-            position = position_dodge(width = 0.8), 
-            vjust = -0.3, # Adjust text position
-            color = "black",
-            size = 3) + 
-  labs(x = "Gender", y = "Proportion", title = "Distribution of Sample and Population by Gender") +
-  theme_minimal() +
-  scale_fill_manual(values = c("orange", "green"))
-```
-
-```{r fig.width=10, fig.height=5}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-distribution-by-race
-#| fig-cap: Distribution of Sample and Population by Race Ethnicity
-# Read in the survey data
-survey_data <- cleaned_data_survey
-# Read in the post-stratification data
-post_strat_data <- cleaned_data_post_strat
-# Calculate percentages for the survey data by race
-survey_race <- survey_data %>%
-  group_by(race) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
-# Calculate percentages for the post-strat data by race
-post_strat_race <- post_strat_data %>%
-  group_by(race) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
-# Combine the race percentages into a single data frame for plotting
-combined_race_percentages <- rbind(
-  data.frame(dataset = 'Survey', race = survey_race$race, percentage = survey_race$percentage),
-  data.frame(dataset = 'Post-Strat', race = post_strat_race$race, percentage = post_strat_race$percentage)
-)
-# Adjust the factor levels for the 'race' column to specify the order
-combined_race_percentages$race <- factor(combined_race_percentages$race, 
-                                                    levels = c("White", "Black", "Asian", "American Indian", "Other"))
-# Plot the race percentages with specified bar width and percentage on y-axis, and label each bar
-ggplot(combined_race_percentages, aes(x = race, y = percentage, fill = dataset)) +
-  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
-  geom_text(aes(label = sprintf("%.1f%%", percentage)),
-            position = position_dodge(width = 0.8), 
-            vjust = -0.3, # Adjust text position
-            color = "black",
-            size = 3) + 
-  labs(x = "Race", y = "Proportion", title = "Distribution of Sample and Population by Race Ethnicity") +
-  theme_minimal() +
-  scale_fill_manual(values = c("orange", "green"))
-```
-
-```{r fig.width=10, fig.height=5}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-distribution-by-education
-#| fig-cap: Distribution of Sample and Population by Education
-# Read in the survey data
-survey_data <- cleaned_data_survey
-# Read in the post-stratification data
-post_strat_data <- cleaned_data_post_strat
-# Calculate percentages for the survey data by education
-survey_education <- survey_data %>%
-  group_by(education) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
-# Calculate percentages for the post-strat data by education
-post_strat_education <- post_strat_data %>%
-  group_by(education) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
-# Combine the education percentages into a single data frame for plotting
-combined_education_percentages <- rbind(
-  data.frame(dataset = 'Survey', education = survey_education$education, percentage = survey_education$percentage),
-  data.frame(dataset = 'Post-Strat', education = post_strat_education$education, percentage = post_strat_education$percentage)
-)
-# Adjust the factor levels for the 'education' column to specify the order
-combined_education_percentages$education <- factor(combined_education_percentages$education, 
-                                                    levels = c("High school or less", "Some college", "College degree", "Postgrad"))
-# Plot the education percentages with the adjusted order
-ggplot(combined_education_percentages, aes(x = education, y = percentage, fill = dataset)) +
-  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
-  geom_text(aes(label = sprintf("%.1f%%", percentage)),
-            position = position_dodge(width = 0.8), 
-            vjust = -0.3, # Adjust text position
-            color = "black",
-            size = 3) + 
-  labs(x = "Education", y = "Proportion", title = "Distribution of Sample and Population by Education") +
-  theme_minimal() +
-  scale_fill_manual(values = c("orange", "green"))
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-voters-intention-to-vote-for-trump
-#| tbl-cap: Voters Intention to Support Trump
-data <- cleaned_data_survey
-
-# Prepare the data with specified order
-trump_counts <- data %>%
-  mutate(consider_trump = factor(consider_trump, levels = c("Yes", "No", "Other"))) %>%
-  count(consider_trump, name = "Number of Respondents") %>%
-  mutate(`Proportion (%)` = (`Number of Respondents` / sum(`Number of Respondents`)) * 100) %>%
-  mutate(`Proportion (%)` = round(`Proportion (%)`, 2)) %>%
-  rename(`Response` = consider_trump)
-
-# Use kable to create the table
-kable(trump_counts, format = "latex", booktabs = TRUE, align = "c") %>%
-  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-voters-intention-of-their-primary-party
-#| tbl-cap: Voters Intention of Their Primary Party
-data <- cleaned_data_survey
-
-# Prepare the data with specified order
-party_preference_counts <- data %>%
-  mutate(primary_party = factor(primary_party, levels = c("Democratic", "Republican", "Other"))) %>%
-  count(primary_party, name = "Number of Respondents") %>%
-  mutate(`Proportion (%)` = (`Number of Respondents` / sum(`Number of Respondents`)) * 100) %>%
-  mutate(`Proportion (%)` = round(`Proportion (%)`, 2)) %>%
-  rename(`Party Preference` = primary_party)
-
-# Use kable to create the table
-kable(party_preference_counts, format = "latex", booktabs = TRUE, align = "c") %>%
-  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
-```
-
 # Model
@@
 
-```{=tex}
 \begin{align} 
@@
 \end{align}
-```
+
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
@@
 
-# Appendix {.unnumbered}
+# Appendix {-}
 

### 2024-03-13 scripts/01-download_data.R

@@ -1,10 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data 
-# Author: Jeongwoo Kim, Jiwon Choi
-# Date: 8 March 2024
-# Contact: jwon.choi@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: 00-simulation-survey.R, 00-simulation-poststratification.R
-
-# We use two datasets, one is the survey data from Voter Study Group, and the other is ACS data from IPUMS USA.
-# We cannot share these files, but the steps we took to obtain them are documented in the README.md file.

### 2024-03-13 scripts/03-test_data.R

@@ -1,93 +1,0 @@
-#### Preamble ####
-# Purpose: Tests the cleaned data from "data/analysis_data"
-# Author: Jeongwoo Kim, Jiwon Choi
-# Date: 9 March 2024
-# Contact: jwon.choi@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: 01-download_data.R, 02-data_cleaning-post-strat.R, 01-data_cleaning-survey.R
-
-#### Workspace setup ####
-library(tidyverse)
-library(lubridate)
-
-### Test for the cleaned post-strat data ###
-cleaned_data <- read_csv("data/analysis_data/usa_00002_cleaned.csv")
-
-# Check if 'stateicp' contains only valid state abbreviations
-all(cleaned_data$stateicp %in% state_abbreviations)
-
-# Check if 'sex' contains only valid categories
-all(cleaned_data$sex %in% sex_categories)
-
-# Check if 'age_group' is categorized correctly
-valid_age_groups <- c("18-29", "30-49", "50-64", "65+")
-all(cleaned_data$age_group %in% valid_age_groups)
-
-# Check if 'race' is categorized correctly
-valid_races <- c("White", "Black", "American Indian", "Asian", "Others")
-all(cleaned_data$race %in% valid_races)
-
-# Check if 'education' is categorized correctly
-valid_education_levels <- c("High School or Less", "Some College", "College Degree", "Postgrad")
-all(cleaned_data$education %in% valid_education_levels)
-
-# Check for absence of 'Unknown' in key columns
-!any(cleaned_data$sex == "Unknown" | cleaned_data$age_group == "Unknown" | cleaned_data$race == "Unknown" | cleaned_data$education == "Unknown")
-
-# Ensure that original 'age' and 'educd' columns are removed
-!"age" %in% names(cleaned_data)
-!"educd" %in% names(cleaned_data)
-
-# Applying is.character to each column and checking if all are TRUE
-columns_to_check <- c("sex", "stateicp", "race", "education", "age_group")
-all(sapply(cleaned_data[columns_to_check], is.character))
-
-
-
-### Test for the cleaned survey data ###
-cleaned_data_survey <- read_csv("data/analysis_data/ns20191003_cleaned.csv")
-
-# Check unique values for 'vote_intention'
-valid_vote_intention <- c("Yes", "No", "Other")
-all(cleaned_data_survey$vote_intention %in% valid_vote_intention)
-
[…]

### 2024-03-13 scripts/post-strat/00-simulation-poststratification.R

@@ -1,0 +1,93 @@
+#### Preamble ####
+# Purpose: Simulates the post-strat data set
+# Author: Jeongwoo Kim, Jiwon Choi
+# Date: 9 March 2024
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+#### Workspace setup ####
+# Load necessary libraries
+library(ggplot2)
+library(dplyr)
+library(tibble)
+
+### Simulation 1) Simulate data for age group distribution ###
+set.seed(123) # Setting seed for reproducibility
+age_groups <- c('18-29', '30-49', '50-69', '70+')
+base_proportion <- c(50, 75, 25, 30) # These are the base proportions for the age groups
+
+simulated_data1 <-
+  tibble(
+    # Create age group data
+    age_group = rep(age_groups, times = 1),
+    # Base proportion data
+    proportion = base_proportion + rnorm(length(age_groups), 0, 5)
+  )
+
+# Create a bar plot with custom styles using ggplot
+plot1 <- ggplot(simulated_data1, aes(x = age_group, y = proportion)) +
+  geom_bar(stat = "identity", position = position_dodge(), fill = "blue") + # Set bar color to blue
+  labs(title = "Distribution of Sample by Age Group",
+       x = "Age Group", y = "Proportion") +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5), # Center the plot title
+        axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Display the plot
+print(plot1)
+
+### Simulation 2) Simulate data for gender distribution ###
+set.seed(123) # Setting seed for reproducibility
+
+# Define gender proportions with some added random variation
+gender_names <- c('male', 'female')
+base_proportions <- c(45, 55) # Base proportions
+
+simulated_data2 <- 
+  tibble(
+    gender = factor(gender_names, levels = gender_names),
+    proportion = base_proportions + runif(length(gender_names), min = -5, max = 5) # Add random variation
+  )
+
+# Create a bar plot with all bars colored blue
+plot2 <- ggplot(simulated_data2, aes(x = gender, y = proportion, fill = gender)) +
+  geom_bar(stat = "identity", color = "black") +
+  scale_fill_manual(values = rep("blue", length(gender_names))) + # Set all bars to blue
+  labs(title = "Distribution of Sample by Gender",
+       x = "Gender", y = "Proportion") +
+  theme_minimal() +
[…]

### 2024-03-13 scripts/post-strat/01-data-cleaning-poststratification.R

@@ -1,8 +1,8 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
+# Purpose: Cleans the raw post-strat data
 # Author: Jeongwoo Kim, Jiwon Choi
 # Date: 9 March 2024
-# Contact: jwon.choi@mail.utoronto.ca
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 01-download_data.R
 
@@ -44,8 +44,8 @@ cleaned_data <- raw_data %>%
          age_group = case_when(
            age >= 18 & age <= 29 ~ "18-29",
            age >= 30 & age <= 49 ~ "30-49",
-           age >= 50 & age <= 64 ~ "50-64",
-           age >= 65 ~ "65+",
+           age >= 50 & age <= 64 ~ "50-69",
+           age >= 70 ~ "70+",
            TRUE ~ "Unknown"
          ),
          race = case_when(
@@ -53,20 +53,20 @@ cleaned_data <- raw_data %>%
            race == 2 ~ "Black",
            race == 3 ~ "American Indian",
            race %in% c(4, 5, 6) ~ "Asian",
-           race %in% c(7, 8, 9) ~ "Others",
+           race %in% c(3, 7, 8, 9) ~ "Other",
            TRUE ~ "Unknown"
          ),
          education = case_when(
-           educd %in% c(0:60, 62, 63) ~ "High School or Less",
-           educd %in% c(64, 65:71, 73:81) ~ "Some College",
-           educd == 72 ~ "College Degree",
-           educd >= 82 ~ "Postgrad",
+           educd %in% c(0:60, 62, 63) ~ "High school or less",
+           educd %in% c(64, 65:71, 73:80) ~ "Some college",
+           educd %in% c(81, 82:113) ~ "College degree",
+           educd >= 114 ~ "Postgrad",
            TRUE ~ "Unknown"
          )) %>%
+  rename(state = stateicp, gender = sex, ) %>%
   select(-age, -educd) %>% # Remove the original 'age' and 'educd' columns to keep only the age group and education group
   # Filter step to remove rows with "Unknown" in any of the transformed columns
-  filter(!if_any(c(sex, age_group, race, education), ~ .x == "Unknown"))
[…]

### 2024-03-13 scripts/post-strat/04-model-poststratification.R

@@ -1,0 +1,33 @@
+#### Preamble ####
+# Purpose: Model political vote intention using logistic regression.
+# Author: Jeongwoo Kim, Jiwon Choi
+# Date: 9 March 2024
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: need to have properly cleaned censusa data
+
+# Load required libraries
+library(tidyverse)
+library(rstanarm)
+
+# Load the trained logistic regression model
+political_preferences <- readRDS("models/consider_trump.rds")
+
+# Load the Census data
+census_data <- read_csv("data/analysis_data/usa_00002_cleaned.csv")
+
+# reduce the size of data due to memory limit
+census_reduced <- 
+  census_data |> 
+  slice_sample(n = 100000)
+
+
+# Use predict() to apply the model to the Census data
+# Here, type = "response" gives probabilities, which is what we want
+predictions <- predict(political_preferences, newdata = census_reduced, type = "response")
+
+# Add these predictions to the census_data dataframe
+census_reduced$predicted_consider_trump_probability <- predictions
+
+# Save the updated census_data with predictions to an RDS file
+saveRDS(census_reduced, file = "models/census_data_with_predictions.rds")

### 2024-03-13 scripts/survey/00-simulation-survey.R

@@ -1,0 +1,95 @@
+#### Preamble ####
+# Purpose: Simulates the survey data set
+# Author: Jeongwoo Kim, Jiwon Choi
+# Date: 9 March 2024
+# Contact: jwoo.kim@mail.utoronto.ca, jwon.choi@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+# Load necessary libraries
+library(ggplot2)
+library(tidyverse)
+
+# Set the seed for reproducibility
+set.seed(123)
+
+# Define the number of samples you want to simulate for vote intention
+n_vote_intention <- 6566
+
+# Define the proportions from your vote intention data
+proportions_vote_intention <- c(Yes = 34.39, No = 55.15, Other = 10.46)
+
+# Adjust the proportions to be out of 1 for vote intention
+proportions_vote_intention <- proportions_vote_intention / sum(proportions_vote_intention)
+
+# Define the responses based on your vote intention categories
+vote_intentions <- c("Yes", "No", "Other")
+
+# Simulate the vote intentions based on the given proportions
+simulated_vote_intentions <- sample(vote_intentions, size = n_vote_intention, replace = TRUE, prob = proportions_vote_intention)
+
+# Calculate the frequency table of the simulated vote intentions
+frequency_table_vote_intention <- table(simulated_vote_intentions)
+
+# Create a tibble of frequencies and proportions for vote intention
+result_table_vote_intention <- tibble(
+  VoteIntention = names(frequency_table_vote_intention),
+  Frequency = as.numeric(frequency_table_vote_intention),
+  Proportion = as.numeric(frequency_table_vote_intention) / sum(frequency_table_vote_intention)
+)
+
+# Print the result table for vote intention
+print(result_table_vote_intention)
+
+# Bar plot for vote intention
+ggplot(data = result_table_vote_intention, aes(x = VoteIntention, y = Frequency, fill = VoteIntention)) +
+  geom_bar(stat = "identity") +
+  theme_minimal() +
+  labs(title = "Bar Plot of Vote Intention to Trump", x = "Vote Intention", y = "Frequency")
+
+# Pie chart for vote intention
+pie(result_table_vote_intention$Frequency, labels = paste(result_table_vote_intention$VoteIntention, ": ", round(result_table_vote_intention$Proportion*100, 2), "%", sep=""),
+    main = "Pie Chart of Vote Intention to Trump")
+
+
+### Simulation 2) Data for the primary party ###
+# Set the seed for reproducibility
+set.seed(123)
+
[…]

### 2024-03-13 paper/paper.qmd
updated results section

--- 2024-03-12 paper/paper.qmd
@@
 
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-```
-
 ```{r}
@@
 
+library(rstanarm)
+
+survey_data$consider_trump_binary <- ifelse(survey_data$consider_trump == "Yes", 1, 0)
+survey_model_now <- glm(consider_trump_binary ~ gender + education + age_group + race + state,data = survey_data,family = binomial(link = "logit"))
+
+# get proportions by demographic, then find proportion of each cell relative to population
+post_props <- post_strat_data %>% 
+  group_by(state, race, age_group, gender, education) %>% 
+  summarise(n = n(), .groups ='drop') %>% 
+  group_by(state) %>% 
+  mutate(prop = n/sum(n))
+
+# generate predictions for each of our grouped cells
+post_props$estimate <- predict.glm(survey_model_now, newdata = post_props,
+                          type = "response")
+
+# create a new column with number of voters from each cell voting Trump
+post_props <- post_props %>% mutate(num_voters = n*estimate)
+
+# get errors for our prediction
+errors <- predict.glm(survey_model_now, newdata = post_props,
+                      type = "response", se.fit = T)
+lower <- errors$fit - errors$se.fit
+upper <- errors$fit + errors$se.fit
+
+# combine the proportions and the errors we generated
+post_props_error <- cbind(post_props, lower, upper)
+post_props_error <- post_props_error %>% 
+  rename("lower" = ...10,
+         "upper" = ...11)
+
+# create a new column with number of voters from each cell voting Biden
+post_props_error <- post_props_error %>% mutate(num_voters_lower = n*lower,
+                                num_voters_upper = n*upper)
+
+# create a new dataset which contains upper and lower estimates for proportions
+# of votes for each state
+vote_trump <- post_props_error %>% mutate(trump_prop=estimate*prop,
+                                       trump_prop_lower = lower*prop,
+                                       trump_prop_upper = upper*prop) %>%
+  group_by(state) %>%
+  summarise(trump_predict = sum(trump_prop),
+            trump_prop_lower = sum(trump_prop_lower),
+            trump_prop_upper = sum(trump_prop_upper))
+
 ```
+
+```{r stateplot} 
+#| echo: false 
+#| message: false
+#| label: fig-modelresults
+#| fig-cap: "Proportion of each State Voting for Trump"
+#| warning: false
+
+# get state by state proportions for our polling data
+survey_props <- survey_data %>% 
+  group_by(state, consider_trump_binary) %>% 
+  summarise(n = n()) %>% 
+  mutate(prop = n/sum(n))
+
+# combine the polling data proportions with the post-stratification data
+combined_trump <- inner_join(survey_props, vote_trump, by = "state")
+combined_trump  <- combined_trump %>% 
+  filter(consider_trump_binary == 1)
+
+# arrange our proportions for each state in descending order
+
+stateprops_plot <- combined_trump %>% arrange(desc(trump_predict)) %>% 
+  mutate(support = ifelse(trump_predict >= 0.5, "Trump", "Biden"))
+
+# plot the results, with a line on 50% because of the electoral college
+
+stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
+                               color = support)) +
+  geom_point(stat = "identity") +
+  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
+  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
+  labs(title = "Estimated Proportion of State Voting for Joe Biden",
+       x = "Percentage", y= "State",
+       color = "Support") +
+  scale_x_continuous(labels = scales::percent) +
+  theme(axis.text.y = element_text(size = 5)) +
+  scale_color_manual(values = c("Blue", "Red"))
+```
+
+Figure \@ref(fig:stateplot) shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms.
+Each dot represents the point estimate of the proportion of support for Biden (blue) or Trump (red) in each state.
+Horizontal Lines extending from the dots represent confidence intervals for these estimates.
+The length of each line indicates the uncertainty associated with each estimate.
+For instance, we can see that this uncertainty lies between 50 percent to slightly higher than 80 percent for Trump in MT (Massachusetts).
+The dashed green line in the middle at the 50 percent mark represents the threshold for majority support.
+On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest at the top to the lowest at the bottom.
+
+From figure \@ref(fig:stateplot), it seems that majority of the states support Biden.
+Only 7 states out of 51 have its point estimate greater than 50 percent for Trump.
+However, the horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans.
+However, excluding these contesting states, our model suggests that only 3 states are definitely in favor of Trump whereas
+35 states are definitely supporting Biden.
+
+```{r educationplot} 
+#| echo: false 
+#| message: false 
+#| warning: false 
+#| fig-cap: "How different education levels of voters affect voting for Trump"
+
+# Aggregating data by education level
+post_props_education <- post_strat_data %>% 
+  group_by(education, race, age_group, gender, state) %>%  # Adjust based on variables of interest
+  summarise(n = n(), .groups ='drop') %>%
+  group_by(education) %>%
+  mutate(prop = n/sum(n))
+
+# Generate predictions for each grouped cell based on education
+post_props_education$estimate <- predict.glm(survey_model_now, newdata = post_props_education, type = "response")
+
+# Calculate the number of voters from each cell voting for Trump
+post_props_education <- post_props_education %>% mutate(num_voters = n * estimate)
+
+# Get errors for the prediction
+errors_education <- predict.glm(survey_model_now, newdata = post_props_education,
+                                type = "response", se.fit = TRUE)
+lower_education <- errors_education$fit - errors_education$se.fit
+upper_education <- errors_education$fit + errors_education$se.fit
+
+# Combine the proportions and the errors generated
+post_props_education_error <- cbind(post_props_education, lower_education, upper_education) %>% rename(lower = ...10, upper = ...11)
+
+# Aggregate predictions by education level
+vote_trump_education <- post_props_education_error %>%
+  mutate(trump_prop = estimate * prop,
+         trump_prop_lower = lower * prop,
+         trump_prop_upper = upper * prop) %>%
+  group_by(education) %>%
+  summarise(trump_predict = sum(trump_prop),
+            trump_prop_lower = sum(trump_prop_lower),
+            trump_prop_upper = sum(trump_prop_upper))
+
+vote_trump_education %>% ggplot(aes(y = reorder(education, trump_predict), x = trump_predict)) +
+  geom_point(stat = "identity") +
+  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
+  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
+  labs(title = "Estimated Proportion of Voters for Trump by Education Level",
+       x = "Percentage", y= "Education",
+       color = "Support") +
+  scale_x_continuous(labels = scales::percent) +
+  theme(axis.text.y = element_text(size = 5)) +
+  scale_color_manual(values = c("Blue", "Red"))
+
+```
+
+Figure \@ref(fig:educationplot) presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
+Each black dot represents the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for Trump.
+The horizontal lines extending to the left and right of each dot represent the confidence intervals around the estimate, which reflect the uncertainty.
+
+It shows that regardless of education level, the level of support for Trump lies below 40 percent.
+The Republican party does not have majority, including the error bars across all education levels.
+Voters with "High school or less" education appear to have the lowest estimated support for Trump, which does not align with various exit polls and analyses from the 2020 election suggesting that Trump had substantial support among voters without a college degree.
+Conversely, Voters with 'Some college' and 'Postgrad' education are the two groups that are more in support of Trump, which is exactly the opposite of what we have expected.
 

### 2024-03-13 paper/paper.qmd
updates on results section, removed the model r script as it was causing error somehow.

@@
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-model <- survey_model
-
-# Assuming 'model' is your model object from rstanarm
-library(broom.mixed)
-
-coefficients <- broom.mixed::tidy(model, conf.int = TRUE)
-
-# Assuming you have knitr and kableExtra already installed and loaded
-library(knitr)
-library(kableExtra)
-
-knitr::kable(coefficients,
-             caption = "Coefficients from the Model",
-             digits = 2) %>%
-  kableExtra::kable_styling(latex_options = "scale_down")
-
-```
-
 ## Model set-up
@@
 ```
+
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
@@
 #| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| message: false
 #| warning: false
@@
 
-# get proportions by demographic, then find proportion of each cell relative to population
+# group by each demographic
 post_props <- post_strat_data %>% 
@@
 
-# generate predictions for each of our grouped cells
+# estimate of the model 
 post_props$estimate <- predict.glm(survey_model_now, newdata = post_props,
@@
 
-# create a new column with number of voters from each cell voting Trump
+# new column with the number of voters for Trump
 post_props <- post_props %>% mutate(num_voters = n*estimate)
 
-# get errors for our prediction
+# error terms 
 errors <- predict.glm(survey_model_now, newdata = post_props,
@@
 
-# combine the proportions and the errors we generated
+# combine errors with proportions
 post_props_error <- cbind(post_props, lower, upper)
@@
 
-# create a new column with number of voters from each cell voting Biden
+# new column of voters for trump 
 post_props_error <- post_props_error %>% mutate(num_voters_lower = n*lower,
@@
 
-# create a new dataset which contains upper and lower estimates for proportions
-# of votes for each state
+# confidence intervals for errors 
 vote_trump <- post_props_error %>% mutate(trump_prop=estimate*prop,
@@
 
-```{r stateplot} 
+```{r} 
 #| echo: false
 #| message: false
-#| label: fig-modelresults
-#| fig-cap: "Proportion of each State Voting for Trump"
 #| warning: false
+#| label: fig-distribution-by-state
+#| fig-cap: Distribution of Sample and Population by Education
 
-# get state by state proportions for our polling data
+# state proportions for survey data
 survey_props <- survey_data %>% 
@@
 
-# combine the polling data proportions with the post-stratification data
+# combining survey data with the post-stratification data
 combined_trump <- inner_join(survey_props, vote_trump, by = "state")
@@
 
-# arrange our proportions for each state in descending order
-
 stateprops_plot <- combined_trump %>% arrange(desc(trump_predict)) %>% 
@@
 
-# plot the results, with a line on 50% because of the electoral college
-
+# make a plot of the result for state 
 stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
@@
 
-Figure \@ref(fig:stateplot) shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms.
+Figure \@ref(fig:fig-distribution-by-state) shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms.
 Each dot represents the point estimate of the proportion of support for Biden (blue) or Trump (red) in each state.
@@
 
-From figure \@ref(fig:stateplot), it seems that majority of the states support Biden.
+From figure, it seems that majority of the states support Biden.
 Only 7 states out of 51 have its point estimate greater than 50 percent for Trump.
@@
 
-```{r educationplot} 
+```{r} 
 #| echo: false
@@
 #| warning: false
-#| fig-cap: "How different education levels of voters affect voting for Trump"
+#| label: fig-distribution-by-educationlevel
+#| fig-cap: How different education levels of voters affect voting for Trump
 
@@
 
-Figure \@ref(fig:educationplot) presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
+Figure \@ref(fig:fig-distribution-by-educationlevel) presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
 Each black dot represents the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for Trump.

### 2024-03-14 paper/paper.qmd
minor updates

@@
 library(tidyverse)
-library(readr)
 library(palmerpenguins)
@@
 library(here)
-library(ggplot2)
-library(dplyr)
-library(knitr)
 library(kableExtra)
 library(statebins)
+library(broom.mixed)
+library(knitr)
 
@@
 These samples provide rich qualitative information on the long-term changes in the population.
-We selected the '' data as the post-stratification dataset for our research.
+We selected the '2019 ACS' data (@citeIPUMS) as the post-stratification dataset for our research.
 The ACS is an ongoing survey that collects data monthly, which is then combined into 1-year, 3-year, and 5-year aggregates.
@@
 
-One strength of the IPUMS survey is the fact that it provides a comprehensive and extensive data with detailed demographic of the U.S.
+One strength of the IPUMS survey is the fact that it provides a data with detailed demographic of the U.S.
 population with social, economic and housing characteristics, which is very useful in our analysis of the 2020 U.S presidential election forecast.
@@
 
-The variables we have decided to use in this analysis from the dataset are 'sex', 'race', 'stateicp', 'age_group', and 'educd'.
-We have filtered out the respondents who have answered 'other' or no available sex data from the raw dataset for simplicity of the analysis.
-Therefore, 'Sex' is categorized into 'Male' and 'Female'.
-'Race' is divided into 'White', 'Black', 'Asian', 'American Indian' and 'Other'.
-As mentioned earlier, the fact that white, black and Asian together make up around 93 percent of the U.S population, this categorization is justified.
-'stateicp' is a record of all the states in the U.S.
-We have used the abbreviations for each state, for example, 'CT' for Connecticut, 'ME' for Maine, 'MA' for Massachusetts.
-We have the data of 55 states in total.
-'age_group' categorized the respondents' age into four groups: '18-29', '30-49', '50-64' and '65+'.
-'education' is divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
-We filtered out those unknowns, and cleaned the dataset by assigning each respondent to the right category accordingly.
+In processing the raw post-stratification dataset, which initially contained approximately 3.2 million records, we refined it down to about 2.3 million records.
+This was achieved through a meticulous selection process, ensuring the data's integrity and relevance for our analysis.
+In our analysis, we've selected the variables 'sex', 'race', 'stateicp', 'age', and 'educd' from the dataset.
+To simplify our analysis, respondents who indicated 'other' or provided no data for their sex have been excluded.
+Consequently, 'sex' has been categorized strictly as 'Male' and 'Female'.
+We've refined 'race' into five categories: 'White', 'Black', 'Asian', 'American Indian', and 'Other', based on the composition of the U.S.
+population, with White, Black, and Asian categories accounting for approximately 93 percent of the total.
+
+The 'stateicp' variable encompasses all U.S.
+states, using their standard abbreviations (e.g., 'CT' for Connecticut), and extends to 55 values to include 'Puerto Rico', 'State groupings (1980 Urban/rural sample)', 'Military/Military Reservations', 'District of Columbia', and an 'State not identified' category.
+
+Age has been grouped into four categories: '18-29', '30-49', '50-64', and '70+'.
+For educational attainment ('educd'), we've created four categories: 'High school or less', 'Some college', 'College degree', and 'Postgrad'.
+
+We excluded any unknown responses to ensure clarity and accuracy in categorization and to enhance clarity and align with survey data, we've renamed 'sex', 'stateicp', 'age', and 'educd' to 'gender', 'state', 'age_group', and 'education', respectively.
+This restructuring aims to streamline our analysis by ensuring each respondent is accurately categorized.
+
+@fig-distribution-by-age-group, @fig-distribution-by-gender, @fig-distribution-by-race, and @fig-distribution-by-education illustrate comparisons between survey data and post-stratification data across different variables.
+In these visuals, orange bars represent post-stratification data, while green bars signify survey data.
+The percentages displayed on each figure are rounded to the nearest tenth, introducing a potential margin of error of ±0.1% in the total values.
+Generally, the survey data aligns closely with the post-stratification data, maintaining a discrepancy of about 10% across most categories.
+Notable exceptions are observed in the '30-49' age group and the 'Some college' education level, where the differences exceed this margin.
+
+```{r}
+# insert figure for state here
+```
 
@@
 
+@tbl-voters-intention-to-vote-for-trump shows the proportion of voters who intend to vote for Donald Trump or not.
+Also, @tbl-voters-intention-of-their-primary-party shows the proportion of voters' supporting party.
+The data used to create these table is from the @citeSurveyData.
+We see that Donald Trump and his party Republican are not expected to win the popular vote before we implement the model.
+
 ```{r}
@@
 #| tbl-cap: Voters Intention of Their Primary Party
+
 data <- cleaned_data_survey
@@
 # Model
-
 For our study, we employ a technique called multilevel regression with post-stratification (MRP).
@@
 
-## Model set-up
+The key steps in MRP involve initially selecting a dataset for model development.
+In this case, we utilized survey data from the Voter Study Group (@citeSurveyData).
+The next step is to construct a model with this smaller dataset; here, we employed logistic regression based on the survey data, formulated as seen in equation \ref{eq:logit}.
+Following model creation, it is then applied to a broader dataset to estimate population characteristics.
+For our analysis, Census data from IPUMS (@citeIPUMS) served as this larger dataset.
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+To predict an individual's likelihood of voting for Donald Trump, we aim to construct a logistic regression model leveraging data from the Voter Study Group (@citeSurveyData) and applying post-stratification with Census Data (@citeIPUMS).
+Given that logistic regression is suited for binary outcomes, we've introduced a variable, 'consider_trump', which assigns a 1 if the respondent indicates a plan to vote for Donald Trump, and a 0 for intentions to vote for other candidates, with 0 encompassing both "No" and "Other" responses.
 
-```{=tex}
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-```
+The logistic regression model takes the form of:
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+\begin{equation}
+log(\frac{\hat{p}}{1 - \hat{p}}) = \beta_{0} + \beta_{1}x_{sex} +
+\beta_{2}x_{agegroup} + \beta_{3}x_{race} + \beta_{4}x_{state} + \beta_{5}x_{education}
+\label{eq:logit}
+\end{equation}
 
-# Results
+After developing our logistic regression model, we'll utilize the `predict` function in R (@citeR) to apply it to the Census data (@citeIPUMS).
+This involves segmenting the Census dataset by our target demographic categories—sex, race, age_group, education level, and state—and then executing the model for each group.
+This process yields probabilities that an individual within a specific group is likely to vote for Donald Trump.
+With these predictions in hand, we can assess potential outcomes, such as the winner of the popular vote or the distribution of electoral college votes.
+Additionally, we'll employ a 95 percent confidence interval to estimate the popular vote percentage, indicating a 95 percent certainty level that the true population value falls within our calculated range.
+This method suggests our findings will have an accuracy margin of +/- 4%.
+
+In the equation \ref{eq:logit}, each $\beta$ is identified as a coefficient to be determined by the regression analysis.
+The chosen variables for this study include sex, age, race, education, and state, selected primarily because the first three have been recognized as consistent indicators of an individual's preferred candidate.
+This rationale stems from observed patterns, such as certain states consistently voting for the Republican party, whereas others oscillate between Democratic and Republican preferences in successive elections.
+Furthermore, we opted for education level over income as a variable, believing that education offers a more definitive characterization of an individual, unlike income.
 
@@
 #| warning: false
+#| label: tbl-model-coefficients
+#| tbl-cap: Coefficients from the Model
+
+model <- survey_model
+
+coefficients <- broom.mixed::tidy(model, conf.int = TRUE)
+
+kable(coefficients, format = "latex", booktabs = TRUE, align = "c") %>%
+  kable_styling(latex_options = "scale_down", font_size = 7)
+
+```
+
+We are running our regression model using the stan_glm() function in R (@citeR).
+The decision to run this model over other models like linear regression was made by the fact that we were predicting a binary variable about a voter's decision.
+Since there are only two possible options our data will likely follow an S shape and a straight line equation will not be helpful to model this relationship.
+Another strength present for logistic regression is that when combined with post-stratification it allows us to take information from under-represented populations and it allows their views to be accounted for more greatly.
+For example, our survey data (@citeSurveyData), includes only 7 observations from Alaska, but using multilevel regression with post-stratification, we can have that expanded to over 4500 people.
+
+Our model does have some weaknesses, since the output must be binary, we cannot account for other candidates or a person deciding not to vote.
+This issue isn't too large because our main goal is to determine which of the two main candidates will be chosen by the people of America.
+Another weakness we do encounter with our model and multi-level regression with post-stratification is that it has a strong dependence on the survey data.
+This is a weakness because if the survey has any gaps or there are any tweaks we need to make, it can change the course of results.
+
+@tbl-model-coefficients shows the estimates for the coefficients that will fit into our logistic regression equation.
+These coefficients will fit into Equation @logit, and were calculated using data from the Voter Study Group (@citeSurveyData).
+The table is made using `kable` from `knitr` (@citeKnitr) and is formatted using `kableExtra` (@citekableExtra)^[data was cleaned using the `tidy` function called `broom` (@citebroom)].
+We can see that our p-values for variables like gender, age, hispanic and education return very significant results while the states have varying levels of significance.
+What our p-values tell us is whether or not our variable is statistically significant in impacting our response variable, in this case, whether the respondent will vote for Biden or not.
+When we look at our p-values, we want them to be as small as possible and can assume they're significant if it's under 0.05.
+As for the coefficients, the value we receive is in terms of log odds.
+When the log odds are positive, it means that the person will likely vote for Biden and if it's negative it means they'll likely vote for Trump.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-coefficient-estimates
+#| fig-cap: Coefficient Estimates
+
+model <- survey_model
+
+coefficients <- broom.mixed::tidy(model, conf.int = TRUE)
+# plot our coefficients with error bars to show the upper and lower
+# estimates for the coefficients
+coefficients %>% ggplot(aes(estimate, term)) +
+  geom_point() +
+  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
+  labs(title = "Coefficient Estimates",
+       x = "Estimate", y = "Coefficient") +
+  theme(axis.text.y = element_text(size = 5))
+```
+
+@fig-coefficient-estimates shows us the coefficients that would fit into equation \ref{eq:logit} using the polling data (@citeSurveyData).
+We also have error bars present, which show the upper and lower estimates for the coefficients.
+What we have to look out for in this scenario is that coefficients with negative values would mean that the person is more likely to vote for Donald Trump (with that characteristic) and positive values mean the person is more likely to vote for Joe Biden.
+@tbl-model-coefficients shows a numerical view of @fig-coefficient-estimates, along with p-values.
+
+Using the outputs of the logistic regression model, we can get an equation that follows the form of \ref{eq:logit}, but with the $\beta$ values filled out.
+This equation is difficult to write out because of the many variables, but in short, if the person's characteristic fits in with a certain variable, it is used in the equation.
+Then the equation is summed up and the probability is found using equation \@ref{eq:prob}.
+
+We can also look at some common assumptions for logistic regression and check if our model follows them or not.
+One assumption is that the data set used is large.
+The data we used to create the model had about 5000 observations, which is high but maybe in the future, we could use a larger data set to increase the accuracy of predictions.
+Another assumption that logistic regression makes is that the observations are independent of each other.
+We know that the observations used for our model were independent, not only because the Voter Study Group ensures that there aren't duplicates but because they assign each respondent a unique ID number.
+
+Lastly, using our coefficients (see @tbl-model-coefficients), we can calculate the upper and lower bounds for the probabilities predicted from our model.
+For our lower probability (white male, from North Dakota, aged 36-49, with some post-secondary education and not hispanic) we get a probability of 22% for supporting Biden.
+For our upper probability (black woman, from Vermont, aged 18-35, with post-secondary or higher education and considers themselves hispanic) we get a probability of 99% for supporting Joe Biden.
+
+# Results
+
+Our results are summarized in @tbl-modelresults.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 
 library(rstanarm)
-
-survey_data$consider_trump_binary <- ifelse(survey_data$consider_trump == "Yes", 1, 0)
-survey_model_now <- glm(consider_trump_binary ~ gender + education + age_group + race + state,data = survey_data,family = binomial(link = "logit"))
-
-# group by each demographic
-post_props <- post_strat_data %>% 
-  group_by(state, race, age_group, gender, education) %>% 
-  summarise(n = n(), .groups ='drop') %>% 
-  group_by(state) %>% 
-  mutate(prop = n/sum(n))
-
-# estimate of the model 
-post_props$estimate <- predict.glm(survey_model_now, newdata = post_props,
-                          type = "response")
-
-# new column with the number of voters for Trump
-post_props <- post_props %>% mutate(num_voters = n*estimate)
-
-# error terms 
-errors <- predict.glm(survey_model_now, newdata = post_props,
-                      type = "response", se.fit = T)
-lower <- errors$fit - errors$se.fit
-upper <- errors$fit + errors$se.fit
-
-# combine errors with proportions
-post_props_error <- cbind(post_props, lower, upper)
-post_props_error <- post_props_error %>% 
-  rename("lower" = ...10,
-         "upper" = ...11)
-
-# new column of voters for trump 
-post_props_error <- post_props_error %>% mutate(num_voters_lower = n*lower,
-                                num_voters_upper = n*upper)
-
-# confidence intervals for errors 
-vote_trump <- post_props_error %>% mutate(trump_prop=estimate*prop,
-                                       trump_prop_lower = lower*prop,
-                                       trump_prop_upper = upper*prop) %>%
-  group_by(state) %>%
-  summarise(trump_predict = sum(trump_prop),
-            trump_prop_lower = sum(trump_prop_lower),
-            trump_prop_upper = sum(trump_prop_upper))
 
@@
 #| echo: false
-#| message: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
-#| label: fig-distribution-by-state
-#| fig-cap: Distribution of Sample and Population by Education
-
-# state proportions for survey data
-survey_props <- survey_data %>% 
-  group_by(state, consider_trump_binary) %>% 
-  summarise(n = n()) %>% 
-  mutate(prop = n/sum(n))
-
-# combining survey data with the post-stratification data
-combined_trump <- inner_join(survey_props, vote_trump, by = "state")
-combined_trump  <- combined_trump %>% 
-  filter(consider_trump_binary == 1)
-
-stateprops_plot <- combined_trump %>% arrange(desc(trump_predict)) %>% 
-  mutate(support = ifelse(trump_predict >= 0.5, "Trump", "Biden"))
-
-# make a plot of the result for state 
-stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
-                               color = support)) +
-  geom_point(stat = "identity") +
-  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
-  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
-  labs(title = "Estimated Proportion of State Voting for Joe Biden",
-       x = "Percentage", y= "State",
-       color = "Support") +
-  scale_x_continuous(labels = scales::percent) +
-  theme(axis.text.y = element_text(size = 5)) +
-  scale_color_manual(values = c("Blue", "Red"))
-```
-
-Figure \@ref(fig:fig-distribution-by-state) shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms.
-Each dot represents the point estimate of the proportion of support for Biden (blue) or Trump (red) in each state.
-Horizontal Lines extending from the dots represent confidence intervals for these estimates.
-The length of each line indicates the uncertainty associated with each estimate.
-For instance, we can see that this uncertainty lies between 50 percent to slightly higher than 80 percent for Trump in MT (Massachusetts).
-The dashed green line in the middle at the 50 percent mark represents the threshold for majority support.
-On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest at the top to the lowest at the bottom.
-
-From figure, it seems that majority of the states support Biden.
-Only 7 states out of 51 have its point estimate greater than 50 percent for Trump.
-However, the horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans.
-However, excluding these contesting states, our model suggests that only 3 states are definitely in favor of Trump whereas
-35 states are definitely supporting Biden.
-
-```{r} 
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-distribution-by-educationlevel
-#| fig-cap: How different education levels of voters affect voting for Trump
-
-# Aggregating data by education level
-post_props_education <- post_strat_data %>% 
-  group_by(education, race, age_group, gender, state) %>%  # Adjust based on variables of interest
-  summarise(n = n(), .groups ='drop') %>%
-  group_by(education) %>%
-  mutate(prop = n/sum(n))
-
-# Generate predictions for each grouped cell based on education
-post_props_education$estimate <- predict.glm(survey_model_now, newdata = post_props_education, type = "response")
-
-# Calculate the number of voters from each cell voting for Trump
-post_props_education <- post_props_education %>% mutate(num_voters = n * estimate)
-
-# Get errors for the prediction
-errors_education <- predict.glm(survey_model_now, newdata = post_props_education,
-                                type = "response", se.fit = TRUE)
-lower_education <- errors_education$fit - errors_education$se.fit
-upper_education <- errors_education$fit + errors_education$se.fit
-
-# Combine the proportions and the errors generated
-post_props_education_error <- cbind(post_props_education, lower_education, upper_education) %>% rename(lower = ...10, upper = ...11)
-
-# Aggregate predictions by education level
-vote_trump_education <- post_props_education_error %>%
-  mutate(trump_prop = estimate * prop,
-         trump_prop_lower = lower * prop,
-         trump_prop_upper = upper * prop) %>%
-  group_by(education) %>%
-  summarise(trump_predict = sum(trump_prop),
-            trump_prop_lower = sum(trump_prop_lower),
-            trump_prop_upper = sum(trump_prop_upper))
-
-vote_trump_education %>% ggplot(aes(y = reorder(education, trump_predict), x = trump_predict)) +
-  geom_point(stat = "identity") +
-  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
-  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
-  labs(title = "Estimated Proportion of Voters for Trump by Education Level",
-       x = "Percentage", y= "Education",
-       color = "Support") +
-  scale_x_continuous(labels = scales::percent) +
-  theme(axis.text.y = element_text(size = 5)) +
-  scale_color_manual(values = c("Blue", "Red"))
 
 ```
-
-Figure \@ref(fig:fig-distribution-by-educationlevel) presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
-Each black dot represents the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for Trump.
-The horizontal lines extending to the left and right of each dot represent the confidence intervals around the estimate, which reflect the uncertainty.
-
-It shows that regardless of education level, the level of support for Trump lies below 40 percent.
-The Republican party does not have majority, including the error bars across all education levels.
-Voters with "High school or less" education appear to have the lowest estimated support for Trump, which does not align with various exit polls and analyses from the 2020 election suggesting that Trump had substantial support among voters without a college degree.
-Conversely, Voters with 'Some college' and 'Postgrad' education are the two groups that are more in support of Trump, which is exactly the opposite of what we have expected.
 

### 2024-03-14 scripts/post-strat/01-data-cleaning-poststratification.R

@@ -53,7 +53,7 @@ cleaned_data <- raw_data %>%
            race == 2 ~ "Black",
            race == 3 ~ "American Indian",
            race %in% c(4, 5, 6) ~ "Asian",
-           race %in% c(3, 7, 8, 9) ~ "Other",
+           race %in% c(7, 8, 9) ~ "Other",
            TRUE ~ "Unknown"
          ),
          education = case_when(
@@ -63,7 +63,7 @@ cleaned_data <- raw_data %>%
            educd >= 114 ~ "Postgrad",
            TRUE ~ "Unknown"
          )) %>%
-  rename(state = stateicp, gender = sex, ) %>%
+  rename(state = stateicp, gender = sex) %>%
   select(-age, -educd) %>% # Remove the original 'age' and 'educd' columns to keep only the age group and education group
   # Filter step to remove rows with "Unknown" in any of the transformed columns
   filter(!if_any(c(gender, age_group, race, education), ~ .x == "Unknown"))

### 2024-03-14 paper/paper.qmd

--- 2024-03-13 paper/paper.qmd
@@
 
-## Model set-up
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+model <- survey_model
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+# Assuming 'model' is your model object from rstanarm
+library(broom.mixed)
 
-```{=tex}
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+coefficients <- broom.mixed::tidy(model, conf.int = TRUE)
+
+# Assuming you have knitr and kableExtra already installed and loaded
+library(knitr)
+library(kableExtra)
+
+knitr::kable(coefficients,
+             caption = "Coefficients from the Model",
+             digits = 2) %>%
+  kableExtra::kable_styling(latex_options = "scale_down")
+
 ```
 
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-coefficient-estimates
+#| fig-cap: Coefficient Estimates
+
+model <- survey_model
+
+coefficients <- broom.mixed::tidy(model, conf.int = TRUE)
+# plot our coefficients with error bars to show the upper and lower
+# estimates for the coefficients
+coefficients %>% ggplot(aes(estimate, term)) +
+  geom_point() +
+  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
+  labs(title = "Coefficient Estimates",
+       x = "Estimate", y = "Coefficient") +
+  theme(axis.text.y = element_text(size = 5))
+```
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### 2024-03-14 paper/paper.qmd
major updates on paper.qmd model section

@@
 survey_model <- readRDS(file = here("models/consider_trump.rds"))
-predicton_model <- readRDS(file = here("models/census_data_with_predictions.rds"))
+prediction_model <- readRDS(file = here("models/census_data_with_predictions.rds"))
 ```
@@
 states, using their standard abbreviations (e.g., 'CT' for Connecticut), and extends to 55 values to include 'Puerto Rico', 'State groupings (1980 Urban/rural sample)', 'Military/Military Reservations', 'District of Columbia', and an 'State not identified' category.
-
 Age has been grouped into four categories: '18-29', '30-49', '50-64', and '70+'.
@@
 
+The key steps in MRP involve initially selecting a dataset for model development.
+In this case, we utilized survey data from the Voter Study Group (@citeSurveyData).
+The next step is to construct a model with this smaller dataset; here, we employed logistic regression based on the survey data, formulated as seen in equation \ref{eq:logit}.
+Following model creation, it is then applied to a broader dataset to estimate population characteristics.
+For our analysis, Census data from IPUMS (@citeIPUMS) served as this larger dataset.
+
+To predict an individual's likelihood of voting for Donald Trump, we aim to construct a logistic regression model leveraging data from the Voter Study Group (@citeSurveyData) and applying post-stratification with Census Data (@citeIPUMS).
+Given that logistic regression is suited for binary outcomes, we've introduced a variable, 'consider_trump', which assigns a 1 if the respondent indicates a plan to vote for Donald Trump, and a 0 for intentions to vote for other candidates, with 0 encompassing both "No" and "Other" responses.
+
+The logistic regression model takes the form of:
+
+\begin{equation}
+log(\frac{\hat{p}}{1 - \hat{p}}) = \beta_{0} + \beta_{1}x_{sex} +
+\beta_{2}x_{agegroup} + \beta_{3}x_{race} + \beta_{4}x_{state} + \beta_{5}x_{education}
+\label{eq:logit}
+\end{equation}
+
+In equation \ref{eq:logit}, each $\beta$ represents a coefficient determined through regression analysis.
+The variables chosen for this project are sex, age, race, education, and state.
+These were selected because sex, age, and race are proven to be reliable indicators of voting preferences.
+This decision is based on patterns such as certain states consistently favoring the Republican party, while others alternate between Democratic and Republican.
+Education was chosen over income because it provides a clearer picture of an individual’s background than income does.
+
+Once the logistic regression model is developed, we'll use the `predict()` function in R (@citeR) to apply our model to Census data (@citeIPUMS), breaking down the dataset into categories based on sex, race, age group, education level, and state.
+This will give us the likelihood of individuals within each category voting for Donald Trump.
+These predictions allow us to analyze potential outcomes like the popular vote winner or the electoral college vote distribution.
+
 ```{r}
@@
 #| warning: false
+#| label: tbl-model-coefficients
+#| tbl-cap: Coefficients from the Model
 model <- survey_model 
-
-# Assuming 'model' is your model object from rstanarm
-library(broom.mixed)
-
 coefficients <- broom.mixed::tidy(model, conf.int = TRUE)
 
-# Assuming you have knitr and kableExtra already installed and loaded
-library(knitr)
-library(kableExtra)
+kable(coefficients, format = "latex", booktabs = TRUE, align = "c") %>%   kable_styling(latex_options = "scale_down", font_size = 7)
+```
 
-knitr::kable(coefficients,
-             caption = "Coefficients from the Model",
-             digits = 2) %>%
-  kableExtra::kable_styling(latex_options = "scale_down")
+@tbl-model-coefficients shows the estimates for the coefficients that will fit into our logistic regression equation.
+These coefficients will fit into Equation \ref{eq:logit}, and were calculated using data from the Voter Study Group (@citeSurveyData).
+The table is made using `kable` from `knitr` (@citeKnitr) and is formatted using `kableExtra` (@citekableExtra).
 
-```
+We use the `stan_glm()` function in R (@citeR) for our regression analysis, specifically because we are dealing with a binary outcome: whether a voter supports Donald Trump or not.
+The nature of our data suggests an S-shaped distribution rather than a linear one, making logistic regression a better fit than linear regression.
+This approach is advantageous, especially when paired with post-stratification, as it allows us to better represent under-represented groups in our analysis.
+For instance, despite having only 7 responses from Alaska in our survey data (@citeSurveyData), through multilevel regression and post-stratification, we can adjust this to effectively represent over 4500 individuals.
+
+However, there are limitations to our model.
+The binary outcome does not allow for consideration of third-party candidates or non-voters, although this limitation is mitigated by our focus on the main candidates.
+More critically, our model's accuracy is heavily dependent on the quality of our survey data.
+Any inaccuracies or the need for adjustments in the survey can significantly impact our findings.
 
@@
 
-model <- survey_model
+coefficients %>%
+  ggplot(aes(estimate, term)) + geom_point() + geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) + labs(title = "Coefficient Estimates", x = "Estimate", y = "Coefficient") +   theme(axis.text.y = element_text(size = 5))
+```
 
-coefficients <- broom.mixed::tidy(model, conf.int = TRUE)
-# plot our coefficients with error bars to show the upper and lower
-# estimates for the coefficients
-coefficients %>% ggplot(aes(estimate, term)) +
-  geom_point() +
-  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
-  labs(title = "Coefficient Estimates",
-       x = "Estimate", y = "Coefficient") +
-  theme(axis.text.y = element_text(size = 5))
+@fig-coefficient-estimates presents the coefficients derived from logistic regression on the survey data (@citeSurveyData).
+It also includes error bars, indicating the confidence interval for each coefficient estimate.
+In interpreting these coefficients, it's essential to understand that positive values suggest a greater likelihood of voting for Donald Trump, whereas negative values indicate a tendency to vote for other candidates, such as Joe Biden.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-prediction-probabilities
+#| tbl-cap: Example of Prediction model
+
+# Select only the first 10 rows for the table
+probability_subset <- head(prediction_model, 10)
+
+# Create the table with kable and format with kable_styling
+kable(probability_subset, format = "latex", booktabs = TRUE, align = "c") %>%
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
 ```
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+By utilizing the results from our logistic regression model, we can formulate an equation that adheres to the structure outlined in equation \ref{eq:logit}, incorporating specific $\beta$ coefficients for each variable.
+Given the number of variables, detailing the equation fully is challenging.
+Essentially, the equation integrates the $\beta$ value of a variable if an individual's characteristic matches that variable.
+@tbl-prediction-probabilities offers examples of how the probability varies based on different variables.
 

### 2024-03-14 paper/paper.qmd
abstract draft

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "In this study, we forecasted the 2020 U.S.
+Presidential Election outcomes using a sophisticated statistical technique called multilevel regression with post-stratification (MRP), applied to a large-scale national survey dataset and census data.
+Our findings reveal how demographic factors such as age, gender, education, race, and geographic location influence voting behavior, with particular emphasis on voter preferences for Donald Trump.
+The analysis highlights significant variations in voter support across different states and educational backgrounds, underscoring the complex interplay of demographics in shaping electoral outcomes.
+This research contributes to our understanding of the American electoral landscape, demonstrating the critical role of demographic diversity in determining the direction of political preferences and election results."
 format: pdf

### 2024-03-14 paper/paper.qmd
working title

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "2020 U.S. Presidential Election Forecast"
+subtitle: "How Multilevel Regression and Post-Stratification Unveil Voter Tendencies"
 author: 

### 2024-03-14 paper/paper.qmd
update github repo link for thanks to

--- 2024-03-14 paper/paper.qmd
@@
   - Jiwon Choi
-thanks: "Code and data are available at: https://github.com/Kjeongwoo99/STA302H_Paper3"
+thanks: "Code and data are available at: https://github.com/Kjeongwoo99/2020_US_Election_Forecast"
 date: today

### 2024-03-14 paper/paper.qmd
discussion update

--- 2024-03-14 paper/paper.qmd
@@
 date-format: long
-abstract: "In this study, we forecasted the 2020 U.S.
-Presidential Election outcomes using a sophisticated statistical technique called multilevel regression with post-stratification (MRP), applied to a large-scale national survey dataset and census data.
-Our findings reveal how demographic factors such as age, gender, education, race, and geographic location influence voting behavior, with particular emphasis on voter preferences for Donald Trump.
-The analysis highlights significant variations in voter support across different states and educational backgrounds, underscoring the complex interplay of demographics in shaping electoral outcomes.
-This research contributes to our understanding of the American electoral landscape, demonstrating the critical role of demographic diversity in determining the direction of political preferences and election results."
+abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
 format: pdf
@@
 library(tidyverse)
+library(readr)
 library(palmerpenguins)
@@
 library(here)
+library(ggplot2)
+library(dplyr)
+library(knitr)
 library(kableExtra)
 library(statebins)
-library(broom.mixed)
-library(knitr)
 
@@
 cleaned_data_survey <- read_csv(file = here("data/analysis_data/ns20191003_cleaned.csv"), show_col_types = FALSE)
+
 survey_model <- readRDS(file = here("models/consider_trump.rds"))
-prediction_model <- readRDS(file = here("models/census_data_with_predictions.rds"))
+predicton_model <- readRDS(file = here("models/census_data_with_predictions.rds"))
 ```
@@
 These samples provide rich qualitative information on the long-term changes in the population.
-We selected the '2019 ACS' data (@citeIPUMS) as the post-stratification dataset for our research.
+We selected the '' data as the post-stratification dataset for our research.
 The ACS is an ongoing survey that collects data monthly, which is then combined into 1-year, 3-year, and 5-year aggregates.
@@
 
-One strength of the IPUMS survey is the fact that it provides a data with detailed demographic of the U.S.
+One strength of the IPUMS survey is the fact that it provides a comprehensive and extensive data with detailed demographic of the U.S.
 population with social, economic and housing characteristics, which is very useful in our analysis of the 2020 U.S presidential election forecast.
@@
 
-In processing the raw post-stratification dataset, which initially contained approximately 3.2 million records, we refined it down to about 2.3 million records.
-This was achieved through a meticulous selection process, ensuring the data's integrity and relevance for our analysis.
-In our analysis, we've selected the variables 'sex', 'race', 'stateicp', 'age', and 'educd' from the dataset.
-To simplify our analysis, respondents who indicated 'other' or provided no data for their sex have been excluded.
-Consequently, 'sex' has been categorized strictly as 'Male' and 'Female'.
-We've refined 'race' into five categories: 'White', 'Black', 'Asian', 'American Indian', and 'Other', based on the composition of the U.S.
-population, with White, Black, and Asian categories accounting for approximately 93 percent of the total.
-
-The 'stateicp' variable encompasses all U.S.
-states, using their standard abbreviations (e.g., 'CT' for Connecticut), and extends to 55 values to include 'Puerto Rico', 'State groupings (1980 Urban/rural sample)', 'Military/Military Reservations', 'District of Columbia', and an 'State not identified' category.
-Age has been grouped into four categories: '18-29', '30-49', '50-64', and '70+'.
-For educational attainment ('educd'), we've created four categories: 'High school or less', 'Some college', 'College degree', and 'Postgrad'.
-
-We excluded any unknown responses to ensure clarity and accuracy in categorization and to enhance clarity and align with survey data, we've renamed 'sex', 'stateicp', 'age', and 'educd' to 'gender', 'state', 'age_group', and 'education', respectively.
-This restructuring aims to streamline our analysis by ensuring each respondent is accurately categorized.
-
-@fig-distribution-by-age-group, @fig-distribution-by-gender, @fig-distribution-by-race, and @fig-distribution-by-education illustrate comparisons between survey data and post-stratification data across different variables.
-In these visuals, orange bars represent post-stratification data, while green bars signify survey data.
-The percentages displayed on each figure are rounded to the nearest tenth, introducing a potential margin of error of ±0.1% in the total values.
-Generally, the survey data aligns closely with the post-stratification data, maintaining a discrepancy of about 10% across most categories.
-Notable exceptions are observed in the '30-49' age group and the 'Some college' education level, where the differences exceed this margin.
-
-```{r}
-# insert figure for state here
-```
+The variables we have decided to use in this analysis from the dataset are 'sex', 'race', 'stateicp', 'age_group', and 'educd'.
+We have filtered out the respondents who have answered 'other' or no available sex data from the raw dataset for simplicity of the analysis.
+Therefore, 'Sex' is categorized into 'Male' and 'Female'.
+'Race' is divided into 'White', 'Black', 'Asian', 'American Indian' and 'Other'.
+As mentioned earlier, the fact that white, black and Asian together make up around 93 percent of the U.S population, this categorization is justified.
+'stateicp' is a record of all the states in the U.S.
+We have used the abbreviations for each state, for example, 'CT' for Connecticut, 'ME' for Maine, 'MA' for Massachusetts.
+We have the data of 55 states in total.
+'age_group' categorized the respondents' age into four groups: '18-29', '30-49', '50-64' and '65+'.
+'education' is divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
+We filtered out those unknowns, and cleaned the dataset by assigning each respondent to the right category accordingly.
 
@@
 
-```{r fig.width=10, fig.height=5}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-distribution-by-gender
-#| fig-cap: Distribution of Sample and Population by Gender
+# Model
 
-# Read in the survey data
-survey_data <- cleaned_data_survey
+The goal of our modelling strategy is twofold. Firstly,...
 
-# Read in the post-stratification data
-post_strat_data <- cleaned_data_post_strat
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
 
-# Calculate percentages for the survey data by gender
-survey_gender <- survey_data %>%
-  group_by(gender) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+## Model set-up
 
-# Calculate percentages for the post-strat data by gender
-post_strat_gender <- post_strat_data %>%
-  group_by(gender) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
 
-# Combine the gender percentages into a single data frame for plotting
-combined_gender_percentages <- rbind(
-  data.frame(dataset = 'Survey', gender = survey_gender$gender, percentage = survey_gender$percentage),
-  data.frame(dataset = 'Post-Strat', gender = post_strat_gender$gender, percentage = post_strat_gender$percentage)
-)
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_i + \gamma_i\\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta &\sim \mbox{Normal}(0, 2.5) \\
+\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
+\end{align}
 
-# Plot the gender percentages with specified bar width and percentage on y-axis, and label each bar
-ggplot(combined_gender_percentages, aes(x = gender, y = percentage, fill = dataset)) +
-  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
-  geom_text(aes(label = sprintf("%.1f%%", percentage)),
-            position = position_dodge(width = 0.8), 
-            vjust = -0.3, # Adjust text position
-            color = "black",
-            size = 3) + 
-  labs(x = "Gender", y = "Proportion", title = "Distribution of Sample and Population by Gender") +
-  theme_minimal() +
-  scale_fill_manual(values = c("orange", "green"))
-```
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
 
-```{r fig.width=10, fig.height=5}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-distribution-by-race
-#| fig-cap: Distribution of Sample and Population by Race Ethnicity
-# Read in the survey data
-survey_data <- cleaned_data_survey
-# Read in the post-stratification data
-post_strat_data <- cleaned_data_post_strat
-# Calculate percentages for the survey data by race
-survey_race <- survey_data %>%
-  group_by(race) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
-# Calculate percentages for the post-strat data by race
-post_strat_race <- post_strat_data %>%
-  group_by(race) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
-# Combine the race percentages into a single data frame for plotting
-combined_race_percentages <- rbind(
-  data.frame(dataset = 'Survey', race = survey_race$race, percentage = survey_race$percentage),
-  data.frame(dataset = 'Post-Strat', race = post_strat_race$race, percentage = post_strat_race$percentage)
-)
-# Adjust the factor levels for the 'race' column to specify the order
-combined_race_percentages$race <- factor(combined_race_percentages$race, 
-                                                    levels = c("White", "Black", "Asian", "American Indian", "Other"))
-# Plot the race percentages with specified bar width and percentage on y-axis, and label each bar
-ggplot(combined_race_percentages, aes(x = race, y = percentage, fill = dataset)) +
-  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
-  geom_text(aes(label = sprintf("%.1f%%", percentage)),
-            position = position_dodge(width = 0.8), 
-            vjust = -0.3, # Adjust text position
-            color = "black",
-            size = 3) + 
-  labs(x = "Race", y = "Proportion", title = "Distribution of Sample and Population by Race Ethnicity") +
-  theme_minimal() +
-  scale_fill_manual(values = c("orange", "green"))
-```
+# Results
 
-```{r fig.width=10, fig.height=5}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-distribution-by-education
-#| fig-cap: Distribution of Sample and Population by Education
-# Read in the survey data
-survey_data <- cleaned_data_survey
-# Read in the post-stratification data
-post_strat_data <- cleaned_data_post_strat
-# Calculate percentages for the survey data by education
-survey_education <- survey_data %>%
-  group_by(education) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
-# Calculate percentages for the post-strat data by education
-post_strat_education <- post_strat_data %>%
-  group_by(education) %>%
-  summarise(count = n()) %>%
-  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
-# Combine the education percentages into a single data frame for plotting
-combined_education_percentages <- rbind(
-  data.frame(dataset = 'Survey', education = survey_education$education, percentage = survey_education$percentage),
-  data.frame(dataset = 'Post-Strat', education = post_strat_education$education, percentage = post_strat_education$percentage)
-)
-# Adjust the factor levels for the 'education' column to specify the order
-combined_education_percentages$education <- factor(combined_education_percentages$education, 
-                                                    levels = c("High school or less", "Some college", "College degree", "Postgrad"))
-# Plot the education percentages with the adjusted order
-ggplot(combined_education_percentages, aes(x = education, y = percentage, fill = dataset)) +
-  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
-  geom_text(aes(label = sprintf("%.1f%%", percentage)),
-            position = position_dodge(width = 0.8), 
-            vjust = -0.3, # Adjust text position
-            color = "black",
-            size = 3) + 
-  labs(x = "Education", y = "Proportion", title = "Distribution of Sample and Population by Education") +
-  theme_minimal() +
-  scale_fill_manual(values = c("orange", "green"))
-```
-
-@tbl-voters-intention-to-vote-for-trump shows the proportion of voters who intend to vote for Donald Trump or not.
-Also, @tbl-voters-intention-of-their-primary-party shows the proportion of voters' supporting party.
-The data used to create these table is from the @citeSurveyData.
-We see that Donald Trump and his party Republican are not expected to win the popular vote before we implement the model.
+Our results are summarized in @tbl-modelresults.
 
@@
 #| echo: false
+#| eval: true
+#| warning: false
 #| message: false
-#| warning: false
-#| label: tbl-voters-intention-to-vote-for-trump
-#| tbl-cap: Voters Intention to Support Trump
-data <- cleaned_data_survey
-
-# Prepare the data with specified order
-trump_counts <- data %>%
-  mutate(consider_trump = factor(consider_trump, levels = c("Yes", "No", "Other"))) %>%
-  count(consider_trump, name = "Number of Respondents") %>%
-  mutate(`Proportion (%)` = (`Number of Respondents` / sum(`Number of Respondents`)) * 100) %>%
-  mutate(`Proportion (%)` = round(`Proportion (%)`, 2)) %>%
-  rename(`Response` = consider_trump)
-
-# Use kable to create the table
-kable(trump_counts, format = "latex", booktabs = TRUE, align = "c") %>%
-  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-voters-intention-of-their-primary-party
-#| tbl-cap: Voters Intention of Their Primary Party
-
-data <- cleaned_data_survey
-
-# Prepare the data with specified order
-party_preference_counts <- data %>%
-  mutate(primary_party = factor(primary_party, levels = c("Democratic", "Republican", "Other"))) %>%
-  count(primary_party, name = "Number of Respondents") %>%
-  mutate(`Proportion (%)` = (`Number of Respondents` / sum(`Number of Respondents`)) * 100) %>%
-  mutate(`Proportion (%)` = round(`Proportion (%)`, 2)) %>%
-  rename(`Party Preference` = primary_party)
-
-# Use kable to create the table
-kable(party_preference_counts, format = "latex", booktabs = TRUE, align = "c") %>%
-  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
-```
-
-# Model
-For our study, we employ a technique called multilevel regression with post-stratification (MRP).
-This approach involves creating a model based on a smaller data set, such as our survey data, and then extending the model's findings to a larger population.
-
-The key steps in MRP involve initially selecting a dataset for model development.
-In this case, we utilized survey data from the Voter Study Group (@citeSurveyData).
-The next step is to construct a model with this smaller dataset; here, we employed logistic regression based on the survey data, formulated as seen in equation \ref{eq:logit}.
-Following model creation, it is then applied to a broader dataset to estimate population characteristics.
-For our analysis, Census data from IPUMS (@citeIPUMS) served as this larger dataset.
-
-To predict an individual's likelihood of voting for Donald Trump, we aim to construct a logistic regression model leveraging data from the Voter Study Group (@citeSurveyData) and applying post-stratification with Census Data (@citeIPUMS).
-Given that logistic regression is suited for binary outcomes, we've introduced a variable, 'consider_trump', which assigns a 1 if the respondent indicates a plan to vote for Donald Trump, and a 0 for intentions to vote for other candidates, with 0 encompassing both "No" and "Other" responses.
-
-The logistic regression model takes the form of:
-
-\begin{equation}
-log(\frac{\hat{p}}{1 - \hat{p}}) = \beta_{0} + \beta_{1}x_{sex} +
-\beta_{2}x_{agegroup} + \beta_{3}x_{race} + \beta_{4}x_{state} + \beta_{5}x_{education}
-\label{eq:logit}
-\end{equation}
-
-In equation \ref{eq:logit}, each $\beta$ represents a coefficient determined through regression analysis.
-The variables chosen for this project are sex, age, race, education, and state.
-These were selected because sex, age, and race are proven to be reliable indicators of voting preferences.
-This decision is based on patterns such as certain states consistently favoring the Republican party, while others alternate between Democratic and Republican.
-Education was chosen over income because it provides a clearer picture of an individual’s background than income does.
-
-Once the logistic regression model is developed, we'll use the `predict()` function in R (@citeR) to apply our model to Census data (@citeIPUMS), breaking down the dataset into categories based on sex, race, age group, education level, and state.
-This will give us the likelihood of individuals within each category voting for Donald Trump.
-These predictions allow us to analyze potential outcomes like the popular vote winner or the electoral college vote distribution.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-model-coefficients
-#| tbl-cap: Coefficients from the Model
-model <- survey_model 
-coefficients <- broom.mixed::tidy(model, conf.int = TRUE)
-
-kable(coefficients, format = "latex", booktabs = TRUE, align = "c") %>%   kable_styling(latex_options = "scale_down", font_size = 7)
-```
-
-@tbl-model-coefficients shows the estimates for the coefficients that will fit into our logistic regression equation.
-These coefficients will fit into Equation \ref{eq:logit}, and were calculated using data from the Voter Study Group (@citeSurveyData).
-The table is made using `kable` from `knitr` (@citeKnitr) and is formatted using `kableExtra` (@citekableExtra).
-
-We use the `stan_glm()` function in R (@citeR) for our regression analysis, specifically because we are dealing with a binary outcome: whether a voter supports Donald Trump or not.
-The nature of our data suggests an S-shaped distribution rather than a linear one, making logistic regression a better fit than linear regression.
-This approach is advantageous, especially when paired with post-stratification, as it allows us to better represent under-represented groups in our analysis.
-For instance, despite having only 7 responses from Alaska in our survey data (@citeSurveyData), through multilevel regression and post-stratification, we can adjust this to effectively represent over 4500 individuals.
-
-However, there are limitations to our model.
-The binary outcome does not allow for consideration of third-party candidates or non-voters, although this limitation is mitigated by our focus on the main candidates.
-More critically, our model's accuracy is heavily dependent on the quality of our survey data.
-Any inaccuracies or the need for adjustments in the survey can significantly impact our findings.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-coefficient-estimates
-#| fig-cap: Coefficient Estimates
-
-coefficients %>%
-  ggplot(aes(estimate, term)) + geom_point() + geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) + labs(title = "Coefficient Estimates", x = "Estimate", y = "Coefficient") +   theme(axis.text.y = element_text(size = 5))
-```
-
-@fig-coefficient-estimates presents the coefficients derived from logistic regression on the survey data (@citeSurveyData).
-It also includes error bars, indicating the confidence interval for each coefficient estimate.
-In interpreting these coefficients, it's essential to understand that positive values suggest a greater likelihood of voting for Donald Trump, whereas negative values indicate a tendency to vote for other candidates, such as Joe Biden.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-prediction-probabilities
-#| tbl-cap: Example of Prediction model
-
-# Select only the first 10 rows for the table
-probability_subset <- head(prediction_model, 10)
-
-# Create the table with kable and format with kable_styling
-kable(probability_subset, format = "latex", booktabs = TRUE, align = "c") %>%
-  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
-```
-
-By utilizing the results from our logistic regression model, we can formulate an equation that adheres to the structure outlined in equation \ref{eq:logit}, incorporating specific $\beta$ coefficients for each variable.
-Given the number of variables, detailing the equation fully is challenging.
-Essentially, the equation integrates the $\beta$ value of a variable if an individual's characteristic matches that variable.
-@tbl-prediction-probabilities offers examples of how the probability varies based on different variables.
-
-# Results
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
 
 library(rstanarm)
-
-survey_data$consider_trump_binary <- ifelse(survey_data$consider_trump == "Yes", 1, 0)
-survey_model_now <- glm(consider_trump_binary ~ gender + education + age_group + race + state,data = survey_data,family = binomial(link = "logit"))
-
-# group by each demographic
-post_props <- post_strat_data %>% 
-  group_by(state, race, age_group, gender, education) %>% 
-  summarise(n = n(), .groups ='drop') %>% 
-  group_by(state) %>% 
-  mutate(prop = n/sum(n))
-
-# estimate of the model 
-post_props$estimate <- predict.glm(survey_model_now, newdata = post_props,
-                          type = "response")
-
-# new column with the number of voters for Trump
-post_props <- post_props %>% mutate(num_voters = n*estimate)
-
-# error terms 
-errors <- predict.glm(survey_model_now, newdata = post_props,
-                      type = "response", se.fit = T)
-lower <- errors$fit - errors$se.fit
-upper <- errors$fit + errors$se.fit
-
-# combine errors with proportions
-post_props_error <- cbind(post_props, lower, upper)
-post_props_error <- post_props_error %>% 
-  rename("lower" = ...10,
-         "upper" = ...11)
-
-# new column of voters for trump 
-post_props_error <- post_props_error %>% mutate(num_voters_lower = n*lower,
-                                num_voters_upper = n*upper)
-
-# confidence intervals for errors 
-vote_trump <- post_props_error %>% mutate(trump_prop=estimate*prop,
-                                       trump_prop_lower = lower*prop,
-                                       trump_prop_upper = upper*prop) %>%
-  group_by(state) %>%
-  summarise(trump_predict = sum(trump_prop),
-            trump_prop_lower = sum(trump_prop_lower),
-            trump_prop_upper = sum(trump_prop_upper))
 
@@
 #| echo: false
-#| message: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
-#| label: fig-distribution-by-state
-#| fig-cap: Distribution of Sample and Population by Education
-
-# state proportions for survey data
-survey_props <- survey_data %>% 
-  group_by(state, consider_trump_binary) %>% 
-  summarise(n = n()) %>% 
-  mutate(prop = n/sum(n))
-
-# combining survey data with the post-stratification data
-combined_trump <- inner_join(survey_props, vote_trump, by = "state")
-combined_trump  <- combined_trump %>% 
-  filter(consider_trump_binary == 1)
-
-stateprops_plot <- combined_trump %>% arrange(desc(trump_predict)) %>% 
-  mutate(support = ifelse(trump_predict >= 0.5, "Trump", "Biden"))
-
-# make a plot of the result for state 
-stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
-                               color = support)) +
-  geom_point(stat = "identity") +
-  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
-  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
-  labs(title = "Estimated Proportion of State Voting for Joe Biden",
-       x = "Percentage", y= "State",
-       color = "Support") +
-  scale_x_continuous(labels = scales::percent) +
-  theme(axis.text.y = element_text(size = 5)) +
-  scale_color_manual(values = c("Blue", "Red"))
-```
-
-Figure \@ref(fig:fig-distribution-by-state) shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms.
-Each dot represents the point estimate of the proportion of support for Biden (blue) or Trump (red) in each state.
-Horizontal Lines extending from the dots represent confidence intervals for these estimates.
-The length of each line indicates the uncertainty associated with each estimate.
-For instance, we can see that this uncertainty lies between 50 percent to slightly higher than 80 percent for Trump in MT (Massachusetts).
-The dashed green line in the middle at the 50 percent mark represents the threshold for majority support.
-On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest at the top to the lowest at the bottom.
-
-From figure, it seems that majority of the states support Biden.
-Only 7 states out of 51 have its point estimate greater than 50 percent for Trump.
-However, the horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans.
-However, excluding these contesting states, our model suggests that only 3 states are definitely in favor of Trump whereas
-35 states are definitely supporting Biden.
-
-```{r} 
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-distribution-by-educationlevel
-#| fig-cap: How different education levels of voters affect voting for Trump
-
-# Aggregating data by education level
-post_props_education <- post_strat_data %>% 
-  group_by(education, race, age_group, gender, state) %>%  # Adjust based on variables of interest
-  summarise(n = n(), .groups ='drop') %>%
-  group_by(education) %>%
-  mutate(prop = n/sum(n))
-
-# Generate predictions for each grouped cell based on education
-post_props_education$estimate <- predict.glm(survey_model_now, newdata = post_props_education, type = "response")
-
-# Calculate the number of voters from each cell voting for Trump
-post_props_education <- post_props_education %>% mutate(num_voters = n * estimate)
-
-# Get errors for the prediction
-errors_education <- predict.glm(survey_model_now, newdata = post_props_education,
-                                type = "response", se.fit = TRUE)
-lower_education <- errors_education$fit - errors_education$se.fit
-upper_education <- errors_education$fit + errors_education$se.fit
-
-# Combine the proportions and the errors generated
-post_props_education_error <- cbind(post_props_education, lower_education, upper_education) %>% rename(lower = ...10, upper = ...11)
-
-# Aggregate predictions by education level
-vote_trump_education <- post_props_education_error %>%
-  mutate(trump_prop = estimate * prop,
-         trump_prop_lower = lower * prop,
-         trump_prop_upper = upper * prop) %>%
-  group_by(education) %>%
-  summarise(trump_predict = sum(trump_prop),
-            trump_prop_lower = sum(trump_prop_lower),
-            trump_prop_upper = sum(trump_prop_upper))
-
-vote_trump_education %>% ggplot(aes(y = reorder(education, trump_predict), x = trump_predict)) +
-  geom_point(stat = "identity") +
-  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
-  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
-  labs(title = "Estimated Proportion of Voters for Trump by Education Level",
-       x = "Percentage", y= "Education",
-       color = "Support") +
-  scale_x_continuous(labels = scales::percent) +
-  theme(axis.text.y = element_text(size = 5)) +
-  scale_color_manual(values = c("Blue", "Red"))
 
@@
 
-Figure \@ref(fig:fig-distribution-by-educationlevel) presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
-Each black dot represents the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for Trump.
-The horizontal lines extending to the left and right of each dot represent the confidence intervals around the estimate, which reflect the uncertainty.
+# Discussion
 
-It shows that regardless of education level, the level of support for Trump lies below 40 percent.
-The Republican party does not have majority, including the error bars across all education levels.
-Voters with "High school or less" education appear to have the lowest estimated support for Trump, which does not align with various exit polls and analyses from the 2020 election suggesting that Trump had substantial support among voters without a college degree.
-Conversely, Voters with 'Some college' and 'Postgrad' education are the two groups that are more in support of Trump, which is exactly the opposite of what we have expected.
+## Analyzing Voter Preferences Across Demographics
 
-# Discussion
+The analysis reveals critical insights into how education, age, race influence voting behavior.
+Contrary to conventional wisdom and past electoral analyses, our findings indicate a divergent pattern in voting behavior among different educational groups.
+The shift in party allegiance among non-college white voters from Democrats to Republicans and the nuanced preferences of voters with 'Some college' and 'Postgrad' education levels challenge the stereotypical narratives that educational attainment leads to an increase in support for the Democrats.
+This group, especially highly educated 'Postgrad' group being the second highest supporters suggest a reevaluation of Democrats' strategies to appeal to educated voters concerned with economic policies and national security.
+However, according to our data, the Republicans fail to attract voters as highest proportion of support among the four educational groups remain under 40 percent.
+
+@tbl-voters-intention-to-vote-for-trump suggests that
+
+shows that majority of Biden's support
+
+The increasing propensity of older voters to lean towards the Democratic nominee, potentially driven by concerns over COVID-19 and healthcare, marks a significant shift in voting patterns.
+Similarly, the racial dynamics observed, with non-white demographics showing strong support for the Democratic party, reflect broader national conversations on race, immigration, and identity.
+
+## Geographical Divides and Electoral Preferences
+
+From @fig-distribution-by-state, we can observe that Trump leads in mid-west of America, Montana (MT), Wyoming (WY), Mississippi (MS), Maine (ME), Missouri (MO), Kentucky (KY), and Utah (UT).
+Even the competing states like Louisiana (LA), New Mexico (NM), Tennessee (TN) are also mid-west part of the country, which are rural areas with low population density.
+This is correctly estimated since in the real world as Trump won in these states: MT, WY, MS, MA, MI, KY, UT.
+He also won in LA and TN, which we considered were competing states.
+We can go more in depth of this study with the benefit of hindsight.
+We estimated that the biggest victory for Trump would be Montana (MT) with just below 70%.
+In reality, he won by 59.6%, which is lower than we thought but still sit within our error values.
+The next state with the highest proportion of support for Trump was Wyoming (WY) from our model with 65%, which is close to the election poll result of 69%.
+We were very successful with Mississippi (MS) with an estimation of 57% and the poll result of 57.5%.
+Maine (ME) was another very close estimate with 52% and the poll result of 53.1%.
+56.8% voted for Trump in Missouri (MO), which was slightly higher than our estimation of 51%.
+We estimated that only 26% of voters in Hawaii (HI) would vote for Trump with error margins of +-10%, which matched 34.3% for Trump in the election.
+41.9% for Trump in Colorado (CO) lie within our estimate of 37% with +-5% margins.
+
+There were some cases where our estimation was incorrect in terms of who won the state but the results lie within our error margins.
+For instance, we estimated Utah (UT) in favor of the Democrats with a slight margin of 51% for Biden but Biden only received 37.6% in the election.
+This still lies within our error values that we estimated to be from 43% to 58%.
+However, there were some errors in our estimation where our model failed to predict the Trump winning in the number of states in general.
+Our model predicted Louisiana (LA) for Biden with 51% but he only received 39.9% of the votes.
+This value is lower than our error margins.
+62% of the residents in Alabama, a Republican stronghold in the last eleven presidential elections had voted for Trump while our model estimated 45% of votes for Trump.
+The republicans won in Florida (FL) with a slight margin of 51.2% but we predicted an estimation of 45% with +-2% margins.
+What we estimated to be the strongest Democrat state was North Dakota (ND) with an estimation of only 23% with +-12% margins but in reality 65.1% of the residents voted for Trump.
+
+We were able to estimate the winning of states with some success
+
+## Weaknesses and Implications for Future Research
+
+Weaknesses and next steps should also be included.
 

### 2024-03-14 paper/paper.qmd
update discussion

@@
 cleaned_data_survey <- read_csv(file = here("data/analysis_data/ns20191003_cleaned.csv"), show_col_types = FALSE)
-
-survey_model <- readRDS(file = here("models/consider_trump.rds"))
-predicton_model <- readRDS(file = here("models/census_data_with_predictions.rds"))
 ```

### 2024-03-15 paper/paper.qmd
update

@@
 
-## Analyzing Voter Preferences Across Demographics
+## Analyzing Voter Preferences Across Education, Race, Gender AND Age
 
-The analysis reveals critical insights into how education, age, race influence voting behavior.
+The analysis reveals critical insights into how education, race, gender and age influence voting behavior.
 Contrary to conventional wisdom and past electoral analyses, our findings indicate a divergent pattern in voting behavior among different educational groups.
-The shift in party allegiance among non-college white voters from Democrats to Republicans and the nuanced preferences of voters with 'Some college' and 'Postgrad' education levels challenge the stereotypical narratives that educational attainment leads to an increase in support for the Democrats.
+The shift in party allegiance among non-college voters from Republicans to Democrats and the the dominance of popularity of the Republicans with 'Some college' and 'Postgrad' education levels challenge the stereotypical narratives that educational attainment leads to an increase in support for the Democrats.
 This group, especially highly educated 'Postgrad' group being the second highest supporters suggest a reevaluation of Democrats' strategies to appeal to educated voters concerned with economic policies and national security.
-However, according to our data, the Republicans fail to attract voters as highest proportion of support among the four educational groups remain under 40 percent.
+However, according to our analysis, the Republicans fail to attract voters as highest proportion of support among the four educational groups remain under 40 percent.
+This almost certainly predicts that Joey Biden will become the president of the United States.
+This coincides with the election results where Biden won 55% of the votes from the college graduates while Trump gained 43%.
+Our analysis of 37% of voters college degree and 37.5% of postgrad suggest are in line with the results.
+However, this model has failed to estimate the winning of Trump from less educated group since we estimated 39% from voters with 'some college degree' and 36.5% from voters with 'High school or less'
+
++ race
+
++ gender: American woman for the Democratic party over the last 40 years [citation].
+
++ age In general, younger people tend to be more liberal and as people get older they become more conservative.
+From our analysis, we observe a trend where ''.
+36% of '18-29' group, 46% of '30-44', 50% of 45-64 group, 52% of '+65' group voted for trump respectively in the last election (washington post + add citation).
+We see a continuous increase in favour of Trump as people get older.
 
@@
 41.9% for Trump in Colorado (CO) lie within our estimate of 37% with +-5% margins.
++ how blue-collar workers became dissatisfied with the democratic government (wilson)
 
@@
 
-We were able to estimate the winning of states with some success
+Big cities and large suburban areas along the west coast and the east coast favored Biden whereas Trump supporters are mostly in the rural areas in the mid-south and mid-north states.
+
+Historically, the Republicans
+
+We were able to estimate the winning of states with some success.
+From the data and analysis outlined above, we believe that our model suggests that Biden will win both the popular vote and the electoral votes and become the president of the United States.
+This was partially right in that he did win both but with a smaller margin that we had expected.
+In the end, he won 306 electoral votes and won 51.3% of the votes to become the 46th US president.
 
 ## Weaknesses and Implications for Future Research
-
-Weaknesses and next steps should also be included.
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
 
@@
 
-# Appendix {-}
+# Appendix
 

### 2024-03-15 paper/paper.qmd
update on results/discussion

@@
   - Jiwon Choi
-thanks: "Code and data are available at: https://github.com/Kjeongwoo99/STA302H_Paper3"
+thanks: "Code and data are available at: https://github.com/Kjeongwoo99/2020_US_Election_Forecast"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "In this study, we forecasted the 2020 U.S.
+Presidential Election outcomes using a sophisticated statistical technique called multilevel regression with post-stratification (MRP), applied to a large-scale national survey dataset and census data.
+Our findings reveal how demographic factors such as age, gender, education, race, and geographic location influence voting behavior, with particular emphasis on voter preferences for Donald Trump.
+The analysis highlights significant variations in voter support across different states and educational backgrounds, underscoring the complex interplay of demographics in shaping electoral outcomes.
+This research contributes to our understanding of the American electoral landscape, demonstrating the critical role of demographic diversity in determining the direction of political preferences and election results."
 format: pdf
@@
 cleaned_data_survey <- read_csv(file = here("data/analysis_data/ns20191003_cleaned.csv"), show_col_types = FALSE)
+survey_model <- readRDS(file = here("models/consider_trump.rds"))
+prediction_model <- readRDS(file = here("models/census_data_with_predictions.rds"))
 ```
@@
 
-This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from from July 2019 to February 2021 @citeSurveyData.
+This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from from July 2019 to February 2021 (@citeSurveyData).
 The sample is weighted to represent the U.S.
-adult population @citeSurveyData.
-Nationscape groups weight on the following important factors: gender, the four major census regions, race, Hispanic ethnicity, household income, education, age, language spoken at home, nativity, 2016 presidential vote, and the urban-rural mix of the respondent's ZIP code @citeSurveyData.
-According to the data, Male make up 48.3% while female make up 51.3% @citeSurveyData.
-74.2% of the respondents are White, 6.8% are Asian/Pacific, 12% are Black @citeSurveyData.
-20.4% are those between 18-29, 33.4% are 30-49, 32.4% are 50-69, 3.3% are 70+ @citeSurveyData.
+adult population (@citeSurveyData).
+Nationscape groups weight on the following important factors: gender, the four major census regions, race, Hispanic ethnicity, household income, education, age, language spoken at home, nativity, 2016 presidential vote, and the urban-rural mix of the respondent's ZIP code (@citeSurveyData).
+According to the data, Male make up 48.3% while female make up 51.3% (@citeSurveyData).
+74.2% of the respondents are White, 6.8% are Asian/Pacific, 12% are Black (@citeSurveyData).
+20.4% are those between 18-29, 33.4% are 30-49, 32.4% are 50-69, 3.3% are 70+ (@citeSurveyData).
 On average, 5.1 percent declined immediately among those who are selected for the survey.
@@
 Education is also an interesting factor.
-In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans @Harris.
-However, there has been a switch in this trend as 61 percent of non-college white voters showed their support wheres just 45 percent of college-educated white voters did in the exit polls @Harris.
-Only 37 percent of those without a degree cast their votes for Democrats while 53 percent with a degree did so @Harris.
+In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans (@Harris).
+However, there has been a switch in this trend as 61 percent of non-college white voters showed their support wheres just 45 percent of college-educated white voters did in the exit polls (@Harris).
+Only 37 percent of those without a degree cast their votes for Democrats while 53 percent with a degree did so (@Harris).
 We categorized education into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
@@
 Race also needs some attention because normally non-white groups are highly in favour of Democrats regardless of candidates and white swing by depending on candidates.
-According to the statistics collected in 2016, 93% of black, 71% of Latino, 68% of Asian support democrats while only 41% of white support democrats @Prokop.
+According to the statistics collected in 2016, 93% of black, 71% of Latino, 68% of Asian support democrats while only 41% of white support democrats (@Prokop).
 As white voters make up 74% of the voting population, it is really important for both parties to attain this demographic group.
@@
 
-The variables we have decided to use in this analysis from the dataset are 'sex', 'race', 'stateicp', 'age_group', and 'educd'.
-We have filtered out the respondents who have answered 'other' or no available sex data from the raw dataset for simplicity of the analysis.
-Therefore, 'Sex' is categorized into 'Male' and 'Female'.
-'Race' is divided into 'White', 'Black', 'Asian', 'American Indian' and 'Other'.
-As mentioned earlier, the fact that white, black and Asian together make up around 93 percent of the U.S population, this categorization is justified.
-'stateicp' is a record of all the states in the U.S.
-We have used the abbreviations for each state, for example, 'CT' for Connecticut, 'ME' for Maine, 'MA' for Massachusetts.
-We have the data of 55 states in total.
-'age_group' categorized the respondents' age into four groups: '18-29', '30-49', '50-64' and '65+'.
-'education' is divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
-We filtered out those unknowns, and cleaned the dataset by assigning each respondent to the right category accordingly.
+In processing the raw post-stratification dataset, which initially contained approximately 3.2 million records, we refined it down to about 2.3 million records.
+This was achieved through a meticulous selection process, ensuring the data's integrity and relevance for our analysis.
+In our analysis, we've selected the variables 'sex', 'race', 'stateicp', 'age', and 'educd' from the dataset.
+To simplify our analysis, respondents who indicated 'other' or provided no data for their sex have been excluded.
+Consequently, 'sex' has been categorized strictly as 'Male' and 'Female'.
+We've refined 'race' into five categories: 'White', 'Black', 'Asian', 'American Indian', and 'Other', based on the composition of the U.S.
+population, with White, Black, and Asian categories accounting for approximately 93 percent of the total.
+
+The 'stateicp' variable encompasses all U.S.
+states, using their standard abbreviations (e.g., 'CT' for Connecticut), and extends to 55 values to include 'Puerto Rico', 'State groupings (1980 Urban/rural sample)', 'Military/Military Reservations', 'District of Columbia', and an 'State not identified' category.
+Age has been grouped into four categories: '18-29', '30-49', '50-64', and '70+'.
+For educational attainment ('educd'), we've created four categories: 'High school or less', 'Some college', 'College degree', and 'Postgrad'.
+
+We excluded any unknown responses to ensure clarity and accuracy in categorization and to enhance clarity and align with survey data, we've renamed 'sex', 'stateicp', 'age', and 'educd' to 'gender', 'state', 'age_group', and 'education', respectively.
+This restructuring aims to streamline our analysis by ensuring each respondent is accurately categorized.
+
+@fig-distribution-by-age-group, @fig-distribution-by-gender, @fig-distribution-by-race, and @fig-distribution-by-education illustrate comparisons between survey data and post-stratification data across different variables.
+In these visuals, orange bars represent post-stratification data, while green bars signify survey data.
+The percentages displayed on each figure are rounded to the nearest tenth, introducing a potential margin of error of ±0.1% in the total values.
+Generally, the survey data aligns closely with the post-stratification data, maintaining a discrepancy of about 10% across most categories.
+Notable exceptions are observed in the '30-49' age group and the 'Some college' education level, where the differences exceed this margin.
+
+```{r}
+# insert figure for state here
+```
 
@@
 
+```{r fig.width=10, fig.height=5}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-gender
+#| fig-cap: Distribution of Sample and Population by Gender
+
+# Read in the survey data
+survey_data <- cleaned_data_survey
+
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+
+# Calculate percentages for the survey data by gender
+survey_gender <- survey_data %>%
+  group_by(gender) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Calculate percentages for the post-strat data by gender
+post_strat_gender <- post_strat_data %>%
+  group_by(gender) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+
+# Combine the gender percentages into a single data frame for plotting
+combined_gender_percentages <- rbind(
+  data.frame(dataset = 'Survey', gender = survey_gender$gender, percentage = survey_gender$percentage),
+  data.frame(dataset = 'Post-Strat', gender = post_strat_gender$gender, percentage = post_strat_gender$percentage)
+)
+
+# Plot the gender percentages with specified bar width and percentage on y-axis, and label each bar
+ggplot(combined_gender_percentages, aes(x = gender, y = percentage, fill = dataset)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
+  geom_text(aes(label = sprintf("%.1f%%", percentage)),
+            position = position_dodge(width = 0.8), 
+            vjust = -0.3, # Adjust text position
+            color = "black",
+            size = 3) + 
+  labs(x = "Gender", y = "Proportion", title = "Distribution of Sample and Population by Gender") +
+  theme_minimal() +
+  scale_fill_manual(values = c("orange", "green"))
+```
+
+```{r fig.width=10, fig.height=5}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-race
+#| fig-cap: Distribution of Sample and Population by Race Ethnicity
+# Read in the survey data
+survey_data <- cleaned_data_survey
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+# Calculate percentages for the survey data by race
+survey_race <- survey_data %>%
+  group_by(race) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+# Calculate percentages for the post-strat data by race
+post_strat_race <- post_strat_data %>%
+  group_by(race) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+# Combine the race percentages into a single data frame for plotting
+combined_race_percentages <- rbind(
+  data.frame(dataset = 'Survey', race = survey_race$race, percentage = survey_race$percentage),
+  data.frame(dataset = 'Post-Strat', race = post_strat_race$race, percentage = post_strat_race$percentage)
+)
+# Adjust the factor levels for the 'race' column to specify the order
+combined_race_percentages$race <- factor(combined_race_percentages$race, 
+                                                    levels = c("White", "Black", "Asian", "American Indian", "Other"))
+# Plot the race percentages with specified bar width and percentage on y-axis, and label each bar
+ggplot(combined_race_percentages, aes(x = race, y = percentage, fill = dataset)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
+  geom_text(aes(label = sprintf("%.1f%%", percentage)),
+            position = position_dodge(width = 0.8), 
+            vjust = -0.3, # Adjust text position
+            color = "black",
+            size = 3) + 
+  labs(x = "Race", y = "Proportion", title = "Distribution of Sample and Population by Race Ethnicity") +
+  theme_minimal() +
+  scale_fill_manual(values = c("orange", "green"))
+```
+
+```{r fig.width=10, fig.height=5}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-education
+#| fig-cap: Distribution of Sample and Population by Education
+# Read in the survey data
+survey_data <- cleaned_data_survey
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+# Calculate percentages for the survey data by education
+survey_education <- survey_data %>%
+  group_by(education) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+# Calculate percentages for the post-strat data by education
+post_strat_education <- post_strat_data %>%
+  group_by(education) %>%
+  summarise(count = n()) %>%
+  mutate(percentage = (count / sum(count)) * 100)  # Calculate percentage
+# Combine the education percentages into a single data frame for plotting
+combined_education_percentages <- rbind(
+  data.frame(dataset = 'Survey', education = survey_education$education, percentage = survey_education$percentage),
+  data.frame(dataset = 'Post-Strat', education = post_strat_education$education, percentage = post_strat_education$percentage)
+)
+# Adjust the factor levels for the 'education' column to specify the order
+combined_education_percentages$education <- factor(combined_education_percentages$education, 
+                                                    levels = c("High school or less", "Some college", "College degree", "Postgrad"))
+# Plot the education percentages with the adjusted order
+ggplot(combined_education_percentages, aes(x = education, y = percentage, fill = dataset)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
+  geom_text(aes(label = sprintf("%.1f%%", percentage)),
+            position = position_dodge(width = 0.8), 
+            vjust = -0.3, # Adjust text position
+            color = "black",
+            size = 3) + 
+  labs(x = "Education", y = "Proportion", title = "Distribution of Sample and Population by Education") +
+  theme_minimal() +
+  scale_fill_manual(values = c("orange", "green"))
+```
+
+@tbl-voters-intention-to-vote-for-trump shows the proportion of voters who intend to vote for Donald Trump or not.
+Also, @tbl-voters-intention-of-their-primary-party shows the proportion of voters' supporting party.
+The data used to create these table is from the @citeSurveyData.
+We see that Donald Trump and his party Republican are not expected to win the popular vote before we implement the model.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-voters-intention-to-vote-for-trump
+#| tbl-cap: Voters Intention to Support Trump
+data <- cleaned_data_survey
+
+# Prepare the data with specified order
+trump_counts <- data %>%
+  mutate(consider_trump = factor(consider_trump, levels = c("Yes", "No", "Other"))) %>%
+  count(consider_trump, name = "Number of Respondents") %>%
+  mutate(`Proportion (%)` = (`Number of Respondents` / sum(`Number of Respondents`)) * 100) %>%
+  mutate(`Proportion (%)` = round(`Proportion (%)`, 2)) %>%
+  rename(`Response` = consider_trump)
+
+# Use kable to create the table
+kable(trump_counts, format = "latex", booktabs = TRUE, align = "c") %>%
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-voters-intention-of-their-primary-party
+#| tbl-cap: Voters Intention of Their Primary Party
+
+data <- cleaned_data_survey
+
+# Prepare the data with specified order
+party_preference_counts <- data %>%
+  mutate(primary_party = factor(primary_party, levels = c("Democratic", "Republican", "Other"))) %>%
+  count(primary_party, name = "Number of Respondents") %>%
+  mutate(`Proportion (%)` = (`Number of Respondents` / sum(`Number of Respondents`)) * 100) %>%
+  mutate(`Proportion (%)` = round(`Proportion (%)`, 2)) %>%
+  rename(`Party Preference` = primary_party)
+
+# Use kable to create the table
+kable(party_preference_counts, format = "latex", booktabs = TRUE, align = "c") %>%
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
+```
+
 # Model
+For our study, we employ a technique called multilevel regression with post-stratification (MRP).
+This approach involves creating a model based on a smaller data set, such as our survey data, and then extending the model's findings to a larger population.
 
-The goal of our modelling strategy is twofold. Firstly,...
+The key steps in MRP involve initially selecting a dataset for model development.
+In this case, we utilized survey data from the Voter Study Group (@citeSurveyData).
+The next step is to construct a model with this smaller dataset; here, we employed logistic regression based on the survey data, formulated as seen in equation \ref{eq:logit}.
+Following model creation, it is then applied to a broader dataset to estimate population characteristics.
+For our analysis, Census data from IPUMS (@citeIPUMS) served as this larger dataset.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+To predict an individual's likelihood of voting for Donald Trump, we aim to construct a logistic regression model leveraging data from the Voter Study Group (@citeSurveyData) and applying post-stratification with Census Data (@citeIPUMS).
+Given that logistic regression is suited for binary outcomes, we've introduced a variable, 'consider_trump', which assigns a 1 if the respondent indicates a plan to vote for Donald Trump, and a 0 for intentions to vote for other candidates, with 0 encompassing both "No" and "Other" responses.
 
-## Model set-up
+The logistic regression model takes the form of:
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+\begin{equation}
+log(\frac{\hat{p}}{1 - \hat{p}}) = \beta_{0} + \beta_{1}x_{sex} +
+\beta_{2}x_{agegroup} + \beta_{3}x_{race} + \beta_{4}x_{state} + \beta_{5}x_{education}
+\label{eq:logit}
+\end{equation}
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+In equation \ref{eq:logit}, each $\beta$ represents a coefficient determined through regression analysis.
+The variables chosen for this project are sex, age, race, education, and state.
+These were selected because sex, age, and race are proven to be reliable indicators of voting preferences.
+This decision is based on patterns such as certain states consistently favoring the Republican party, while others alternate between Democratic and Republican.
+Education was chosen over income because it provides a clearer picture of an individual’s background than income does.
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+Once the logistic regression model is developed, we'll use the `predict()` function in R (@citeR) to apply our model to Census data (@citeIPUMS), breaking down the dataset into categories based on sex, race, age group, education level, and state.
+This will give us the likelihood of individuals within each category voting for Donald Trump.
+These predictions allow us to analyze potential outcomes like the popular vote winner or the electoral college vote distribution.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-model-coefficients
+#| tbl-cap: Coefficients from the Model
+model <- survey_model 
+coefficients <- broom.mixed::tidy(model, conf.int = TRUE)
+
+kable(coefficients, format = "latex", booktabs = TRUE, align = "c") %>%   kable_styling(latex_options = "scale_down", font_size = 7)
+```
+
+@tbl-model-coefficients shows the estimates for the coefficients that will fit into our logistic regression equation.
+These coefficients will fit into Equation \ref{eq:logit}, and were calculated using data from the Voter Study Group (@citeSurveyData).
+The table is made using `kable` from `knitr` (@citeKnitr) and is formatted using `kableExtra` (@citekableExtra).
+
+We use the `stan_glm()` function in R (@citeR) for our regression analysis, specifically because we are dealing with a binary outcome: whether a voter supports Donald Trump or not.
+The nature of our data suggests an S-shaped distribution rather than a linear one, making logistic regression a better fit than linear regression.
+This approach is advantageous, especially when paired with post-stratification, as it allows us to better represent under-represented groups in our analysis.
+For instance, despite having only 7 responses from Alaska in our survey data (@citeSurveyData), through multilevel regression and post-stratification, we can adjust this to effectively represent over 4500 individuals.
+
+However, there are limitations to our model.
+The binary outcome does not allow for consideration of third-party candidates or non-voters, although this limitation is mitigated by our focus on the main candidates.
+More critically, our model's accuracy is heavily dependent on the quality of our survey data.
+Any inaccuracies or the need for adjustments in the survey can significantly impact our findings.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-coefficient-estimates
+#| fig-cap: Coefficient Estimates
+
+coefficients %>%
+  ggplot(aes(estimate, term)) + geom_point() + geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) + labs(title = "Coefficient Estimates", x = "Estimate", y = "Coefficient") +   theme(axis.text.y = element_text(size = 5))
+```
+
+@fig-coefficient-estimates presents the coefficients derived from logistic regression on the survey data (@citeSurveyData).
+It also includes error bars, indicating the confidence interval for each coefficient estimate.
+In interpreting these coefficients, it's essential to understand that positive values suggest a greater likelihood of voting for Donald Trump, whereas negative values indicate a tendency to vote for other candidates, such as Joe Biden.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-prediction-probabilities
+#| tbl-cap: Example of Prediction model
+
+# Select only the first 10 rows for the table
+probability_subset <- head(prediction_model, 10)
+
+# Create the table with kable and format with kable_styling
+kable(probability_subset, format = "latex", booktabs = TRUE, align = "c") %>%
+  kable_styling(latex_options = c("striped", "scale_down", "hold_position"), full_width = FALSE)
+```
+
+By utilizing the results from our logistic regression model, we can formulate an equation that adheres to the structure outlined in equation \ref{eq:logit}, incorporating specific $\beta$ coefficients for each variable.
+Given the number of variables, detailing the equation fully is challenging.
+Essentially, the equation integrates the $\beta$ value of a variable if an individual's characteristic matches that variable.
+@tbl-prediction-probabilities offers examples of how the probability varies based on different variables.
 
@@
 
-Our results are summarized in @tbl-modelresults.
-
 ```{r}
 #| echo: false
-#| eval: true
+#| message: false
 #| warning: false
-#| message: false
 
 library(rstanarm)
+
+survey_data$consider_trump_binary <- ifelse(survey_data$consider_trump == "Yes", 1, 0)
+survey_model_now <- glm(consider_trump_binary ~ gender + education + age_group + race + state,data = survey_data,family = binomial(link = "logit"))
+
+# group by each demographic
+post_props <- post_strat_data %>% 
+  group_by(state, race, age_group, gender, education) %>% 
+  summarise(n = n(), .groups ='drop') %>% 
+  group_by(state) %>% 
+  mutate(prop = n/sum(n))
+
+# estimate of the model 
+post_props$estimate <- predict.glm(survey_model_now, newdata = post_props,
+                          type = "response")
+
+# new column with the number of voters for Trump
+post_props <- post_props %>% mutate(num_voters = n*estimate)
+
+# error terms 
+errors <- predict.glm(survey_model_now, newdata = post_props,
+                      type = "response", se.fit = T)
+lower <- errors$fit - errors$se.fit
+upper <- errors$fit + errors$se.fit
+
+# combine errors with proportions
+post_props_error <- cbind(post_props, lower, upper)
+post_props_error <- post_props_error %>% 
+  rename("lower" = ...10,
+         "upper" = ...11)
+
+# new column of voters for trump 
+post_props_error <- post_props_error %>% mutate(num_voters_lower = n*lower,
+                                num_voters_upper = n*upper)
+
+# confidence intervals for errors 
+vote_trump <- post_props_error %>% mutate(trump_prop=estimate*prop,
+                                       trump_prop_lower = lower*prop,
+                                       trump_prop_upper = upper*prop) %>%
+  group_by(state) %>%
+  summarise(trump_predict = sum(trump_prop),
+            trump_prop_lower = sum(trump_prop_lower),
+            trump_prop_upper = sum(trump_prop_upper))
 
@@
 #| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| message: false
 #| warning: false
+#| label: fig-distribution-by-state
+#| fig-cap: Distribution of Sample and Population by Education
+
+# state proportions for survey data
+survey_props <- survey_data %>% 
+  group_by(state, consider_trump_binary) %>% 
+  summarise(n = n()) %>% 
+  mutate(prop = n/sum(n))
+
+# combining survey data with the post-stratification data
+combined_trump <- inner_join(survey_props, vote_trump, by = "state")
+combined_trump  <- combined_trump %>% 
+  filter(consider_trump_binary == 1)
+
+stateprops_plot <- combined_trump %>% arrange(desc(trump_predict)) %>% 
+  mutate(support = ifelse(trump_predict >= 0.5, "Trump", "Biden"))
+
+# make a plot of the result for state 
+stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
+                               color = support)) +
+  geom_point(stat = "identity") +
+  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
+  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
+  labs(title = "Estimated Proportion of State Voting for Joe Biden",
+       x = "Percentage", y= "State",
+       color = "Support") +
+  scale_x_continuous(labels = scales::percent) +
+  theme(axis.text.y = element_text(size = 5)) +
+  scale_color_manual(values = c("Blue", "Red"))
+```
+
+Figure \@ref(fig:fig-distribution-by-state) shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms.
+Each dot represents the point estimate of the proportion of support for Biden (blue) or Trump (red) in each state.
+Horizontal Lines extending from the dots represent confidence intervals for these estimates.
+The length of each line indicates the uncertainty associated with each estimate.
+For instance, we can see that this uncertainty lies between 50 percent to slightly higher than 80 percent for Trump in MT (Massachusetts).
+The dashed green line in the middle at the 50 percent mark represents the threshold for majority support.
+On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest at the top to the lowest at the bottom.
+
+From figure, it seems that majority of the states support Biden.
+Only 7 states out of 51 have its point estimate greater than 50 percent for Trump.
+However, the horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans.
+However, excluding these contesting states, our model suggests that only 3 states are definitely in favor of Trump whereas
+35 states are definitely supporting Biden.
+
+```{r} 
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-distribution-by-educationlevel
+#| fig-cap: How different education levels of voters affect voting for Trump
+
+# Aggregating data by education level
+post_props_education <- post_strat_data %>% 
+  group_by(education, race, age_group, gender, state) %>%  # Adjust based on variables of interest
+  summarise(n = n(), .groups ='drop') %>%
+  group_by(education) %>%
+  mutate(prop = n/sum(n))
+
+# Generate predictions for each grouped cell based on education
+post_props_education$estimate <- predict.glm(survey_model_now, newdata = post_props_education, type = "response")
+
+# Calculate the number of voters from each cell voting for Trump
+post_props_education <- post_props_education %>% mutate(num_voters = n * estimate)
+
+# Get errors for the prediction
+errors_education <- predict.glm(survey_model_now, newdata = post_props_education,
+                                type = "response", se.fit = TRUE)
+lower_education <- errors_education$fit - errors_education$se.fit
+upper_education <- errors_education$fit + errors_education$se.fit
+
+# Combine the proportions and the errors generated
+post_props_education_error <- cbind(post_props_education, lower_education, upper_education) %>% rename(lower = ...10, upper = ...11)
+
+# Aggregate predictions by education level
+vote_trump_education <- post_props_education_error %>%
+  mutate(trump_prop = estimate * prop,
+         trump_prop_lower = lower * prop,
+         trump_prop_upper = upper * prop) %>%
+  group_by(education) %>%
+  summarise(trump_predict = sum(trump_prop),
+            trump_prop_lower = sum(trump_prop_lower),
+            trump_prop_upper = sum(trump_prop_upper))
+
+vote_trump_education %>% ggplot(aes(y = reorder(education, trump_predict), x = trump_predict)) +
+  geom_point(stat = "identity") +
+  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
+  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
+  labs(title = "Estimated Proportion of Voters for Trump by Education Level",
+       x = "Percentage", y= "Education",
+       color = "Support") +
+  scale_x_continuous(labels = scales::percent) +
+  theme(axis.text.y = element_text(size = 5)) +
+  scale_color_manual(values = c("Blue", "Red"))
 
 ```
+
+Figure \@ref(fig:fig-distribution-by-educationlevel) presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
+Each black dot represents the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for Trump.
+The horizontal lines extending to the left and right of each dot represent the confidence intervals around the estimate, which reflect the uncertainty.
+
+It shows that regardless of education level, the level of support for Trump lies below 40 percent.
+The Republican party does not have majority, including the error bars across all education levels.
+Voters with "High school or less" education appear to have the lowest estimated support for Trump, which does not align with various exit polls and analyses from the 2020 election suggesting that Trump had substantial support among voters without a college degree.
+Conversely, Voters with 'Some college' and 'Postgrad' education are the two groups that are more in support of Trump, which is exactly the opposite of what we have expected.
 
@@
 
+## Weaknesses and next steps
+
+Weaknesses and next steps should also be included.
+
 \newpage
@@
 
-# Appendix
+# Appendix {.unnumbered}
 

### 2024-03-15 paper/paper.qmd
added figure for state

@@
 
-@fig-distribution-by-age-group, @fig-distribution-by-gender, @fig-distribution-by-race, and @fig-distribution-by-education illustrate comparisons between survey data and post-stratification data across different variables.
-In these visuals, orange bars represent post-stratification data, while green bars signify survey data.
+@fig-data-distribution-by-state, @fig-distribution-by-age-group, @fig-distribution-by-gender, @fig-distribution-by-race, and @fig-distribution-by-education illustrate comparisons between survey data and post-stratification data across different variables.
+In these figures, orange bars represent post-stratification data, while green bars signify survey data.
 The percentages displayed on each figure are rounded to the nearest tenth, introducing a potential margin of error of ±0.1% in the total values.
@@
 
-```{r}
-# insert figure for state here
+```{r fig.width=8.5, fig.height=6}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-data-distribution-by-state
+#| fig-cap: Distribution of Sample and Population by State
+# Read in the survey data
+survey_data <- cleaned_data_survey
+
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+
+state_survey <- survey_data %>% 
+  group_by(state) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "Survey",
+         group = "state")
+
+state_post_strat <- post_strat_data %>% 
+  group_by(state) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "Post-Strat",
+         group = "state")
+
+variables <- rbind(state_survey, state_post_strat)
+
+# plot our results, grouped by variable with each line representing
+# the proportions for the polling and post-strat data
+
+variables %>% ggplot(aes(as.factor(state), pct, group=as.factor(type), linetype = as.factor(type))) + 
+  geom_line() + 
+  facet_wrap(~group, ncol = 1, scales = "free_x") + 
+  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5)) + # Corrected theme function
+  labs(x = "State", y = "Proportion", linetype = "Data Set",
+       title = "Distribution of Sample and Population by State") +
+  scale_y_continuous(labels = scales::percent)
+
 ```
@@
 Also, @tbl-voters-intention-of-their-primary-party shows the proportion of voters' supporting party.
-The data used to create these table is from the @citeSurveyData.
+The data used to create these table is from the Democracy Fund + UCLA Nationscape [@citeSurveyData].
 We see that Donald Trump and his party Republican are not expected to win the popular vote before we implement the model.

### 2024-03-15 paper/paper.qmd
typo

@@
 
-This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from from July 2019 to February 2021 (@citeSurveyData).
+This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from from July 2019 to February 2021 [@citeSurveyData].
 The sample is weighted to represent the U.S.
-adult population (@citeSurveyData).
-Nationscape groups weight on the following important factors: gender, the four major census regions, race, Hispanic ethnicity, household income, education, age, language spoken at home, nativity, 2016 presidential vote, and the urban-rural mix of the respondent's ZIP code (@citeSurveyData).
-According to the data, Male make up 48.3% while female make up 51.3% (@citeSurveyData).
-74.2% of the respondents are White, 6.8% are Asian/Pacific, 12% are Black (@citeSurveyData).
-20.4% are those between 18-29, 33.4% are 30-49, 32.4% are 50-69, 3.3% are 70+ (@citeSurveyData).
+adult population [@citeSurveyData].
+Nationscape groups weight on the following important factors: gender, the four major census regions, race, Hispanic ethnicity, household income, education, age, language spoken at home, nativity, 2016 presidential vote, and the urban-rural mix of the respondent's ZIP code [@citeSurveyData].
+According to the data, Male make up 48.3% while female make up 51.3% [@citeSurveyData].
+74.2% of the respondents are White, 6.8% are Asian/Pacific, 12% are Black [@citeSurveyData].
+20.4% are those between 18-29, 33.4% are 30-49, 32.4% are 50-69, 3.3% are 70+ [@citeSurveyData].
 On average, 5.1 percent declined immediately among those who are selected for the survey.

### 2024-03-15 paper/paper.qmd
updated reference, some minor changes to the qmd.

@@
 survey_data$consider_trump_binary <- ifelse(survey_data$consider_trump == "Yes", 1, 0)
-survey_model_now <- glm(consider_trump_binary ~ gender + education + age_group + race + state,data = survey_data,family = binomial(link = "logit"))
 
@@
 # estimate of the model 
-post_props$estimate <- predict.glm(survey_model_now, newdata = post_props,
+post_props$estimate <- predict(survey_model, newdata = post_props,
                           type = "response")
@@
 # error terms 
-errors <- predict.glm(survey_model_now, newdata = post_props,
+errors <- predict(survey_model, newdata = post_props,
                       type = "response", se.fit = T)
@@
   geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
-  labs(title = "Estimated Proportion of State Voting for Joe Biden",
+  labs(title = "Estimated Proportion of State Voting for Donal Trump",
        x = "Percentage", y= "State",
@@
 
-Figure \@ref(fig:fig-distribution-by-state) shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms.
+Figure @fig-distribution-by-state shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms.
 Each dot represents the point estimate of the proportion of support for Biden (blue) or Trump (red) in each state.
-Horizontal Lines extending from the dots represent confidence intervals for these estimates.
+Horizontal lines extending from the dots represent error bars for these estimates.
 The length of each line indicates the uncertainty associated with each estimate.
-For instance, we can see that this uncertainty lies between 50 percent to slightly higher than 80 percent for Trump in MT (Massachusetts).
-The dashed green line in the middle at the 50 percent mark represents the threshold for majority support.
-On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest at the top to the lowest at the bottom.
+For instance, we can see that this uncertainty lies between 55% to slightly higher than 80% for Trump in MT (Massachusetts).
+The dashed green line in the middle at the 50% mark represents the threshold for majority support.
+On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest to the lowest.
 
-From figure, it seems that majority of the states support Biden.
+From @fig-distribution-by-state, it seems that the majority of the states support Biden.
 Only 7 states out of 51 have its point estimate greater than 50 percent for Trump.
-However, the horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans.
-However, excluding these contesting states, our model suggests that only 3 states are definitely in favor of Trump whereas
-35 states are definitely supporting Biden.
+The horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans.
+Excluding these contesting states, however, our model suggests that only 3 states are definitely in favor of Trump whereas 35 states are definitely supporting Biden.
 
@@
 # Generate predictions for each grouped cell based on education
-post_props_education$estimate <- predict.glm(survey_model_now, newdata = post_props_education, type = "response")
+post_props_education$estimate <- predict(survey_model, newdata = post_props_education, type = "response")
 
@@
 # Get errors for the prediction
-errors_education <- predict.glm(survey_model_now, newdata = post_props_education,
+errors_education <- predict(survey_model, newdata = post_props_education,
                                 type = "response", se.fit = TRUE)
@@
 
-Figure \@ref(fig:fig-distribution-by-educationlevel) presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
+@fig-distribution-by-educationlevel presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
 Each black dot represents the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for Trump.
-The horizontal lines extending to the left and right of each dot represent the confidence intervals around the estimate, which reflect the uncertainty.
+The horizontal lines extending to the left and right of each dot represent the error terms around the estimate, which reflect the uncertainty.
 
-It shows that regardless of education level, the level of support for Trump lies below 40 percent.
-The Republican party does not have majority, including the error bars across all education levels.
-Voters with "High school or less" education appear to have the lowest estimated support for Trump, which does not align with various exit polls and analyses from the 2020 election suggesting that Trump had substantial support among voters without a college degree.
+It shows that regardless of education level, the level of support for Trump lies below 40% The Republican party does not have the majority, including the error bars across all education levels.
+Voters with "High school or less" education level appear to have the lowest estimated support for Trump, which does not align with various exit polls and analyses from the 2020 election suggesting that Trump had substantial support among voters without a college degree.
 Conversely, Voters with 'Some college' and 'Postgrad' education are the two groups that are more in support of Trump, which is exactly the opposite of what we have expected.
@@
 This group, especially highly educated 'Postgrad' group being the second highest supporters suggest a reevaluation of Democrats' strategies to appeal to educated voters concerned with economic policies and national security.
-However, according to our analysis, the Republicans fail to attract voters as highest proportion of support among the four educational groups remain under 40 percent.
-This almost certainly predicts that Joey Biden will become the president of the United States.
-This coincides with the election results where Biden won 55% of the votes from the college graduates while Trump gained 43%.
+However, according to our analysis, the Republicans fail to attract voters as highest proportion of support among the four educational groups remain under 40%.
+This almost certainly predicts that Joe Biden will become the president of the United States.
+This coincides with the election results where Biden won 55% of the votes from the college graduates while Trump gained 43% [@theWashingtonPost].
 Our analysis of 37% of voters college degree and 37.5% of postgrad suggest are in line with the results.
-However, this model has failed to estimate the winning of Trump from less educated group since we estimated 39% from voters with 'some college degree' and 36.5% from voters with 'High school or less'
+However, this model has failed to estimate the winning of Trump from less educated group since we estimated 39% from voters with 'some college degree' and 36.5% from voters with 'High school or less' whereas the post-election results show 50% for Trump from 'Some college or less' [@theWashingtonPost].
 

### 2024-03-15 paper/paper.qmd
edited discussion, added reference

@@
 
-+ gender: American woman for the Democratic party over the last 40 years [citation].
+@tbl-model-coefficients shows how males were more likely than females to have a favorable view of Trump, which correctly captured the tendency of voters by sex since 53% of men favored Trump whereas 42% of women favored Trump [@theWashingtonPost].
+Gender does not act as a uniform lens through which policy positions are viewed, however, it is still an important issue especially regarding Trump in this election due to his words and actions that may viewed as sexist.
+Historically, women have been leaning more liberal than men.
+This gender gap has varied across different elections but has remained a consistent feature of the American political landscape since the 1980s [@mary].
+While men's preferences of a smaller government remained stable, women's preference for a bigger government increased [@pewresearchcenter].
+This shift contributed to a widening gap in presidential job approval with men remaining more supportive attitude of Trump.
+Solanas (2018) suggests that Trump's adminstrations increased investment in health, education, and security for girls, teenagers, migrant women, and support measures for pregnant women, as well as the promotion of women in STEM fields and efforts against gender violence.
+However, this is no secret that Trump is very conservative in treating gender equality.
+His openly sexist comments and behaviours presaged potential limitations on women's rights and freedoms, which may have contributed to the gender gap further [@elcano].
 
-+ age In general, younger people tend to be more liberal and as people get older they become more conservative.
+In general, younger people tend to be more liberal and as people get older they become more conservative.
 From our analysis, we observe a trend where ''.
-36% of '18-29' group, 46% of '30-44', 50% of 45-64 group, 52% of '+65' group voted for trump respectively in the last election (washington post + add citation).
+36% of '18-29' group, 46% of '30-44', 50% of 45-64 group, 52% of '+65' group voted for trump respectively in the last election [@theWashingtonPost].
 We see a continuous increase in favour of Trump as people get older.
@@
 @tbl-voters-intention-to-vote-for-trump suggests that
-
-shows that majority of Biden's support
 

### 2024-03-15 paper/paper.qmd
added weakness/further research, reference

@@
 
-## Analyzing Voter Preferences Across Education, Race, Gender AND Age
+## Analyzing Voter Preferences Across Education and Gender
 
@@
 
-+ race
-
 @tbl-model-coefficients shows how males were more likely than females to have a favorable view of Trump, which correctly captured the tendency of voters by sex since 53% of men favored Trump whereas 42% of women favored Trump [@theWashingtonPost].
@@
 His openly sexist comments and behaviours presaged potential limitations on women's rights and freedoms, which may have contributed to the gender gap further [@elcano].
-
-In general, younger people tend to be more liberal and as people get older they become more conservative.
-From our analysis, we observe a trend where ''.
-36% of '18-29' group, 46% of '30-44', 50% of 45-64 group, 52% of '+65' group voted for trump respectively in the last election [@theWashingtonPost].
-We see a continuous increase in favour of Trump as people get older.
-
-@tbl-voters-intention-to-vote-for-trump suggests that
 
@@
 41.9% for Trump in Colorado (CO) lie within our estimate of 37% with +-5% margins.
-+ how blue-collar workers became dissatisfied with the democratic government (wilson)
 
@@
 What we estimated to be the strongest Democrat state was North Dakota (ND) with an estimation of only 23% with +-12% margins but in reality 65.1% of the residents voted for Trump.
+In general, big cities and large suburban areas along the west coast and the east coast favored Biden whereas Trump supporters are mostly in the rural areas in the mid-south and mid-north states from the election results.
 
-Big cities and large suburban areas along the west coast and the east coast favored Biden whereas Trump supporters are mostly in the rural areas in the mid-south and mid-north states.
+Our model was able to capture some challenges of long-held assumptions about electoral tendecy by educaiton.
+Biden improved his performance among suburban voters and White non-college voters, demonstrating a broadening appeal across different demographic groups, which is what our model suggested [@pewresearchcenter1].
+Trump maintained strong support among rural voters and White evangelicals, which underscore the importance of educational divides, with significant differences in candidate support based on education levels [@pewresearchcenter1].
+In addition, while Trump's stronghold among White men without a college degree loosened somewhat, he still won a significant majority of this demographic, which is not captured in our model [@pewresearchcenter1].
 
-Historically, the Republicans
-
-We were able to estimate the winning of states with some success.
-From the data and analysis outlined above, we believe that our model suggests that Biden will win both the popular vote and the electoral votes and become the president of the United States.
+We were able to estimate the winning states for each candidate with some success.
+From the data and analysis outlined above, we believe that our model suggests that Biden would have won both the popular vote and the electoral votes and become the president of the United States.
 This was partially right in that he did win both but with a smaller margin that we had expected.
-In the end, he won 306 electoral votes and won 51.3% of the votes to become the 46th US president.
+In the end, he won 306 electoral votes and won 51.3% of the votes to become the 46th US president [@CNN].
 
@@
 
-## Weaknesses and next steps
+One major limitation of this research is the reliance on the quality of the survey data collected in October 2019.
+Any inaccuracies  COVID-19 affected the collection of the survey data, so we had to use the 2019 data.
+The election was held in November 2020, so there exists a time gap of one year since the data collection and the election, which may have affected the voting as one year gives enough time for voters to change their mind.
+In addition, any inaccuracies or biases within the survey could significantly impact the model's forecasts, + data lost in the process.
 
-Weaknesses and next steps should also be included.
+Future research should explore models that can accommodate multiple electoral outcomes, including third-party voting and more variables if possible.
+Using more accurate or applicable real-time data sources could complement traditional survey methods and offer fresh insights into voter preferences.
+Societal norms and political landscapes change over time.
+Therefore, future research should focus on emerging social issues that may influence voter behavior.
+In the 2024 Presidential election, climate change, technology policy and social justice could play more significant roles in shaping electoral outcomes.
 

### 2024-03-15 paper/paper.qmd
added introduction

@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+The U.S.
+Presidential Election is a pivotal world political issue given its impacts on social, economic and cultural dimension across the world.
+The 2020 Election was unique in that it was held during COVID-19, characterized by unprecedented voter turnout and divided political gap between the Republicans and the Democrats.
+Joe Biden was elected as the President of the United States after 4 years of the former President Donald Trump era.
+Our research uses statistical methodologies to analyze how demographics factors such as education, state, and gender influence voter behavior and resulted in the victory of the Democratic party.
+By utilizing multilevel regression with post-stratification (MRP) to a large-scale national survey data and census information, we explain the roles of gender, education, and geographic location in shaping voter preferences, with a special focus on the support for Donald Trump.
 
-The remainder of this paper is structured as follows. @sec-data....
+Understanding the electoral outcomes is crucial, not only for politicians but for the broader public and policymakers aiming to grasp the evolving landscape of American politics.
+Traditional analyses often fall short of capturing the relationships between voter demographics and electoral preferences.
+There are growing gaps between the groups in each education, gender and state level and this study responds to this need by employing MRP, a technique that allows accurate inferences about the voting behavior of diverse population segments.
+
+Our findings reveal significant variations in voter support across different states and educational backgrounds and gender.
+The analysis points to a divergent pattern in voting behavior among different educational groups, challenging stereotypical narratives about educational attainment and party allegiance.
+We find that there is an increase in support for the Republicans from the voters with high education level in this election.
+We also find the gender gap between men and women where men are more supportive of Trump while women tend to favor Biden.
+Lastly, we discover the the strong preference for Biden from big cities in the west and the east coast while Trump's supporters are concentrated in the mid-west and suburb cities.
+
+The importance of our study extends beyond academic circles, offering valuable perspectives for political strategists, journalists, and citizens alike.
+By providing a clearer picture of the demographic factors that influenced the 2020 election, our research aids in the development of more informed political strategies and fosters a deeper understanding of American democracy.
+Furthermore, our methodology showcases the potential of advanced statistical techniques to enhance the accuracy and depth of electoral forecasts.
+
+The paper begins by introducing the broader context and motivation behind our study.
+It then discusses the specifics of the data sources and the variables that we considered important.
+This is followed by presentation of the data and discussion of our findings regarding education, gender and state in voter preference.
 

### 2024-03-15 paper/paper.qmd
introduction

@@
 Education is also an interesting factor.
-In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans (@Harris).
-However, there has been a switch in this trend as 61 percent of non-college white voters showed their support wheres just 45 percent of college-educated white voters did in the exit polls (@Harris).
-Only 37 percent of those without a degree cast their votes for Democrats while 53 percent with a degree did so (@Harris).
+In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans [@Harris].
+However, there has been a switch in this trend as 61 percent of non-college white voters showed their support wheres just 45 percent of college-educated white voters did in the exit polls [@Harris].
+Only 37 percent of those without a degree cast their votes for Democrats while 53 percent with a degree did so [@Harris].
 We categorized education into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
@@
 Race also needs some attention because normally non-white groups are highly in favour of Democrats regardless of candidates and white swing by depending on candidates.
-According to the statistics collected in 2016, 93% of black, 71% of Latino, 68% of Asian support democrats while only 41% of white support democrats (@Prokop).
+According to the statistics collected in 2016, 93% of black, 71% of Latino, 68% of Asian support democrats while only 41% of white support democrats [@Prokop].
 As white voters make up 74% of the voting population, it is really important for both parties to attain this demographic group.
@@
 These samples provide rich qualitative information on the long-term changes in the population.
-We selected the '2019 ACS' data (@citeIPUMS) as the post-stratification dataset for our research.
+We selected the '2019 ACS' data [@citeIPUMS] as the post-stratification dataset for our research.
 The ACS is an ongoing survey that collects data monthly, which is then combined into 1-year, 3-year, and 5-year aggregates.
@@
 The key steps in MRP involve initially selecting a dataset for model development.
-In this case, we utilized survey data from the Voter Study Group (@citeSurveyData).
+In this case, we utilized survey data from the Voter Study Group [@citeSurveyData].
 The next step is to construct a model with this smaller dataset; here, we employed logistic regression based on the survey data, formulated as seen in equation \ref{eq:logit}.
 Following model creation, it is then applied to a broader dataset to estimate population characteristics.
-For our analysis, Census data from IPUMS (@citeIPUMS) served as this larger dataset.
+For our analysis, Census data from IPUMS [@citeIPUMS] served as this larger dataset.
 
-To predict an individual's likelihood of voting for Donald Trump, we aim to construct a logistic regression model leveraging data from the Voter Study Group (@citeSurveyData) and applying post-stratification with Census Data (@citeIPUMS).
+To predict an individual's likelihood of voting for Donald Trump, we aim to construct a logistic regression model leveraging data from the Voter Study Group [@citeSurveyData] and applying post-stratification with Census Data [@citeIPUMS].
 Given that logistic regression is suited for binary outcomes, we've introduced a variable, 'consider_trump', which assigns a 1 if the respondent indicates a plan to vote for Donald Trump, and a 0 for intentions to vote for other candidates, with 0 encompassing both "No" and "Other" responses.
@@
 
-Once the logistic regression model is developed, we'll use the `predict()` function in R (@citeR) to apply our model to Census data (@citeIPUMS), breaking down the dataset into categories based on sex, race, age group, education level, and state.
+Once the logistic regression model is developed, we'll use the `predict()` function in R [@citeR] to apply our model to Census data [@citeIPUMS], breaking down the dataset into categories based on sex, race, age group, education level, and state.
 This will give us the likelihood of individuals within each category voting for Donald Trump.
@@
 @tbl-model-coefficients shows the estimates for the coefficients that will fit into our logistic regression equation.
-These coefficients will fit into Equation \ref{eq:logit}, and were calculated using data from the Voter Study Group (@citeSurveyData).
-The table is made using `kable` from `knitr` (@citeKnitr) and is formatted using `kableExtra` (@citekableExtra).
+These coefficients will fit into Equation \ref{eq:logit}, and were calculated using data from the Voter Study Group [@citeSurveyData].
+The table is made using `kable` from `knitr` [@citeKnitr] and is formatted using `kableExtra` [@citekableExtra].
 
-We use the `stan_glm()` function in R (@citeR) for our regression analysis, specifically because we are dealing with a binary outcome: whether a voter supports Donald Trump or not.
+We use the `stan_glm()` function in R [@citeR] for our regression analysis, specifically because we are dealing with a binary outcome: whether a voter supports Donald Trump or not.
 The nature of our data suggests an S-shaped distribution rather than a linear one, making logistic regression a better fit than linear regression.
 This approach is advantageous, especially when paired with post-stratification, as it allows us to better represent under-represented groups in our analysis.
-For instance, despite having only 7 responses from Alaska in our survey data (@citeSurveyData), through multilevel regression and post-stratification, we can adjust this to effectively represent over 4500 individuals.
+For instance, despite having only 7 responses from Alaska in our survey data [@citeSurveyData], through multilevel regression and post-stratification, we can adjust this to effectively represent over 4500 individuals.
 
@@
 
-@fig-coefficient-estimates presents the coefficients derived from logistic regression on the survey data (@citeSurveyData).
+@fig-coefficient-estimates presents the coefficients derived from logistic regression on the survey data [@citeSurveyData].
 It also includes error bars, indicating the confidence interval for each coefficient estimate.
@@
 From @fig-distribution-by-state, it seems that the majority of the states support Biden.
-Only 7 states out of 51 have its point estimate greater than 50 percent for Trump.
+Only 7 states out of 51 have its point estimate greater than 50% for Trump.
 The horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans.
@@
 56.8% voted for Trump in Missouri (MO), which was slightly higher than our estimation of 51%.
-We estimated that only 26% of voters in Hawaii (HI) would vote for Trump with error margins of +-10%, which matched 34.3% for Trump in the election.
-41.9% for Trump in Colorado (CO) lie within our estimate of 37% with +-5% margins.
+We estimated that only 26% of voters in Hawaii (HI) would vote for Trump with error margins of ±10%, which matched 34.3% for Trump in the election.
+41.9% for Trump in Colorado (CO) lie within our estimate of 37% with ±5% margins.
 
@@
 62% of the residents in Alabama, a Republican stronghold in the last eleven presidential elections had voted for Trump while our model estimated 45% of votes for Trump.
-The republicans won in Florida (FL) with a slight margin of 51.2% but we predicted an estimation of 45% with +-2% margins.
-What we estimated to be the strongest Democrat state was North Dakota (ND) with an estimation of only 23% with +-12% margins but in reality 65.1% of the residents voted for Trump.
+The republicans won in Florida (FL) with a slight margin of 51.2% but we predicted an estimation of 45% with ±2% margins.
+What we estimated to be the strongest Democrat state was North Dakota (ND) with an estimation of only 23% with ±12% margins but in reality 65.1% of the residents voted for Trump.
 In general, big cities and large suburban areas along the west coast and the east coast favored Biden whereas Trump supporters are mostly in the rural areas in the mid-south and mid-north states from the election results.

### 2024-03-15 paper/paper.qmd
update paper file

@@
 We have used two datasets for this study.
-One is the U.S election survey data of Democracy Fund + UCLA Nationscape dataset from the Voter Study Group, conducted on October 3, 2019.
-Second is the census data from IPUMS America Census Service, which is used as the post-stratification data for the survey data to adjust the weight.
+One is the U.S election survey data of Democracy Fund + UCLA Nationscape dataset from the Voter Study Group [@citeSurveyData], conducted on October 3, 2019.
+Second is the census data from IPUMS America Census Service [@citeIPUMS], which is used as the post-stratification data for the survey data to adjust the weight.
 
@@
 In general, the west and the east coasts are democrat supporters whereas south are conservative supporters.
-
-## Post-stratification Data
-
-IPUMS ("Integrated Public Use Microdata Series") is a website that offers database of samples of the American population from the American Community Surveys of 2000-present.
-These samples provide rich qualitative information on the long-term changes in the population.
-We selected the '2019 ACS' data [@citeIPUMS] as the post-stratification dataset for our research.
-The ACS is an ongoing survey that collects data monthly, which is then combined into 1-year, 3-year, and 5-year aggregates.
-It then uses stratified sampling where the U.S population is broken down into sub-groups and initial weights are assigned to each respondent.
-
-One strength of the IPUMS survey is the fact that it provides a data with detailed demographic of the U.S.
-population with social, economic and housing characteristics, which is very useful in our analysis of the 2020 U.S presidential election forecast.
-The longitudinal data of this survey also allows researchers to analyze trends over time.
-The U.S.
-Census Bureau offers credibility of the data with high quality checks.
-The post-stratification process ensures correcting for sampling biases and non-response.
-On the other hand, since the survey relies on self-report, there lies a risk of response bias inherently.
-While it is an ongoing survey, there is still a time lag between the data collection and data availability.
-However, the large sample size, consistency and reliability of the data collection, the integrated data over time with post-stratification can justify the decision to utilize IPUMS data over other sources.
-
-In processing the raw post-stratification dataset, which initially contained approximately 3.2 million records, we refined it down to about 2.3 million records.
-This was achieved through a meticulous selection process, ensuring the data's integrity and relevance for our analysis.
-In our analysis, we've selected the variables 'sex', 'race', 'stateicp', 'age', and 'educd' from the dataset.
-To simplify our analysis, respondents who indicated 'other' or provided no data for their sex have been excluded.
-Consequently, 'sex' has been categorized strictly as 'Male' and 'Female'.
-We've refined 'race' into five categories: 'White', 'Black', 'Asian', 'American Indian', and 'Other', based on the composition of the U.S.
-population, with White, Black, and Asian categories accounting for approximately 93 percent of the total.
-
-The 'stateicp' variable encompasses all U.S.
-states, using their standard abbreviations (e.g., 'CT' for Connecticut), and extends to 55 values to include 'Puerto Rico', 'State groupings (1980 Urban/rural sample)', 'Military/Military Reservations', 'District of Columbia', and an 'State not identified' category.
-Age has been grouped into four categories: '18-29', '30-49', '50-64', and '70+'.
-For educational attainment ('educd'), we've created four categories: 'High school or less', 'Some college', 'College degree', and 'Postgrad'.
-
-We excluded any unknown responses to ensure clarity and accuracy in categorization and to enhance clarity and align with survey data, we've renamed 'sex', 'stateicp', 'age', and 'educd' to 'gender', 'state', 'age_group', and 'education', respectively.
-This restructuring aims to streamline our analysis by ensuring each respondent is accurately categorized.
-
-@fig-data-distribution-by-state, @fig-distribution-by-age-group, @fig-distribution-by-gender, @fig-distribution-by-race, and @fig-distribution-by-education illustrate comparisons between survey data and post-stratification data across different variables.
-In these figures, orange bars represent post-stratification data, while green bars signify survey data.
-The percentages displayed on each figure are rounded to the nearest tenth, introducing a potential margin of error of ±0.1% in the total values.
-Generally, the survey data aligns closely with the post-stratification data, maintaining a discrepancy of about 10% across most categories.
-Notable exceptions are observed in the '30-49' age group and the 'Some college' education level, where the differences exceed this margin.
-
-```{r fig.width=8.5, fig.height=6}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-data-distribution-by-state
-#| fig-cap: Distribution of Sample and Population by State
-# Read in the survey data
-survey_data <- cleaned_data_survey
-
-# Read in the post-stratification data
-post_strat_data <- cleaned_data_post_strat
-
-state_survey <- survey_data %>% 
-  group_by(state) %>% 
-  summarise(n = n()) %>% 
-  mutate(pct = n/sum(n), type = "Survey",
-         group = "state")
-
-state_post_strat <- post_strat_data %>% 
-  group_by(state) %>% 
-  summarise(n = n()) %>% 
-  mutate(pct = n/sum(n), type = "Post-Strat",
-         group = "state")
-
-variables <- rbind(state_survey, state_post_strat)
-
-# plot our results, grouped by variable with each line representing
-# the proportions for the polling and post-strat data
-
-variables %>% ggplot(aes(as.factor(state), pct, group=as.factor(type), linetype = as.factor(type))) + 
-  geom_line() + 
-  facet_wrap(~group, ncol = 1, scales = "free_x") + 
-  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5)) + # Corrected theme function
-  labs(x = "State", y = "Proportion", linetype = "Data Set",
-       title = "Distribution of Sample and Population by State") +
-  scale_y_continuous(labels = scales::percent)
-
-```
 
@@
 
+```{r fig.width=8.5, fig.height=5}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: fig-data-distribution-by-state
+#| fig-cap: Distribution of Sample and Population by State
+# Read in the survey data
+survey_data <- cleaned_data_survey
+
+# Read in the post-stratification data
+post_strat_data <- cleaned_data_post_strat
+
+state_survey <- survey_data %>% 
+  group_by(state) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "Survey",
+         group = "state")
+
+state_post_strat <- post_strat_data %>% 
+  group_by(state) %>% 
+  summarise(n = n()) %>% 
+  mutate(pct = n/sum(n), type = "Post-Strat",
+         group = "state")
+
+variables <- rbind(state_survey, state_post_strat)
+
+# plot our results, grouped by variable with each line representing
+# the proportions for the polling and post-strat data
+
+variables %>% ggplot(aes(as.factor(state), pct, group=as.factor(type), linetype = as.factor(type))) + 
+  geom_line() + 
+  facet_wrap(~group, ncol = 1, scales = "free_x") + 
+  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5)) + # Corrected theme function
+  labs(x = "State", y = "Proportion", linetype = "Data Set",
+       title = "Distribution of Sample and Population by State") +
+  scale_y_continuous(labels = scales::percent)
+
+```
+
 ```{r fig.width=10, fig.height=5}
@@
 ```
+## Post-stratification Data
+
+IPUMS ("Integrated Public Use Microdata Series") is a website that offers database of samples of the American population from the American Community Surveys of 2000-present.
+These samples provide rich qualitative information on the long-term changes in the population.
+We selected the '2019 ACS' data [@citeIPUMS] as the post-stratification dataset for our research.
+The ACS is an ongoing survey that collects data monthly, which is then combined into 1-year, 3-year, and 5-year aggregates.
+It then uses stratified sampling where the U.S population is broken down into sub-groups and initial weights are assigned to each respondent.
+
+One strength of the IPUMS survey is the fact that it provides a data with detailed demographic of the U.S.
+population with social, economic and housing characteristics, which is very useful in our analysis of the 2020 U.S presidential election forecast.
+The longitudinal data of this survey also allows researchers to analyze trends over time.
+The U.S.
+Census Bureau offers credibility of the data with high quality checks.
+The post-stratification process ensures correcting for sampling biases and non-response.
+On the other hand, since the survey relies on self-report, there lies a risk of response bias inherently.
+While it is an ongoing survey, there is still a time lag between the data collection and data availability.
+However, the large sample size, consistency and reliability of the data collection, the integrated data over time with post-stratification can justify the decision to utilize IPUMS data over other sources.
+
+In processing the raw post-stratification dataset, which initially contained approximately 3.2 million records, we refined it down to about 2.3 million records.
+This was achieved through a meticulous selection process, ensuring the data's integrity and relevance for our analysis.
+In our analysis, we've selected the variables 'sex', 'race', 'stateicp', 'age', and 'educd' from the dataset.
+To simplify our analysis, respondents who indicated 'other' or provided no data for their sex have been excluded.
+Consequently, 'sex' has been categorized strictly as 'Male' and 'Female'.
+We've refined 'race' into five categories: 'White', 'Black', 'Asian', 'American Indian', and 'Other', based on the composition of the U.S.
+population, with White, Black, and Asian categories accounting for approximately 93 percent of the total.
+
+The 'stateicp' variable encompasses all U.S.
+states, using their standard abbreviations (e.g., 'CT' for Connecticut), and extends to 55 values to include 'Puerto Rico', 'State groupings (1980 Urban/rural sample)', 'Military/Military Reservations', 'District of Columbia', and an 'State not identified' category.
+Age has been grouped into four categories: '18-29', '30-49', '50-64', and '70+'.
+For educational attainment ('educd'), we've created four categories: 'High school or less', 'Some college', 'College degree', and 'Postgrad'.
+
+We excluded any unknown responses to ensure clarity and accuracy in categorization and to enhance clarity and align with survey data, we've renamed 'sex', 'stateicp', 'age', and 'educd' to 'gender', 'state', 'age_group', and 'education', respectively.
+This restructuring aims to streamline our analysis by ensuring each respondent is accurately categorized.
+
+All figures showed in the data section illustrate comparisons between survey data and post-stratification data across different variables.
+First, the data comparison by state shows that the correspondence between the survey and post-stratification data for each state is quite good, further emphasize the overall reliability of the survey methodology in capturing diverse demographic characteristics.
+(see @fig-data-distribution-by-state) In other four figures, orange bars represent post-stratification data, while green bars signify survey data.
+The percentages displayed on each figure are rounded to the nearest tenth, introducing a potential margin of error of ±0.1% in the total values.
+Generally, the survey data aligns closely with the post-stratification data, maintaining a discrepancy of about 10% across most categories.
+Notable exceptions are observed in the '30-49' age group and the 'Some college' education level, where the differences exceed this margin.
+(see @fig-distribution-by-age-group, @fig-distribution-by-gender, @fig-distribution-by-race, and @fig-distribution-by-education)
 
@@
 @tbl-voters-intention-to-vote-for-trump shows the proportion of voters who intend to vote for Donald Trump or not.
+Apparently, more than half of the respondents, which is 55.30% choose not to vote for Donald Trump.
 Also, @tbl-voters-intention-of-their-primary-party shows the proportion of voters' supporting party.
@@
 
-\appendix
-
-# Appendix {.unnumbered}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-```
-
-\newpage
-
 # References

### 2024-03-15 paper/paper.qmd
Update paper.qmd

--- 2024-03-15 paper/paper.qmd
@@
 # Introduction
-
 The U.S.
@@
 # Data {#sec-data}
-
 We have used two datasets for this study.
-One is the U.S election survey data of Democracy Fund + UCLA Nationscape dataset from the Voter Study Group, conducted on October 3, 2019.
-Second is the census data from IPUMS America Census Service, which is used as the post-stratification data for the survey data to adjust the weight.
+One is the U.S election survey data of Democracy Fund + UCLA Nationscape dataset from the Voter Study Group [@citeSurveyData], conducted on October 3, 2019.
+Second is the census data from IPUMS America Census Service [@citeIPUMS], which is used as the post-stratification data for the survey data to adjust the weight.
 
 ## Survey Data
-
 This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from from July 2019 to February 2021 [@citeSurveyData].
@@
 ## Post-stratification Data
-
 IPUMS ("Integrated Public Use Microdata Series") is a website that offers database of samples of the American population from the American Community Surveys of 2000-present.
@@
 population, with White, Black, and Asian categories accounting for approximately 93 percent of the total.
-
-The 'stateicp' variable encompasses all U.S.
-states, using their standard abbreviations (e.g., 'CT' for Connecticut), and extends to 55 values to include 'Puerto Rico', 'State groupings (1980 Urban/rural sample)', 'Military/Military Reservations', 'District of Columbia', and an 'State not identified' category.
-Age has been grouped into four categories: '18-29', '30-49', '50-64', and '70+'.
-For educational attainment ('educd'), we've created four categories: 'High school or less', 'Some college', 'College degree', and 'Postgrad'.
-
-We excluded any unknown responses to ensure clarity and accuracy in categorization and to enhance clarity and align with survey data, we've renamed 'sex', 'stateicp', 'age', and 'educd' to 'gender', 'state', 'age_group', and 'education', respectively.
-This restructuring aims to streamline our analysis by ensuring each respondent is accurately categorized.
-
-@fig-data-distribution-by-state, @fig-distribution-by-age-group, @fig-distribution-by-gender, @fig-distribution-by-race, and @fig-distribution-by-education illustrate comparisons between survey data and post-stratification data across different variables.
-In these figures, orange bars represent post-stratification data, while green bars signify survey data.
-The percentages displayed on each figure are rounded to the nearest tenth, introducing a potential margin of error of ±0.1% in the total values.
-Generally, the survey data aligns closely with the post-stratification data, maintaining a discrepancy of about 10% across most categories.
-Notable exceptions are observed in the '30-49' age group and the 'Some college' education level, where the differences exceed this margin.
-
-```{r fig.width=8.5, fig.height=6}
-#| message: false
-#| echo: false
-#| warning: false
-#| label: fig-data-distribution-by-state
-#| fig-cap: Distribution of Sample and Population by State
-# Read in the survey data
-survey_data <- cleaned_data_survey
-
-# Read in the post-stratification data
-post_strat_data <- cleaned_data_post_strat
-
-state_survey <- survey_data %>% 
-  group_by(state) %>% 
-  summarise(n = n()) %>% 
-  mutate(pct = n/sum(n), type = "Survey",
-         group = "state")
-
-state_post_strat <- post_strat_data %>% 
-  group_by(state) %>% 
-  summarise(n = n()) %>% 
-  mutate(pct = n/sum(n), type = "Post-Strat",
-         group = "state")
-
-variables <- rbind(state_survey, state_post_strat)
-
-# plot our results, grouped by variable with each line representing
-# the proportions for the polling and post-strat data
-
-variables %>% ggplot(aes(as.factor(state), pct, group=as.factor(type), linetype = as.factor(type))) + 
-  geom_line() + 
-  facet_wrap(~group, ncol = 1, scales = "free_x") + 
-  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5)) + # Corrected theme function
-  labs(x = "State", y = "Proportion", linetype = "Data Set",
-       title = "Distribution of Sample and Population by State") +
-  scale_y_continuous(labels = scales::percent)
-
-```
 
@@
 
+The 'stateicp' variable encompasses all U.S.
+states, using their standard abbreviations (e.g., 'CT' for Connecticut), and extends to 55 values to include 'Puerto Rico', 'State groupings (1980 Urban/rural sample)', 'Military/Military Reservations', 'District of Columbia', and an 'State not identified' category.
+Age has been grouped into four categories: '18-29', '30-49', '50-64', and '70+'.
+For educational attainment ('educd'), we've created four categories: 'High school or less', 'Some college', 'College degree', and 'Postgrad'.
+
+We excluded any unknown responses to ensure clarity and accuracy in categorization and to enhance clarity and align with survey data, we've renamed 'sex', 'stateicp', 'age', and 'educd' to 'gender', 'state', 'age_group', and 'education', respectively.
+This restructuring aims to streamline our analysis by ensuring each respondent is accurately categorized.
+
+All figures showed in the data section illustrate comparisons between survey data and post-stratification data across different variables.
+First, the data comparison by state shows that the correspondence between the survey and post-stratification data for each state is quite good, further emphasize the overall reliability of the survey methodology in capturing diverse demographic characteristics.
+(see @fig-data-distribution-by-state) In other four figures, orange bars represent post-stratification data, while green bars signify survey data.
+The percentages displayed on each figure are rounded to the nearest tenth, introducing a potential margin of error of ±0.1% in the total values.
+Generally, the survey data aligns closely with the post-stratification data, maintaining a discrepancy of about 10% across most categories.
+Notable exceptions are observed in the '30-49' age group and the 'Some college' education level, where the differences exceed this margin.
+(see @fig-distribution-by-age-group, @fig-distribution-by-gender, @fig-distribution-by-race, and @fig-distribution-by-education)
+
 @tbl-voters-intention-to-vote-for-trump shows the proportion of voters who intend to vote for Donald Trump or not.
+Apparently, more than half of the respondents, which is 55.30% choose not to vote for Donald Trump.
 Also, @tbl-voters-intention-of-their-primary-party shows the proportion of voters' supporting party.
@@
 ## Analyzing Voter Preferences Across Education and Gender
-
 The analysis reveals critical insights into how education, race, gender and age influence voting behavior.
@@
 ## Geographical Divides and Electoral Preferences
-
 From @fig-distribution-by-state, we can observe that Trump leads in mid-west of America, Montana (MT), Wyoming (WY), Mississippi (MS), Maine (ME), Missouri (MO), Kentucky (KY), and Utah (UT).
@@
 ## Weaknesses and Implications for Future Research 
-
 One major limitation of this research is the reliance on the quality of the survey data collected in October 2019.

### 2024-03-15 paper/paper.qmd
typos, errors

--- 2024-03-15 paper/paper.qmd
@@
 ---
-title: "2020 U.S. Presidential Election Forecast"
-subtitle: "How Multilevel Regression and Post-Stratification Unveil Voter Tendencies"
+title: "A Detailed Analysis of 2020 U.S. Presidential Election: A Demographic Analysis Through Multilevel Regression and Post-Stratification"
+subtitle: "The Impact of Gender, Educatoin, Geographic location on Voter Preferences"
 author: 
@@
 Presidential Election outcomes using a sophisticated statistical technique called multilevel regression with post-stratification (MRP), applied to a large-scale national survey dataset and census data.
-Our findings reveal how demographic factors such as age, gender, education, race, and geographic location influence voting behavior, with particular emphasis on voter preferences for Donald Trump.
-The analysis highlights significant variations in voter support across different states and educational backgrounds, underscoring the complex interplay of demographics in shaping electoral outcomes.
+Our findings reveal how demographic factors such as gender, education, and geographic location influence voting behavior, with particular emphasis on voter preferences for Donald Trump.
+The analysis highlights significant variations in voter support across gender, states and educational backgrounds and how they shape electoral outcomes.
 This research contributes to our understanding of the American electoral landscape, demonstrating the critical role of demographic diversity in determining the direction of political preferences and election results."
@@
 The U.S.
-Presidential Election is a pivotal world political issue given its impacts on social, economic and cultural dimension across the world.
-The 2020 Election was unique in that it was held during COVID-19, characterized by unprecedented voter turnout and divided political gap between the Republicans and the Democrats.
+Presidential Election is a significant political issue given its impacts on social, economic and cultural dimensions across the world.
+The 2020 Election was unique in that it was held during COVID-19, characterized by unprecedented voter turnout and a divided political gap between the Republicans and the Democrats.
 Joe Biden was elected as the President of the United States after 4 years of the former President Donald Trump era.
-Our research uses statistical methodologies to analyze how demographics factors such as education, state, and gender influence voter behavior and resulted in the victory of the Democratic party.
-By utilizing multilevel regression with post-stratification (MRP) to a large-scale national survey data and census information, we explain the roles of gender, education, and geographic location in shaping voter preferences, with a special focus on the support for Donald Trump.
+Our research uses statistical methodologies to analyze how demographic factors such as education, state, and gender influence voter behavior and result in the victory of the Democratic party.
+By utilizing multilevel regression with post-stratification (MRP) to large-scale national survey data and census information, we explain the roles of gender, education, and geographic location in shaping voter preferences, focusing on support for Donald Trump.
 
@@
 Traditional analyses often fall short of capturing the relationships between voter demographics and electoral preferences.
-There are growing gaps between the groups in each education, gender and state level and this study responds to this need by employing MRP, a technique that allows accurate inferences about the voting behavior of diverse population segments.
+There are growing gaps between the groups in each education, gender, and state level and this study responds to this need by employing MRP.
+This technique allows accurate inferences about the voting behavior of diverse population segments.
+The estimand of our research is the true effect of demographic factors— gender, education, and geographic location—on voter preferences for Donald Trump in the 2020 U.S.
+Presidential Election.
+This effect represents the real-world influence that the demographic variables affect the likelihood of an individual voting for Donald Trump with all other variables held constant.
+It shows the estimated shift in voter preference towards Donald Trump to one unit change in each demographic factor.
+We aim to estimate this true effect as accurately as possible using available data.
 
@@
 The analysis points to a divergent pattern in voting behavior among different educational groups, challenging stereotypical narratives about educational attainment and party allegiance.
-We find that there is an increase in support for the Republicans from the voters with high education level in this election.
+We find that there is an increase in support for the Republicans from the voters with high education levels in this election.
 We also find the gender gap between men and women where men are more supportive of Trump while women tend to favor Biden.
-Lastly, we discover the the strong preference for Biden from big cities in the west and the east coast while Trump's supporters are concentrated in the mid-west and suburb cities.
+Lastly, we discover the the strong preference for Biden from big cities in the West and the East Coast while Trump's supporters are concentrated in the mid-west and suburban cities.
 
-The importance of our study extends beyond academic circles, offering valuable perspectives for political strategists, journalists, and citizens alike.
+The importance of our study extends beyond academic, offering valuable insights for political strategists, journalists, and citizens.
 By providing a clearer picture of the demographic factors that influenced the 2020 election, our research aids in the development of more informed political strategies and fosters a deeper understanding of American democracy.
-Furthermore, our methodology showcases the potential of advanced statistical techniques to enhance the accuracy and depth of electoral forecasts.
 
@@
 It then discusses the specifics of the data sources and the variables that we considered important.
-This is followed by presentation of the data and discussion of our findings regarding education, gender and state in voter preference.
+This is followed by the presentation of the data and a discussion of our findings regarding education, gender, and state in voter preference.
 
@@
 We have used two datasets for this study.
-One is the U.S election survey data of Democracy Fund + UCLA Nationscape dataset from the Voter Study Group [@citeSurveyData], conducted on October 3, 2019.
+One is the U.S.
+election survey data of Democracy Fund + UCLA Nationscape dataset from the Voter Study Group [@citeSurveyData], conducted on October 3, 2019.
 Second is the census data from IPUMS America Census Service [@citeIPUMS], which is used as the post-stratification data for the survey data to adjust the weight.
@@
 
-This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from from July 2019 to February 2021 [@citeSurveyData].
+This survey data is an 18-month election study conducted by UCLA researchers with roughly 6250 online interviews each from July 2019 to February 2021 [@citeSurveyData].
 The sample is weighted to represent the U.S.
@@
 Nationscape groups weight on the following important factors: gender, the four major census regions, race, Hispanic ethnicity, household income, education, age, language spoken at home, nativity, 2016 presidential vote, and the urban-rural mix of the respondent's ZIP code [@citeSurveyData].
-According to the data, Male make up 48.3% while female make up 51.3% [@citeSurveyData].
-74.2% of the respondents are White, 6.8% are Asian/Pacific, 12% are Black [@citeSurveyData].
-20.4% are those between 18-29, 33.4% are 30-49, 32.4% are 50-69, 3.3% are 70+ [@citeSurveyData].
-On average, 5.1 percent declined immediately among those who are selected for the survey.
-16.7 percent of the respondents did not complete the survey.
-Another 5.9 percent were categorized as speeding or straight-line which means they completed the survey in less than 6 minutes or selected the same response for every question in the three policy question batteries.
-Leaving these out leave 72.4 percent of the original sample for the analysis.
+According to the data, males make up 48.3% while females make up 51.3% [@citeSurveyData].
+74.2% of the respondents are White, 6.8% are Asian/Pacific, and 12% are Black [@citeSurveyData].
+20.4% are those between 18-29, 33.4% are 30-49, 32.4% are 50-69, and 3.3% are 70+ [@citeSurveyData].
+On average, 5.1% declined immediately among those who were selected for the survey.
+16.7% of the respondents did not complete the survey.
+Another 5.9% were categorized as speeding or straight-line which means they completed the survey in less than 6 minutes or selected the same response for every question in the three policy question batteries [@citeSurveyData].
+Leaving these out leave 72.4% of the original sample for the analysis.
 
-The Nationscape survey's strength lies in its methodological rigor - the effectiveness in collecting large samples from the U.S.
-citizen and its weighting strategy desinged to mirror the U.S.
-adult population by including weight factors such as age, gender, race and income and more.
+The Nationscape survey's strength lies in its methodological rigor - the effectiveness of collecting large samples from U.S.
+citizens and its weighting strategy designed to mirror the U.S.
+adult population by including weight factors such as age, gender, race, and more.
 As they filter out inaccurate or missing data, it makes sure that the data collected are accurate and ensures data integrity.
-While other datasets such as the General Social Survey (GSS) and the American National Election Studies (ANES) are available, the Nationscape dataset's frequency (surveys collected every week) give it an advantage in analyzing electoral trends and shifts in real-time.
+While other datasets such as the General Social Survey (GSS) and the American National Election Studies (ANES) are available, the Nationscape dataset's frequency (surveys collected every week) gives it an advantage in analyzing electoral trends and shifts in real-time.
 Its' extensive sample size also justifies the choice of this dataset.
 
-For our analysis, we decided to focus on five demographics: age, gender, education, race and state.
+For our analysis, we decided to choose five demographics: age, gender, education, race, and state and focus on gender, education and state.
 Age is important because in general, voters tend to become more conservative as they get older.
@@
 
-Gender is also an important category because in general, men tend be more conservative and women tend to be more liberal.
+Gender is also an important category because in general, men tend to be more conservative and women tend to be more liberal.
 Recently, gender issues are growing social issues and this may affect the election, hence we wanted to explore how this affects our model.
@@
 In the past, non-college white voters used to support Democrats while college-educated white voters supported Republicans [@Harris].
-However, there has been a switch in this trend as 61 percent of non-college white voters showed their support wheres just 45 percent of college-educated white voters did in the exit polls [@Harris].
-Only 37 percent of those without a degree cast their votes for Democrats while 53 percent with a degree did so [@Harris].
-We categorized education into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
+However, there has been a switch in this trend as 61% of non-college white voters showed their support whereas just 45% of college-educated white voters did in the exit polls [@Harris].
 
@@
 
-IPUMS ("Integrated Public Use Microdata Series") is a website that offers database of samples of the American population from the American Community Surveys of 2000-present.
+IPUMS ("Integrated Public Use Microdata Series") is a website that offers a database of samples of the American population from the American Community Surveys of 2000-present.
 These samples provide rich qualitative information on the long-term changes in the population.
-We selected the '2019 ACS' data [@citeIPUMS] as the post-stratification dataset for our research.
+We selected the '2019 ACS' data [@citeIPUMS] as the post-stratification dataset for our research to avoid any potential effects of COVID-19.
 The ACS is an ongoing survey that collects data monthly, which is then combined into 1-year, 3-year, and 5-year aggregates.
@@
 All figures showed in the data section illustrate comparisons between survey data and post-stratification data across different variables.
-First, the data comparison by state shows that the correspondence between the survey and post-stratification data for each state is quite good, further emphasize the overall reliability of the survey methodology in capturing diverse demographic characteristics.
+First, the data comparison by state shows that the correspondence between the survey and post-stratification data for each state is quite accurate, further emphasizing the overall reliability of the survey methodology in capturing diverse demographic characteristics.
 (see @fig-data-distribution-by-state) In other four figures, orange bars represent post-stratification data, while green bars signify survey data.
@@
 @tbl-voters-intention-to-vote-for-trump shows the proportion of voters who intend to vote for Donald Trump or not.
-Apparently, more than half of the respondents, which is 55.30% choose not to vote for Donald Trump.
-Also, @tbl-voters-intention-of-their-primary-party shows the proportion of voters' supporting party.
+More than half of the respondents, which is 55.30% choose not to vote for Donald Trump.
+Also, @tbl-voters-intention-of-their-primary-party shows the proportion of voters' supporting each party.
 The data used to create these table is from the Democracy Fund + UCLA Nationscape [@citeSurveyData].
@@
 In this case, we utilized survey data from the Voter Study Group [@citeSurveyData].
-The next step is to construct a model with this smaller dataset; here, we employed logistic regression based on the survey data, formulated as seen in equation \ref{eq:logit}.
-Following model creation, it is then applied to a broader dataset to estimate population characteristics.
+The next step is to construct a model with this smaller dataset; here, we employed a logistic regression based on the survey data, formulated as seen in equation \ref{eq:logit}.
+It is then followed by appling post-stratification to a broader dataset to estimate population characteristics.
 For our analysis, Census data from IPUMS [@citeIPUMS] served as this larger dataset.
-
-To predict an individual's likelihood of voting for Donald Trump, we aim to construct a logistic regression model leveraging data from the Voter Study Group [@citeSurveyData] and applying post-stratification with Census Data [@citeIPUMS].
-Given that logistic regression is suited for binary outcomes, we've introduced a variable, 'consider_trump', which assigns a 1 if the respondent indicates a plan to vote for Donald Trump, and a 0 for intentions to vote for other candidates, with 0 encompassing both "No" and "Other" responses.
+Given that logistic regression is suited for binary outcomes, we've introduced a variable, 'consider_trump', which assigns a 1 if the respondent indicates a willingness to vote for Donald Trump, and a 0 for intentions to vote for other candidates, with 0 encompassing both "No" and "Other" responses.
 
@@
 \begin{equation}
-log(\frac{\hat{p}}{1 - \hat{p}}) = \beta_{0} + \beta_{1}x_{sex} +
+log(\frac{\hat{p}}{1 - \hat{p}}) = \beta_{0} + \beta_{1}x_{gender} +
 \beta_{2}x_{agegroup} + \beta_{3}x_{race} + \beta_{4}x_{state} + \beta_{5}x_{education}
@@
 In equation \ref{eq:logit}, each $\beta$ represents a coefficient determined through regression analysis.
-The variables chosen for this project are sex, age, race, education, and state.
-These were selected because sex, age, and race are proven to be reliable indicators of voting preferences.
+The variables chosen for this project are gender, age, race, education, and state.
+These were selected because gender, age, and race are proven to be reliable indicators of voting preferences.
 This decision is based on patterns such as certain states consistently favoring the Republican party, while others alternate between Democratic and Republican.
@@
 
-Once the logistic regression model is developed, we'll use the `predict()` function in R [@citeR] to apply our model to Census data [@citeIPUMS], breaking down the dataset into categories based on sex, race, age group, education level, and state.
+Once the logistic regression model is developed, we'll use the `predict()` function in R [@citeR] to apply our model to Census data [@citeIPUMS], breaking down the dataset into categories based on gender, race, age group, education level, and state.
 This will give us the likelihood of individuals within each category voting for Donald Trump.
 These predictions allow us to analyze potential outcomes like the popular vote winner or the electoral college vote distribution.
+
+We use the `stan_glm()` function in R [@citeR] for our regression analysis, specifically because we are dealing with a binary outcome: whether a voter supports Donald Trump or not.
+The nature of our data suggests an S-shaped distribution rather than a linear one, making logistic regression a better fit than linear regression.
+This approach is advantageous, especially when paired with post-stratification, as it allows us to better represent under-represented groups in our analysis.
+For instance, despite having only 7 responses from Alaska in our survey data [@citeSurveyData], through multilevel regression and post-stratification, we can adjust this to effectively represent over 4500 individuals.
+
+However, there are limitations to our model.
+The binary outcome does not allow for consideration of third-party candidates or non-voters, although this limitation is mitigated by our focus on the main candidates.
+More critically, our model's accuracy is heavily dependent on the quality of our survey data.
+Any inaccuracies or the need for adjustments in the survey can significantly impact our findings.
+
+# Results
 
@@
 
-We use the `stan_glm()` function in R [@citeR] for our regression analysis, specifically because we are dealing with a binary outcome: whether a voter supports Donald Trump or not.
-The nature of our data suggests an S-shaped distribution rather than a linear one, making logistic regression a better fit than linear regression.
-This approach is advantageous, especially when paired with post-stratification, as it allows us to better represent under-represented groups in our analysis.
-For instance, despite having only 7 responses from Alaska in our survey data [@citeSurveyData], through multilevel regression and post-stratification, we can adjust this to effectively represent over 4500 individuals.
-
-However, there are limitations to our model.
-The binary outcome does not allow for consideration of third-party candidates or non-voters, although this limitation is mitigated by our focus on the main candidates.
-More critically, our model's accuracy is heavily dependent on the quality of our survey data.
-Any inaccuracies or the need for adjustments in the survey can significantly impact our findings.
-
 ```{r}
@@
 It also includes error bars, indicating the confidence interval for each coefficient estimate.
-In interpreting these coefficients, it's essential to understand that positive values suggest a greater likelihood of voting for Donald Trump, whereas negative values indicate a tendency to vote for other candidates, such as Joe Biden.
+In interpreting these coefficients, it is essential to understand that positive values suggest a greater likelihood of voting for Donald Trump, whereas negative values indicate a tendency to vote for other candidates, such as Joe Biden.
 
@@
 @tbl-prediction-probabilities offers examples of how the probability varies based on different variables.
-
-# Results
 
@@
 stateprops_plot <- combined_trump %>% arrange(desc(trump_predict)) %>% 
-  mutate(support = ifelse(trump_predict >= 0.5, "Trump", "Biden"))
+  mutate(support = ifelse(trump_predict >= 0.5, "Yes", "No"))
 
-# make a plot of the result for state 
+# a state plot of the result
 stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
@@
        x = "Percentage", y= "State",
-       color = "Support") +
+       color = "Support Trump") +
   scale_x_continuous(labels = scales::percent) +
@@
 
-Figure @fig-distribution-by-state shows us the estimated of proportion of support for Trump and Biden by state using MRP with the inclusion of error terms.
-Each dot represents the point estimate of the proportion of support for Biden (blue) or Trump (red) in each state.
+@fig-distribution-by-state shows us the estimated of proportion of support for Trump or not by state using MRP with the inclusion of error terms.
+Each dot represents the point estimate of the proportion of support for Trump (red) or support for any other candidates including Joe Biden (blue) in each state.
 Horizontal lines extending from the dots represent error bars for these estimates.
@@
 
-From @fig-distribution-by-state, it seems that the majority of the states support Biden.
-Only 7 states out of 51 have its point estimate greater than 50% for Trump.
+From @fig-distribution-by-state, it seems that the majority of the states do not support Trump.
+Only 7 states out of 51 have a point estimate greater than 50% for Trump.
 The horizontal lines of confidence intervals of some states overlapping the green mark give some hope for the Republicans.
-Excluding these contesting states, however, our model suggests that only 3 states are definitely in favor of Trump whereas 35 states are definitely supporting Biden.
+Excluding these contesting states, however, our model suggests that only 3 states are definitely in favor of Trump whereas 35 states are not considering supporting for Trump.
 
@@
 
-@fig-distribution-by-educationlevel presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', 'Postgrad'.
+@fig-distribution-by-educationlevel presents the estimated proportion of voters for Trump by education level, divided into four categories: 'High school or less', 'Some college', 'College degree', and 'Postgrad'.
 Each black dot represents the point estimate of the proportion of voters within the corresponding education category who are predicted to vote for Trump.
@@
 It shows that regardless of education level, the level of support for Trump lies below 40% The Republican party does not have the majority, including the error bars across all education levels.
-Voters with "High school or less" education level appear to have the lowest estimated support for Trump, which does not align with various exit polls and analyses from the 2020 election suggesting that Trump had substantial support among voters without a college degree.
-Conversely, Voters with 'Some college' and 'Postgrad' education are the two groups that are more in support of Trump, which is exactly the opposite of what we have expected.
+Voters with a "High school or less" education level appear to have the lowest estimated support for Trump, which does not align with various exit polls and analyses from the 2020 election suggesting that Trump had substantial support among voters without a college degree.
+Conversely, Voters with 'Some college' and 'Postgrad' education are the two groups that are more in support of Trump, which is exactly the opposite of what we expected.
 
@@
 
-## Analyzing Voter Preferences Across Education and Gender
+## Analyzing Voter Preferences Across Education
 
@@
 
+## Voter Preferences Across Gender
+
 @tbl-model-coefficients shows how males were more likely than females to have a favorable view of Trump, which correctly captured the tendency of voters by sex since 53% of men favored Trump whereas 42% of women favored Trump [@theWashingtonPost].
@@
 His openly sexist comments and behaviours presaged potential limitations on women's rights and freedoms, which may have contributed to the gender gap further [@elcano].
-
-The increasing propensity of older voters to lean towards the Democratic nominee, potentially driven by concerns over COVID-19 and healthcare, marks a significant shift in voting patterns.
-Similarly, the racial dynamics observed, with non-white demographics showing strong support for the Democratic party, reflect broader national conversations on race, immigration, and identity.
 
@@
 In the end, he won 306 electoral votes and won 51.3% of the votes to become the 46th US president [@CNN].
+however, + add why this happened
 

### 2024-03-16 paper/paper.qmd
minor changes

@@
 #| warning: false
+#| label: fig-distribution-by-state
+#| fig-cap: Distribution of Sample and Population by State
 
@@
 # estimate of the model 
-post_props$estimate <- predict(survey_model, newdata = post_props,
-                          type = "response")
+post_props$estimate <- predict(survey_model, newdata = post_props, type = "response")
 
@@
 # error terms 
-errors <- predict(survey_model, newdata = post_props,
-                      type = "response", se.fit = T)
+errors <- predict(survey_model, newdata = post_props, type = "response", se.fit = T)
 lower <- errors$fit - errors$se.fit
@@
 post_props_error <- cbind(post_props, lower, upper)
-post_props_error <- post_props_error %>% 
-  rename("lower" = ...10,
-         "upper" = ...11)
+post_props_error <- post_props_error %>% rename("lower" = ...10, "upper" = ...11)
 
 # new column of voters for trump 
-post_props_error <- post_props_error %>% mutate(num_voters_lower = n*lower,
-                                num_voters_upper = n*upper)
+post_props_error <- post_props_error %>% 
+  mutate(num_voters_lower = n*lower, num_voters_upper = n*upper)
 
 # confidence intervals for errors 
-vote_trump <- post_props_error %>% mutate(trump_prop=estimate*prop,
+vote_trump <- post_props_error %>% 
+  mutate(trump_prop=estimate*prop, 
          trump_prop_lower = lower*prop,
@@
             trump_prop_upper = sum(trump_prop_upper))
-
-```
-
-```{r} 
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-distribution-by-state
-#| fig-cap: Distribution of Sample and Population by Education
 
@@
 combined_trump <- inner_join(survey_props, vote_trump, by = "state")
-combined_trump  <- combined_trump %>% 
-  filter(consider_trump_binary == 1)
+combined_trump  <- combined_trump %>% filter(consider_trump_binary == 1)
 
@@
 
-# a state plot of the result
-stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
-                               color = support)) +
-  geom_point(stat = "identity") +
+# define colors 
+support_colors <- c("No" = "Blue", "Yes" = "Red")
+
+# state plot
+ggplot(stateprops_plot, aes(y = reorder(state, trump_predict), 
+                            x = trump_predict, color = support)) +
+  geom_point() +
   geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
-  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
-  labs(title = "Estimated Proportion of State Voting for Donal Trump",
+  geom_vline(xintercept = 0.5, linetype = "dashed", color = "orange") +
+  scale_x_continuous(labels = scales::percent) +
+  labs(title = "Estimated Proportion of State Voting for Donald Trump",
        x = "Percentage", y= "State",
-       color = "Support Trump") +
-  scale_x_continuous(labels = scales::percent) +
-  theme(axis.text.y = element_text(size = 5)) +
+       color = "Candidate") +
+  theme(axis.text.y = element_text(size = 5), legend.position = "right") +
   scale_color_manual(values = c("Blue", "Red"))
@@
 For instance, we can see that this uncertainty lies between 55% to slightly higher than 80% for Trump in MT (Massachusetts).
-The dashed green line in the middle at the 50% mark represents the threshold for majority support.
+The dashed orange line in the middle at the 50% mark represents the threshold for majority support.
 On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest to the lowest.
@@
 
-# Aggregating data by education level
+# group by education level
 post_props_education <- post_strat_data %>% 
-  group_by(education, race, age_group, gender, state) %>%  # Adjust based on variables of interest
+  group_by(education, race, age_group, gender, state) %>%  
   summarise(n = n(), .groups ='drop') %>%
@@
 
-# Generate predictions for each grouped cell based on education
+# predictions based on education
 post_props_education$estimate <- predict(survey_model, newdata = post_props_education, type = "response")
 
-# Calculate the number of voters from each cell voting for Trump
+# number of voters for Trump
 post_props_education <- post_props_education %>% mutate(num_voters = n * estimate)
 
-# Get errors for the prediction
+# errors for the prediction
 errors_education <- predict(survey_model, newdata = post_props_education,
@@
 
-# Combine the proportions and the errors generated
-post_props_education_error <- cbind(post_props_education, lower_education, upper_education) %>% rename(lower = ...10, upper = ...11)
+# combine the proportions and the errors
+post_props_education_error <- cbind(post_props_education, lower_education, upper_education) %>% 
+  rename(lower = ...10, upper = ...11)
 
-# Aggregate predictions by education level
+# predictions by education level
 vote_trump_education <- post_props_education_error %>%
@@
 
-vote_trump_education %>% ggplot(aes(y = reorder(education, trump_predict), x = trump_predict)) +
+# education plot
+  ggplot(vote_trump_education, aes(y = reorder(education, trump_predict), 
+                                   x = trump_predict)) +
   geom_point(stat = "identity") +
-  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
+  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper, width = 0.2)) +
   geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
-  labs(title = "Estimated Proportion of Voters for Trump by Education Level",
-       x = "Percentage", y= "Education",
-       color = "Support") +
   scale_x_continuous(labels = scales::percent) +
-  theme(axis.text.y = element_text(size = 5)) +
-  scale_color_manual(values = c("Blue", "Red"))
+  labs(title = "Estimated Proportion of Voters for Trump by Education",
+       x = "Percentage", y= "Education") + 
+  theme_minimal() + theme(axis.text.y = element_text(size = 12))
 

### 2024-03-16 paper/paper.qmd
review

@@
 ---
-title: "A Detailed Analysis of 2020 U.S. Presidential Election: A Demographic Analysis Through Multilevel Regression and Post-Stratification"
+title: "A Detailed Demographic Analysis of 2020 U.S. Presidential Election Through Multilevel Regression and Post-Stratification"
 subtitle: "The Impact of Gender, Educatoin, Geographic location on Voter Preferences"
@@
   - Jiwon Choi
-thanks: "Code and data are available at: https://github.com/Kjeongwoo99/2020_US_Election_Forecast"
+thanks: "Code and data are available at: https://github.com/Kjeongwoo99/2020_US_Election_Result_Analysis"
 date: today
@@
 
-The analysis reveals critical insights into how education, race, gender and age influence voting behavior.
+The analysis reveals critical insights into how education, race, gender, and age influence voting behavior.
 Contrary to conventional wisdom and past electoral analyses, our findings indicate a divergent pattern in voting behavior among different educational groups.
-The shift in party allegiance among non-college voters from Republicans to Democrats and the the dominance of popularity of the Republicans with 'Some college' and 'Postgrad' education levels challenge the stereotypical narratives that educational attainment leads to an increase in support for the Democrats.
-This group, especially highly educated 'Postgrad' group being the second highest supporters suggest a reevaluation of Democrats' strategies to appeal to educated voters concerned with economic policies and national security.
-However, according to our analysis, the Republicans fail to attract voters as highest proportion of support among the four educational groups remain under 40%.
+The shift in party allegiance among non-college voters from Republicans to Democrats and the dominance of popularity of the Republicans with 'Some college' and 'Postgrad' education levels challenge the stereotypical narratives that educational attainment leads to an increase in support for the Democrats.
+This group, especially the highly educated 'Postgrad' group being the second highest supporters suggests a reevaluation of Democrats' strategies to appeal to educated voters concerned with economic policies and national security.
+However, according to our analysis, the Republicans failed to attract voters as the highest proportion of support among the four educational groups remains under 40%.
 This almost certainly predicts that Joe Biden will become the president of the United States.
 This coincides with the election results where Biden won 55% of the votes from the college graduates while Trump gained 43% [@theWashingtonPost].
-Our analysis of 37% of voters college degree and 37.5% of postgrad suggest are in line with the results.
-However, this model has failed to estimate the winning of Trump from less educated group since we estimated 39% from voters with 'some college degree' and 36.5% from voters with 'High school or less' whereas the post-election results show 50% for Trump from 'Some college or less' [@theWashingtonPost].
+Our analysis of 37% of voters with college degrees and 37.5% of postgrad suggests this is in line with the results.
+However, this model has failed to estimate the winning of Trump from the less educated group since we estimated 39% from voters with 'some college degree' and 36.5% from voters with 'High school or less' whereas the post-election results show 50% for Trump from 'Some college or less' [@theWashingtonPost].
 
@@
 This gender gap has varied across different elections but has remained a consistent feature of the American political landscape since the 1980s [@mary].
-While men's preferences of a smaller government remained stable, women's preference for a bigger government increased [@pewresearchcenter].
+While men's preferences for a smaller government remained stable, women's preference for a bigger government increased [@pewresearchcenter].
 This shift contributed to a widening gap in presidential job approval with men remaining more supportive attitude of Trump.
-Solanas (2018) suggests that Trump's adminstrations increased investment in health, education, and security for girls, teenagers, migrant women, and support measures for pregnant women, as well as the promotion of women in STEM fields and efforts against gender violence.
+Solanas suggests that Trump's administration increased investment in health, education, and security for girls, teenagers, and migrant women, and support measures for pregnant women, as well as the promotion of women in STEM fields and efforts against gender violence [@elcano].
 However, this is no secret that Trump is very conservative in treating gender equality.
@@
 
-From @fig-distribution-by-state, we can observe that Trump leads in mid-west of America, Montana (MT), Wyoming (WY), Mississippi (MS), Maine (ME), Missouri (MO), Kentucky (KY), and Utah (UT).
-Even the competing states like Louisiana (LA), New Mexico (NM), Tennessee (TN) are also mid-west part of the country, which are rural areas with low population density.
-This is correctly estimated since in the real world as Trump won in these states: MT, WY, MS, MA, MI, KY, UT.
+From @fig-distribution-by-state, we can observe that Trump leads in the mid-west of America, Montana (MT), Wyoming (WY), Mississippi (MS), Maine (ME), Missouri (MO), Kentucky (KY), and Utah (UT).
+Even the competing states like Louisiana (LA), New Mexico (NM), and Tennessee (TN) are also mid-west part of the country, which are rural areas with low population density.
+This is correctly estimated since in the real world where Trump won in these states: MT, WY, MS, MA, MI, KY, and UT.
 He also won in LA and TN, which we considered were competing states.
-We can go more in depth of this study with the benefit of hindsight.
+We can go more in-depth into this study with the benefit of hindsight.
 We estimated that the biggest victory for Trump would be Montana (MT) with just below 70%.
-In reality, he won by 59.6%, which is lower than we thought but still sit within our error values.
+In reality, he won by 59.6%, which is lower than we thought but still sits within our error values.
 The next state with the highest proportion of support for Trump was Wyoming (WY) from our model with 65%, which is close to the election poll result of 69%.
@@
 For instance, we estimated Utah (UT) in favor of the Democrats with a slight margin of 51% for Biden but Biden only received 37.6% in the election.
-This still lies within our error values that we estimated to be from 43% to 58%.
+This still lies within our error values which we estimated to be from 43% to 58%.
 However, there were some errors in our estimation where our model failed to predict the Trump winning in the number of states in general.
@@
 62% of the residents in Alabama, a Republican stronghold in the last eleven presidential elections had voted for Trump while our model estimated 45% of votes for Trump.
-The republicans won in Florida (FL) with a slight margin of 51.2% but we predicted an estimation of 45% with ±2% margins.
-What we estimated to be the strongest Democrat state was North Dakota (ND) with an estimation of only 23% with ±12% margins but in reality 65.1% of the residents voted for Trump.
+The Republicans won in Florida (FL) with a slight margin of 51.2% but we predicted an estimation of 45% with ±2% margins.
+What we estimated to be the strongest Democrat state was North Dakota (ND) with an estimation of only 23% with ±12% margins but in reality, 65.1% of the residents voted for Trump.
 In general, big cities and large suburban areas along the west coast and the east coast favored Biden whereas Trump supporters are mostly in the rural areas in the mid-south and mid-north states from the election results.
 
-Our model was able to capture some challenges of long-held assumptions about electoral tendecy by educaiton.
+Our model was able to capture some challenges of long-held assumptions about electoral tendency by education.
 Biden improved his performance among suburban voters and White non-college voters, demonstrating a broadening appeal across different demographic groups, which is what our model suggested [@pewresearchcenter1].
-Trump maintained strong support among rural voters and White evangelicals, which underscore the importance of educational divides, with significant differences in candidate support based on education levels [@pewresearchcenter1].
+Trump maintained strong support among rural voters and White evangelicals, which emphasizes the importance of educational divides, with significant differences in candidate support based on education levels [@pewresearchcenter1].
 In addition, while Trump's stronghold among White men without a college degree loosened somewhat, he still won a significant majority of this demographic, which is not captured in our model [@pewresearchcenter1].
@@
 From the data and analysis outlined above, we believe that our model suggests that Biden would have won both the popular vote and the electoral votes and become the president of the United States.
-This was partially right in that he did win both but with a smaller margin that we had expected.
+This was partially right in that he did win both but with a smaller margin than we had expected.
 In the end, he won 306 electoral votes and won 51.3% of the votes to become the 46th US president [@CNN].
@@
 One major limitation of this research is the reliance on the quality of the survey data collected in October 2019.
-Any inaccuracies  COVID-19 affected the collection of the survey data, so we had to use the 2019 data.
-The election was held in November 2020, so there exists a time gap of one year since the data collection and the election, which may have affected the voting as one year gives enough time for voters to change their mind.
-In addition, any inaccuracies or biases within the survey could significantly impact the model's forecasts, + data lost in the process.
+Any inaccuracies of COVID-19 affected the collection of the survey data, so we had to use the 2019 data.
+The election was held in November 2020, so there exists a time gap of one year between the data collection and the election, which may have affected the voting as one year gives enough time for voters to change their minds.
+In addition, any inaccuracies or biases within the survey could significantly impact the model's forecasts, especially, through the data cleaning process, which involved dropping observations with missing data.
+We reduced the survey dataset from 6,146 to 5,570 observations and the post-stratification dataset from 3,239,553 to 2,334,234 observations.
+This resulted in a loss of approximately 10% of the survey data and 30% of the post-stratification data.
+Such reductions could potentially impact the accuracy of our forecast results.
 
@@
 Therefore, future research should focus on emerging social issues that may influence voter behavior.
-In the 2024 Presidential election, climate change, technology policy and social justice could play more significant roles in shaping electoral outcomes.
+In the 2024 Presidential election, climate change, technology policy, and social justice could play more significant roles in shaping electoral outcomes.
 

### 2024-03-16 paper/paper.qmd
final commit!

@@
        x = "Percentage", y= "Education") + 
-  theme_minimal() + theme(axis.text.y = element_text(size = 12))
-
+  theme_minimal() +
+  theme(axis.text.y = element_text(size = 12),
+        plot.title = element_text(hjust = 2)) # Move title to the right
 ```
@@
 
-The analysis reveals critical insights into how education, race, gender, and age influence voting behavior.
+The analysis reveals critical insights into how education influence voting behavior.
 Contrary to conventional wisdom and past electoral analyses, our findings indicate a divergent pattern in voting behavior among different educational groups.
@@
 However, this model has failed to estimate the winning of Trump from the less educated group since we estimated 39% from voters with 'some college degree' and 36.5% from voters with 'High school or less' whereas the post-election results show 50% for Trump from 'Some college or less' [@theWashingtonPost].
+This is odd because the higher the education completed, the more likely an individual to vote for democratic party in general.
+One possible explanation for this is that there are some issues with the data collection where voting for Trump is underestimated as the the proportion of voters for him remain under 40% across all educational groups.
+Another possibility is that the gap between educational groups in party preference has actually been reduced from the past.
+In fact, support from white men without a college degree reduced somewhat although he still won a majority of this demographic [@pewresearchcenter1].
 
@@
 In the end, he won 306 electoral votes and won 51.3% of the votes to become the 46th US president [@CNN].
-however, + add why this happened
+However, our model underestimates the proportion of Trump voters around 8 - 10%, compared to the actual 48% of votes he got in the election.
+In the cleaning process, we had to reduce both the number of survey data and post-stratification data, that could have led to this difference.
 
@@
 
-Future research should explore models that can accommodate multiple electoral outcomes, including third-party voting and more variables if possible.
+Our analysis also involves some weaknesses of using predictive modeling to fully grasp the complexity of how elections work.
+Although our model picks up on some trends and changes in voter demographics, it also shows that predicting election outcomes accurately, particularly in states where the results could go either way, is difficult.
+The differences between what our model predicted and the actual election results underline the importance of continuously improving our models.
+This means adding more detailed data and possibly trying different methods to better understand how people vote.
+This experience has made it clear that to make our election predictions more accurate and trustworthy in the future, we need to take into account various factors, such as education levels, differences between urban and rural areas, and the evolving political scene.
+Therefore, future research should explore models that can accommodate multiple electoral outcomes, including third-party voting and more variables if possible.
 Using more accurate or applicable real-time data sources could complement traditional survey methods and offer fresh insights into voter preferences.

### 2024-03-16 paper/paper.qmd
Final commit!!!

@@
 date-format: long
-abstract: "In this study, we forecasted the 2020 U.S.
+abstract: "In this study, we analyzed the 2020 U.S.
 Presidential Election outcomes using a sophisticated statistical technique called multilevel regression with post-stratification (MRP), applied to a large-scale national survey dataset and census data.
@@
 library(palmerpenguins)
-library(ggrepel)
-library(maps)
 library(here)
 library(kableExtra)
-library(statebins)
 library(broom.mixed)
 library(knitr)
+library(rstanarm)
 
@@
 Second is the census data from IPUMS America Census Service [@citeIPUMS], which is used as the post-stratification data for the survey data to adjust the weight.
+Data was collected and analyzed using R statistical programming software [@citeR], with additional packages like tidyverse [@tidyverse], rstanarm [@rstanarm] knitr [@citeKnitr], here [@here], and many others for support.
 
@@
 
-library(rstanarm)
-
 survey_data$consider_trump_binary <- ifelse(survey_data$consider_trump == "Yes", 1, 0)

### 2024-03-20 paper/paper.qmd
updated README, abstract and introduction

@@
 abstract: "In this study, we analyzed the 2020 U.S.
-Presidential Election outcomes using a sophisticated statistical technique called multilevel regression with post-stratification (MRP), applied to a large-scale national survey dataset and census data.
+Presidential Election outcomes using multilevel regression with post-stratification (MRP), applied to a large-scale national survey dataset and census data.
 Our findings reveal how demographic factors such as gender, education, and geographic location influence voting behavior, with particular emphasis on voter preferences for Donald Trump.
@@
 The paper begins by introducing the broader context and motivation behind our study.
-It then discusses the specifics of the data sources and the variables that we considered important.
-This is followed by the presentation of the data and a discussion of our findings regarding education, gender, and state in voter preference.
+@sec-data then discusses specifics of the data sources and the variables that we considered important.
+@sec-model introduces the specifics of the logistic regression model with post-stratification methods.
+This is followed by the presentation of the data in @sec-results and discussions of our findings regarding education, gender, and state in voter preference in @sec-discussion.
 
@@
 
-# Model
+# Model {#sec-model}
 For our study, we employ a technique called multilevel regression with post-stratification (MRP).
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 

### 2024-03-20 paper/paper.qmd
small typos

@@
 title: "A Detailed Demographic Analysis of 2020 U.S. Presidential Election Through Multilevel Regression and Post-Stratification"
-subtitle: "The Impact of Gender, Educatoin, Geographic location on Voter Preferences"
+subtitle: "The Impact of Gender, Education, and Geographic Location on Voter Preferences"
 author: 
@@
 We also find the gender gap between men and women where men are more supportive of Trump while women tend to favor Biden.
-Lastly, we discover the the strong preference for Biden from big cities in the West and the East Coast while Trump's supporters are concentrated in the mid-west and suburban cities.
+Lastly, we discover the strong preference for Biden from big cities in the West and the East Coast while Trump's supporters are concentrated in the mid-west and suburban cities.
 
@@
 The next step is to construct a model with this smaller dataset; here, we employed a logistic regression based on the survey data, formulated as seen in equation \ref{eq:logit}.
-It is then followed by appling post-stratification to a broader dataset to estimate population characteristics.
+It is then followed by applying post-stratification to a broader dataset to estimate population characteristics.
 For our analysis, Census data from IPUMS [@citeIPUMS] served as this larger dataset.
@@
 This is odd because the higher the education completed, the more likely an individual to vote for democratic party in general.
-One possible explanation for this is that there are some issues with the data collection where voting for Trump is underestimated as the the proportion of voters for him remain under 40% across all educational groups.
+One possible explanation for this is that there are some issues with the data collection where voting for Trump is underestimated as the proportion of voters for him remain under 40% across all educational groups.
 Another possibility is that the gap between educational groups in party preference has actually been reduced from the past.
